### 2021-03-10 eda.Rmd
Loaded data and libraries, and displayed first several rows

---
title: "R Notebook"
output: html_notebook
---

```{r include=FALSE}
knitr::opts_chunk$set('echo'=FALSE)

library(dplyr)
library(ggplot2)

load('streetrx.rdata')
```

```{r}
head(streetrx, 15)
```

### 2021-03-10 eda.Rmd
Beginning of analysis

@@
 ---
+
+# Todo
+
+- Check for missing data across columns (including response), as well as alternate ways data might be NA
+- Perform EDA to sanity check values, check types, etc.
+
+# Changes to data
+
+- Filtered data such that `api_temp = oxymorphone`
+- Dropped `country` and `api_temp`
+- Removed observations with missing `ppm` (14)
 
@@
 library(ggplot2)
+library(lubridate)
 
 load('streetrx.rdata')
+streetrx <- streetrx %>% filter(api_temp == 'oxymorphone')
+
+# Only 1 country (USA), and we're only working with 
+drop_cols <- c('country', 'api_temp')
+streetrx <- streetrx %>% select(-any_of(drop_cols))
 ```
@@
 ```{r}
-head(streetrx, 15)
+set.seed(42)
+slice_sample(streetrx, n=20)
 ```
+
+# Initial observations (brain dump)
+
+- At least one of the response variables (`ppm`) is missing - we'll need to remove these observations
+- `yq_pdate` and `price_date` contain redundant information - possibly best to break these 2 columns apart into quarter, day, month, and year columns.
+Likely day won't be important after accounting for month and year - check a random set of same month/year pairings and see if there's a trend?
+- `city` frequently missing
+- `state` contains 59 levels: will need to look into this
+- `country` column not needed (only working with USA)
+- `source` looks frequently missing and it's not clear how many levels there are: look into
+- `api_temp`: Not needed (we're only working with oxymorphone)
+- `mgstr`: Will need to validate this - shouldn't be negative or 0 (check histogram)
+- `bulk_purchase`: Remove the text
+- `primary_reason`: Looks missing frequently
+
+# Checking Columns One-byone
+
+The big one. First, let's take a look at which fields have missing values and how frequently they occur. First, the response variable
+
+```{r}
+sum(is.na(streetrx$ppm))
+```
+
+We'll need to drop these, because a missing response provides no assistance to our model:
+
+```{r}
+streetrx <- streetrx %>% filter(!is.na(ppm))
+```
+
+Any other weird values?
+
+```{r}
+ggplot(streetrx, aes(x=ppm)) + geom_histogram(bins = 50)
+```
+
+There are some very extreme values here:
+
+```{r}
+streetrx %>% filter(ppm > 10)
+```
+
+I'm certainly not an expert on drugs.
+But based on the previous prices, these are likely erroneously recorded, or not serious entries.
+Keep an eye on and consider dropping - the corresponding points could be influential

### 2021-03-10 eda.Rmd
Eda on state

@@
 - Removed observations with missing `ppm` (14)
+- Dropped unused levels
 
@@
 streetrx <- streetrx %>% filter(api_temp == 'oxymorphone')
+
+# Drop unusued levels:
+streetrx <- droplevels(streetrx)
 
@@
 Keep an eye on and consider dropping - the corresponding points could be influential
+
+## City
+
+First, check for standard missing values
+
+```{r}
+sum(is.na(streetrx$city))
+```
+
+Some issues:
+
+- Missing city is the most common (1413 times)
+- Nonstandard casing (sometimes lowercase)
+- Numbers occasionally appear
+- State occasionally included in city
+
+Honestly this column is a bit of a mess, with many values appearing a very minimal number of times:
+
+```{r}
+table(streetrx$city, exclude = '') %>% hist()
+```
+
+Don't include for now
+
+## State

### 2021-03-10 eda.Rmd
Analysis on state

@@
 - Dropped unused levels
+- Dropped observations where `state == 'USA'`
 
@@
 ## State
+
+Missing values?
+
+```{r}
+sum(is.na(streetrx$state))
+```
+
+No - let's check a table then:
+
+```{r}
+table(streetrx$state)
+```
+
+Ok, we have 52 levels but we *should* have 50.
+D.C.
+is on there which is fine.
+The other incorrect value is a state of "USA".
+We'll probably have to drop these, or we could take a look at their cities and manually enter the state
+
+```{r}
+streetrx %>% filter(state == 'USA')
+```
+
+City is missing, region is missing - we don't have any location information on these observations: drop
+
+```{r}
+streetrx <- streetrx %>% filter(state != 'USA')
+```

### 2021-03-10 eda.Rmd
EDA on region

@@
 ```{r}
-streetrx <- streetrx %>% filter(state != 'USA')
+streetrx <- streetrx %>% filter(state != 'USA') %>% droplevels()
 ```
+
+## USA region
+
+```{r}
+table(streetrx$USA_region)
+```
+
+Let's visualize the estimated density for these 4 groups. For visualization purposes, we omit ppm greater than 10 (which is largely the entire data anyawy)
+
+```{r}
+small_ppm <- streetrx %>% filter(ppm < 10)
+ggplot(small_ppm, aes(x=ppm, color=USA_region)) + geom_density(bw=0.5)
+```
+
+They all look pretty similar actually *except* for South, which seems to be higher priced than the other on average

### 2021-03-10 eda.Rmd
Work on source EDA

@@
 
-They all look pretty similar actually *except* for South, which seems to be higher priced than the other on average
+They all look pretty similar actually *except* for South, which seems to be higher priced than the other on average.
+
+## Source
+
+Only 18 levels - let's check the table output:
+
+```{r}
+table(streetrx$source)
+```
+
+Let's recode some of these:
+
+```{r}
+# Recode "missing" values to NA
+levels(streetrx$source)[levels(streetrx$source) == ''] <- NA_character_
+streetrx$source <- recode_factor(streetrx$source, 'w'=NA_character_,
+                                 'Idk'=NA_character_)
+
+levels(streetrx$source)[grepl('bluelight', levels(streetrx$source), fixed=TRUE)] <- 'bluelight'
+levels(streetrx$source)[grepl('opiophile', levels(streetrx$source), fixed=TRUE)] <- 'opiophile'
+```
+
+There's still work to be done on this one. Ultimately about 1/3 of these values are missing - we might not be able to save it from that.

### 2021-03-11 eda.Rmd
EDA work on more variables

@@
 - Filtered data such that `api_temp = oxymorphone`
-- Dropped `country` and `api_temp`
+- Dropped `country`, `api_temp`, and 'form_temp`
 - Removed observations with missing `ppm` (14)
@@
 
-# Only 1 country (USA), and we're only working with 
-drop_cols <- c('country', 'api_temp')
+# 1 country (USA), 1 drug (oxymorphone), and always as a pill/tablet (except once)
+drop_cols <- c('country', 'api_temp', 'form_temp')
 streetrx <- streetrx %>% select(-any_of(drop_cols))
@@
 Keep an eye on and consider dropping - the corresponding points could be influential
+
+## yq_pdate
+
+This variable contains the year of sale as well as the quarter (1-4). We already have the year extracted, let's get the quarter. First are all the dates as expected?
+
+```{r}
+summary(streetrx$yq_pdate)
+```
+
+Well, we certainly have a year that is likely wrong (1969), but at least all of these values contain 5 digits
+
+```{r}
+# To find the final number (quarter), simply mod 10
+streetrx$quarter <- streetrx$yq_pdate %% 10
+```
+
+Let's take a look at the density:
+
+```{r}
+small_ppm <- streetrx %>% filter(ppm < 10)
+ggplot(small_ppm, aes(x=ppm, color=as.factor(quarter))) + geom_density(bw=0.8)
+```
+
+## Price Date
+
+Firstly, change this to a date and extract its component parts:
+
+```{r}
+streetrx$price_date <- as.Date(streetrx$price_date, format= '%m/%d/%y')
+streetrx <- streetrx %>% mutate(year = lubridate::year(price_date),
+                                month = lubridate::month(price_date),
+                                day = lubridate::day(price_date))
+```
 
@@
 There's still work to be done on this one. Ultimately about 1/3 of these values are missing - we might not be able to save it from that.
+
+## mgstr
+
+Dosage strength in mg of the units purchased. Let's check a summary:
+
+```{r}
+summary(streetrx$mgstr)
+```
+
+Not actually a continuous variable:
+
+```{r}
+table(streetrx$mgstr)
+```
+
+## bulk_purchase
+
+```{r}
+table(streetrx$bulk_purchase)
+```
+
+```{r}
+ggplot(small_ppm, aes(x=ppm, fill=bulk_purchase, color=bulk_purchase)) +
+    geom_histogram(position="identity", alpha=0.5)
+```

### 2021-03-11 eda.Rmd
Work on bulk_purchase EDA

@@
 ```{r}
-ggplot(small_ppm, aes(x=ppm, fill=bulk_purchase, color=bulk_purchase)) +
-    geom_histogram(position="identity", alpha=0.5)
+ggplot(small_ppm, aes(x=bulk_purchase, y=ppm)) + geom_boxplot()
 ```
+
+It looks like bulk purchases are generally associated with cheaper prices. But are these differences in means significant? We may apply the t-test because the CLT kicks in for these large groups
+
+```{r}
+t.test(ppm ~ bulk_purchase, data = small_ppm)
+```
+
+So bulk purchase looks to be a significant predictor *if* we exclude the extreme ppm values

### 2021-03-13 eda.Rmd
Added some results from discussion

@@
 
+Keep, log transform
+
 The big one. First, let's take a look at which fields have missing values and how frequently they occur. First, the response variable
@@
 ```{r}
-ggplot(streetrx, aes(x=ppm)) + geom_histogram(bins = 50)
+ggplot(streetrx, aes(x=log(ppm))) + geom_histogram(bins = 50)
 ```
@@
 streetrx <- streetrx %>% mutate(year = lubridate::year(price_date),
-                                month = lubridate::month(price_date),
+                                month = as.factor(lubridate::month(price_date)),
                                 day = lubridate::day(price_date))
+```
+
+I wouldn't think we should include year, as presumably we're going to be using this for future data, for which we'll have a new year.
+What about month?
+We'll use median because there are some extreme ppm values
+
+```{r}
+streetrx %>% group_by(month) %>% summarise(med_ppm = median(ppm))
+```
+
+An F test shows that we cannot reject the possibility that this variance is simply due to chance:
+
+```{r}
+aov(ppm ~ month, data = streetrx) %>% summary()
 ```
@@
 
+Recode missing as "missing"
+
 ```{r}
@@
 ## Source
+
+Missing (empty), internet (interet pharmacy, url, google), heard it, personal, other
 

### 2021-03-13 eda.Rmd
Remaining EDA changes

@@
 streetrx <- streetrx %>% filter(api_temp == 'oxymorphone')
+streetrx <- streetrx %>% filter(!is.na(ppm))
+streetrx$log_ppm <- log(streetrx$ppm)
 
@@
 
-Keep, log transform
+**Keep, log transform**
 
-The big one. First, let's take a look at which fields have missing values and how frequently they occur. First, the response variable
+Our response. First, let's take a look at which fields have missing values and how frequently they occur. First, the response variable
 
@@
 # Recode "missing" values to NA
-levels(streetrx$source)[levels(streetrx$source) == ''] <- NA_character_
-streetrx$source <- recode_factor(streetrx$source, 'w'=NA_character_,
-                                 'Idk'=NA_character_)
+levels(streetrx$source)[levels(streetrx$source) == ''] <- 'Missing'
+streetrx$source <- recode_factor(streetrx$source, 'w'='Missing',
+                                 'Idk'='Missing')
 
-levels(streetrx$source)[grepl('bluelight', levels(streetrx$source), fixed=TRUE)] <- 'bluelight'
-levels(streetrx$source)[grepl('opiophile', levels(streetrx$source), fixed=TRUE)] <- 'opiophile'
+streetrx$source <- recode_factor(streetrx$source,
+                                 'Www.drugforum.com' = 'Internet',
+                                 'Google.com' = 'Internet',
+                                 'opiophile' = 'Internet',
+                                 'silkroad' = 'Internet',
+                                 'bluelight' = 'Internet',
+                            'http://www.kentucky.com/2011/08/05/1835152/union-college-presidents-wife.html' = 'Internet')
+
+for(website in c('bluelight', 'opiophile', 'silkroad')) {
+    levels(streetrx$source)[grepl(website, levels(streetrx$source), fixed=TRUE)] <- website    
+}
 ```
@@
 
+```{r}
+mgstr_grouped <- streetrx %>% group_by(mgstr) %>% summarise(avg=mean(log_ppm))
+ggplot(mgstr_grouped, aes(x=mgstr, y=avg)) + geom_line()
+```
+
+```{r}
+ggplot(streetrx, aes(x=log_ppm)) + geom_histogram(bins = 50)
+```
+
 ## bulk_purchase
@@
 ```{r}
-ggplot(small_ppm, aes(x=bulk_purchase, y=ppm)) + geom_boxplot()
+ggplot(streetrx, aes(x=bulk_purchase, y=log_ppm)) + geom_boxplot()
 ```
@@
 ```{r}
-t.test(ppm ~ bulk_purchase, data = small_ppm)
+t.test(log_ppm ~ bulk_purchase, data = streetrx)
 ```
 
-So bulk purchase looks to be a significant predictor *if* we exclude the extreme ppm values
+So the two bulk purchase values look to have significantly different log ppm prices
+
+## Primary Reason
+
+```{r}
+table(streetrx$Primary_Reason)
+```
+
+First of all, it seems like "reporter did not answer the question" and missing should belong to the same category. Let's see:
+
+```{r}
+t.test(streetrx[streetrx$Primary_Reason == '',]$log_ppm,
+       streetrx[streetrx$Primary_Reason == '0 Reporter did not answer the question',]$log_ppm)
+```

### 2021-03-14 cjuracek_model_building.Rmd
Added model fitting with lme4 and brms

---
title: "Cole Juracek - Model Building"
output: html_notebook
---

```{r}
library(lme4)
library(lmerTest)
library(brms)
library(dplyr)

load('mod_streetrx.Rda')
streetrx$year <- as.numeric(as.character(streetrx$year))

# Need to relabel single observation
streetrx$source <- recode_factor(streetrx$source, 'STREET PRICE' = 'Heard it')

# Similar reason
levels(streetrx$Primary_Reason)[levels(streetrx$Primary_Reason) == ''] <- '0 Reporter did not answer the question'
```

# Simple models

No interactions, random intercept on state (no other random slopes), no Bayesian specification:

Predicting: log_ppm

Which variables to we want to include?

- State (as a random effect)
- USA_region
- source
- mgstr
- bulk_purchase
- Primary reason
- year

```{r}
#model_1 <- lmer(log_ppm ~ (1 | state) + (source + mgstr + bulk_purchase + year)^2, data=streetrx)
model_1 <- lmer(log_ppm ~ (1 | state) + source + mgstr + bulk_purchase + year, data=streetrx)
BIC(model_1)
```

Add new random effects:

```{r}
model_2 <- lmer(log_ppm ~ (1 | state) + USA_region + source + mgstr + bulk_purchase +
                Primary_Reason + year, data=streetrx)
#BIC(model_2)
step(model_2)
```

```{r}
model_bms <- brm(data = streetrx, family = gaussian,
    log_ppm ~ (1 | state) + source + mgstr + bulk_purchase + year,
    prior = c(prior(normal(0, 10), class = Intercept),
              prior(normal(0, 10), class = b),
              prior(cauchy(0, 1), class = sd),
              prior(cauchy(0, 1), class = sigma)),
    iter = 100, warmup = 10, chains = 2, cores = 2,
    seed = 13)
```

### 2021-03-14 cjuracek_model_building.Rmd
Small tweaks

@@
 ```{r}
-#model_1 <- lmer(log_ppm ~ (1 | state) + (source + mgstr + bulk_purchase + year)^2, data=streetrx)
 model_1 <- lmer(log_ppm ~ (1 | state) + source + mgstr + bulk_purchase + year, data=streetrx)
@@
                 Primary_Reason + year, data=streetrx)
-#BIC(model_2)
 step(model_2)

### 2021-03-15 hm-cs1-work-cleaned-3_14.Rmd
Rmd file including EDA, model selection, diagnostics...

---
title: "hm-cs1"
output: html_document
---

```{r message=FALSE}
library(tidyverse)
library(lme4)
```

```{r}
load("streetrx.RData")
```

```{r}
df <- streetrx %>% filter(api_temp=="oxymorphone")
```

# Cleaning individual variables

* Response: ppm
* Grouping variables to consider: city, state, USA_region
* Non-group predictors to consider: Continuous: yq_pdate, mgstr, ; Categorical: source, bulk_purchase, Primary_Reason
* Can drop: price_date, country (all USA), api_temp, form_temp

```{r}
# cleaning state:

# table(df$state)

# remove levels corresponding to states with 0 observations in df
df$state <- droplevels(df$state)
# table(df$state)

# drop North Dakota - one observation
df <- df %>% filter(state != "North Dakota")
df$state <- droplevels(df$state)
# table(df$state)

# replace state "USA" with "Unknown"
df$state <- recode_factor(df$state, "USA" = "Unknown")
# table(df$state)

```

```{r}
# cleaning region:

# # looks fine:
# table(df$USA_region)
# 
# # confirms that the unknown state observations are same as those with unknown region
# table(df$USA_region, df$state)
```

Change date to numeric (easier for interpretation?):

```{r}
# hist(as.numeric(df$price_date))

# this sets 
df$date_num <- as.numeric(as.Date(df$price_date, "%m/%d/%y")) - 10957 # sets 1/1/2000 as zero

tempdf <- df %>% select(date_num, price_date)

# this filters out all data before the 1/1/2000, which probably shouldn't be included
df <- df %>% filter(date_num >= 0)

hist(df$date_num)
```

```{r}
hist(df$mgstr)
# fine as is?
```

```{r}
# recode sources

# remove unused levels
df$source <- droplevels(df$source)

source_df <- data.frame(table(df$source))

# levels(df$source)
levels(df$source)[1] <- "Unknown"

source_df <- data.frame(table(df$source))

df$source <- recode_factor(df$source, "Unknown" = "Unknown",
                           "Internet" = "Internet",
                           "Internet Pharmacy" = "Internet",
                           "Personal" = "Personal",
                           "Heard it" = "Heard it",
                           "Idk" = "Unknown",
                           "w" = "Unknown",
                           "STREET PRICE" = "Heard it",
                           .default = "Internet") # everything else is a URL

source_df_2 <- data.frame(table(df$source))

print(source_df)
print(source_df_2)
```

```{r}
table(df$bulk_purchase, useNA = "always")
# looks fine
```

```{r}

# remove unused levels
df$Primary_Reason <- droplevels(df$Primary_Reason)

reason_df <- data.frame(table(df$Primary_Reason))

# levels(df$Primary_Reason)
levels(df$Primary_Reason)[1] <- "Other or unknown"

reason_df <- data.frame(table(df$Primary_Reason))

df$Primary_Reason <- recode_factor(df$Primary_Reason, "Other or unknown" = "Other or unknown",
                                   "9 To self-treat my pain" = "Self-treat",
                                   "3 To prevent or treat withdrawal" = "Self-treat",
                                   "10 To treat a medical condition other than pain" = "Self-treat",
                                   "4 For enjoyment/to get high" = "Enjoyment",
                           .default = "Other or unknown")

reason_df_2 <- data.frame(table(df$Primary_Reason))

print(reason_df)
print(reason_df_2)

```

```{r}
# log transform for ppm looks good
hist(log(df$ppm))

# transform
df$log_ppm <- log(df$ppm)
```

## More EDA

Thinking about what to group by...
State best? Regions kind of similar, cities have many small sample sizes...

```{r}
ggplot(df, aes(x=reorder(state, state, length), y=log_ppm, fill=state)) +
geom_boxplot() +
labs(title="log PPM by state", x="state",y="Log PPM")

ggplot(df, aes(x=reorder(USA_region, USA_region, length), y=log_ppm, fill=USA_region)) +
geom_boxplot() +
labs(title="log PPM by region", x="region",y="Log PPM")

ggplot(df[df$state=="Florida",], aes(x=reorder(city, city, length), y=log_ppm)) +
geom_boxplot() +
labs(title="log PPM by city (FL)", x="city",y="Log PPM")

ggplot(df[df$state=="California",], aes(x=reorder(city, city, length), y=log_ppm)) +
geom_boxplot() +
labs(title="log PPM by city (CA)", x="city",y="Log PPM")
```

Closer look at by-state values of price

```{r}
ggplot(df[df$USA_region == "Northeast" ,], aes(x=reorder(state, state, length), y=log_ppm, fill=state)) +
geom_boxplot() +
labs(title="log PPM by state - Northeast", x="state",y="Log PPM")

ggplot(df[df$USA_region == "West" ,], aes(x=reorder(state, state, length), y=log_ppm, fill=state)) +
geom_boxplot() +
labs(title="log PPM by state - West", x="state",y="Log PPM")

ggplot(df[df$USA_region == "South" ,], aes(x=reorder(state, state, length), y=log_ppm, fill=state)) +
geom_boxplot() +
labs(title="log PPM by state - South", x="state",y="Log PPM")

ggplot(df[df$USA_region == "Midwest" ,], aes(x=reorder(state, state, length), y=log_ppm, fill=state)) +
geom_boxplot() +
labs(title="log PPM by state - Midwest", x="state",y="Log PPM")

ggplot(df[df$USA_region == "Other/Unknown" ,], aes(x=reorder(state, state, length), y=log_ppm, fill=state)) +
geom_boxplot() +
labs(title="log PPM by state - Other/Unknown", x="state",y="Log PPM")
```

Looking more at cities...

```{r}
# some states have few observations for even most populous cities... 
temp <- df$city[df$state=="Massachusetts"]
length(temp)

# look at the big cities overall 
city_temp <- df %>% group_by(city) %>% summarize(city_count = n()) %>% arrange(-city_count)
city_temp[1:20, ]
hist(city_temp$city_count[-1])
View(city_temp)

# look at the big states  
state_temp <- df %>% group_by(state) %>% summarize(state_count = n()) %>% arrange(-state_count)
state_temp[1:20, ]
hist(state_temp$state_count[-1])
View(state_temp)
```

Use random slopes (by region)?
Looks like no for continuous variables, maybe for categorical ones.

```{r}
ggplot(data = df, aes(x = date_num, y = log_ppm)) + 
  geom_point() +
  facet_grid(~ USA_region)

ggplot(data = df, aes(x = mgstr, y = log_ppm)) + 
  geom_point() +
  facet_grid(~ USA_region)

ggplot(data = df, aes(x = source, y = log_ppm)) + 
  geom_boxplot() +
  facet_grid(~ USA_region)

ggplot(data = df, aes(x = bulk_purchase, y = log_ppm)) + 
  geom_boxplot() +
  facet_grid(~ USA_region)

ggplot(data = df, aes(x = Primary_Reason, y = log_ppm)) + 
  geom_boxplot() +
  facet_grid(~ USA_region)
```

Random slopes by state? Sampling a few states at a time for visibility.

For continuous variables looks like no, maybe for categorical variables.

```{r}
ggplot(data = df[df$state=="Florida" | df$state=="Texas" | df$state=="North Carolina",], aes(x = date_num, y = log_ppm)) + 
  geom_point() +
  facet_grid(~ state)

ggplot(data = df[df$state=="Massachusetts" | df$state=="Connecticut" | df$state=="New York",], aes(x = date_num, y = log_ppm)) + 
  geom_point() +
  facet_grid(~ state)

ggplot(data = df[df$state=="California" | df$state=="Oregon" | df$state=="Arizona",], aes(x = date_num, y = log_ppm)) + 
  geom_point() +
  facet_grid(~ state)

ggplot(data = df[df$state=="Michigan" | df$state=="Wisconsin" | df$state=="Kansas",], aes(x = date_num, y = log_ppm)) + 
  geom_point() +
  facet_grid(~ state)
```

```{r}
# random slopes for mgstr?

ggplot(data = df[df$state=="Florida" | df$state=="Texas" | df$state=="North Carolina",], aes(x = mgstr, y = log_ppm)) + 
  geom_point() +
  facet_grid(~ state)

ggplot(data = df[df$state=="Massachusetts" | df$state=="Connecticut" | df$state=="New York",], aes(x = mgstr, y = log_ppm)) + 
  geom_point() +
  facet_grid(~ state)

ggplot(data = df[df$state=="California" | df$state=="Oregon" | df$state=="Arizona",], aes(x = mgstr, y = log_ppm)) + 
  geom_point() +
  facet_grid(~ state)

ggplot(data = df[df$state=="Michigan" | df$state=="Wisconsin" | df$state=="Kansas",], aes(x = mgstr, y = log_ppm)) + 
  geom_point() +
  facet_grid(~ state)
```

```{r}
# categorical variables -- looking at South only
ggplot(data = df[df$USA_region == "South",], aes(x = bulk_purchase, y = log_ppm)) + 
  geom_boxplot() +
  facet_grid(~ state)

ggplot(data = df[df$USA_region == "South",], aes(x = Primary_Reason, y = log_ppm)) + 
  geom_boxplot() +
  facet_grid(~ state)

ggplot(data = df[df$USA_region == "South",], aes(x = source, y = log_ppm)) + 
  geom_boxplot() +
  facet_grid(~ state)
```

Modeling:

Include random intercepts for state, but should region also be included as fixed effect? -- No.
Are grouping by region or by city better than by state? -- No.

```{r}
mod1 <- lmer(data=df, log_ppm ~ (1 | state)  + date_num + mgstr + bulk_purchase + Primary_Reason + source)
mod2 <- lmer(data=df, log_ppm ~ (1 | state) + USA_region + date_num + mgstr + bulk_purchase + Primary_Reason + source)
anova(mod1, mod2)
anova(mod1_2, mod1)
# confirsms we may be able to leave region out
# all other predictors in mod1 significant

mod_s <- lmer(data=df, log_ppm ~ (1 | state)  + date_num + mgstr + bulk_purchase + Primary_Reason + source)
mod_r <- lmer(data=df, log_ppm ~ (1 | USA_region)  + date_num + mgstr + bulk_purchase + Primary_Reason + source)
mod_c <- lmer(data=df, log_ppm ~ (1 | city)  + date_num + mgstr + bulk_purchase + Primary_Reason + source)

BIC(mod_s) # state grouping is lowest - confirms our suspicion that state was the best grouping variable
BIC(mod_r)
BIC(mod_c)
```

Try random slopes for each categorical variable. 
Do they give better model? -- No.

```{r}
# none of the random slopes for the categorical variables seem to add to the model 
# all predictors significant in mod1 so keep all 
# mod1 is final model

mod1 <- lmer(data=df, log_ppm ~ (1 | state) + date_num + mgstr + bulk_purchase + Primary_Reason + source)
mod2_bulk <- lmer(data=df, log_ppm ~ (bulk_purchase| state) + date_num + mgstr + bulk_purchase + Primary_Reason + source)
mod2_reason <- lmer(data=df, log_ppm ~ (Primary_Reason| state) + date_num + mgstr + bulk_purchase + Primary_Reason + source)
mod2_source <- lmer(data=df, log_ppm ~ (source| state) + date_num + mgstr + bulk_purchase + Primary_Reason + source)

anova(mod1, mod2_bulk)
anova(mod1, mod2_reason)
anova(mod1, mod2_source)
```

Diagnostics of the (potential) final model: 
Good sign -- No states are outliers.

```{r}
library(influence.ME)
mod1_inf <- influence(mod1,"state")
print(2/sqrt(length(unique(df$state))))
# dfbetas(mod1_inf)
cooks.distance(mod1_inf,sort=TRUE)

# no states have values above the cutoff
```

### 2021-03-18 STA 610 Case Study 1 Report_Team 4.Rmd

---
title: "STA610 Case Study 1 Team 4 Report"
author: 
  - Cole Juracek (Coordinator) 
  - Lauren Palazzo (Programmer) 
  - Lingyu Zhou (Checker) 
  - Fan Zhu (Presenter)
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
library(lme4)
library(gridExtra)
library(grid)
library(ggplot2)
library(lattice)
#devtools::install_github("goodekat/redres")
library(redres)
```

```{r, echo = FALSE}
load("streetrx.RData")
df <- streetrx %>% filter(api_temp=="oxymorphone", !is.na(ppm))
```

```{r, echo = FALSE}
# cleaning state:

# table(df$state)

# remove levels corresponding to states with 0 observations in df
df$state <- droplevels(df$state)
# table(df$state)

# drop North Dakota - one observation
df <- df %>% filter(state != "North Dakota")
df$state <- droplevels(df$state)
# table(df$state)

# replace state "USA" with "Unknown"
df$state <- recode_factor(df$state, "USA" = "Unknown")
# table(df$state)

# log transform ppm
df$log_ppm <- log(df$ppm)
```

## EDA
#### PPM

```{r, fig.height=3, fig.align='center'}
hist_ppm <- ggplot(df, aes(x=ppm)) + 
 geom_histogram(aes(y=..density..), colour="black", fill="white", bins = 50)+
 geom_density(alpha=.2, fill="red")

hist_log_ppm <- ggplot(df, aes(x=log_ppm)) + 
 geom_histogram(aes(y=..density..), colour="black", fill="white", bins = 50)+
 geom_density(alpha=.2, fill="blue")

grid.arrange(hist_ppm, hist_log_ppm, ncol=2)
```
The histogram of the original ppm is extremely right skewed, while the distribution of the log(ppm) looks roughly normal. Thus, we will use the log(ppm) as the dependent variable in our model.

```{r, echo = FALSE}
# Change date to numeric (easier for interpretation)

# this sets 
df$date_num <- as.numeric(as.Date(df$price_date, "%m/%d/%y")) - 10957 # sets 1/1/2000 as zero

tempdf <- df %>% select(date_num, price_date)

# this filters out all data before the 1/1/2000, which probably shouldn't be included
df <- df %>% filter(date_num >= 0)

```

#### Sources

We recoded the categorical variable "sources" into the following 4 levels.
```{r, echo = FALSE}
# recode sources

# remove unused levels
df$source <- droplevels(df$source)

source_df <- data.frame(table(df$source))

# levels(df$source)
levels(df$source)[1] <- "Unknown"

source_df <- data.frame(table(df$source))

df$source <- recode_factor(df$source, "Unknown" = "Unknown",
                           "Internet" = "Internet",
                           "Internet Pharmacy" = "Internet",
                           "Personal" = "Personal",
                           "Heard it" = "Heard it",
                           "Idk" = "Unknown",
                           "w" = "Unknown",
                           "STREET PRICE" = "Heard it",
                           .default = "Internet") # everything else is a URL

source_df_2 <- data.frame(table(df$source))

knitr::kable(source_df_2, col.names = c("Level", "Freq"))
```

```{r}
#table(df$bulk_purchase, useNA = "always")
# looks fine
```

#### Primary_Reason

We recoded the categorical variable "Primary_Reason" into the following 3 levels.
```{r}
# remove unused levels
df$Primary_Reason <- droplevels(df$Primary_Reason)

reason_df <- data.frame(table(df$Primary_Reason))

# levels(df$Primary_Reason)
levels(df$Primary_Reason)[1] <- "Other or unknown"

reason_df <- data.frame(table(df$Primary_Reason))

df$Primary_Reason <- recode_factor(df$Primary_Reason, "Other or unknown" = "Other or unknown",
                                   "9 To self-treat my pain" = "Self-treat",
                                   "3 To prevent or treat withdrawal" = "Self-treat",
                                   "10 To treat a medical condition other than pain" = "Self-treat",
                                   "4 For enjoyment/to get high" = "Enjoyment",
                           .default = "Other or unknown")

reason_df_2 <- data.frame(table(df$Primary_Reason))

#print(reason_df)
knitr::kable(reason_df_2, col.names = c("Level", "Freq"))
```

#### State
```{r, fig.align='center', fig.height=6}
ggplot(df, aes(x=log_ppm, y=reorder(state, state, length), fill=state)) +
geom_boxplot() +
labs(title="state", x="log PPM by state",y="Log PPM")
```
The boxplot of log PPM by state is arranged in the order of increasing sample size from the bottom to the top.

#### Region

```{r, fig.align='center'}
ggplot(df, aes(x=reorder(USA_region, USA_region, length), y=log_ppm, fill=USA_region)) +
geom_boxplot() +
labs(title="log PPM by region", x="region",y="Log PPM")
```

#### City
```{r, fig.align='center'}
fl_lppmbycity <- ggplot(df[df$state=="Florida",], aes(x=log_ppm, y=reorder(city, city, length))) +
geom_boxplot() +
labs(title="log PPM by city (FL)", x="Log PPM",y="city")

ca_lppmbycity <- ggplot(df[df$state=="California",], aes(x=reorder(city, city, length), y=log_ppm)) +
geom_boxplot() +
labs(title="log PPM by city (CA)", x="Log PPM",y="city")

grid.arrange(fl_lppmbycity, ca_lppmbycity, ncol= 2)
```

#### mgstr
```{r, fig.align='center'}
# random slopes for mgstr?

ggplot(data = df[df$state=="Florida" | df$state=="Texas" | df$state=="North Carolina",], aes(x = mgstr, y = log_ppm)) + 
  geom_point() +
  facet_grid(~ state)
```

#### bulk_purchase
```{r, fig.align='center'}
# categorical variables -- looking at South only
ggplot(data = df[df$USA_region == "South",], aes(x = bulk_purchase, y = log_ppm)) + 
  geom_boxplot() +
  facet_grid(~ state)
```

#### date_num
```{r, fig.align='center'}
# categorical variables -- looking at South only
ggplot(data = df[df$state=="Florida" | df$state=="Texas" | df$state=="North Carolina",], aes(x = date_num, y = log_ppm)) + 
  geom_point() +
  facet_grid(~ state)
```

## Model
#### Model Specification

$$y_{ij} = \beta_{0,j} + \beta_{1} date_num + \beta_{2} mgstr + \beta_{3} bulk_purchase + \beta_{4} Primary_Reason + \beta_{5} source + \epsilon_{ij}$$
$$
\beta_{0,j} = \beta_0 + b_{0,j}, b_{0,j} \sim N(0,\tau^2)
$$
$$
\epsilon_{ij} \perp b_{0,j} \overset{iid}{\sim} N(0,\sigma^2)
$$

* i stands for individuals and j stands for each state. 
* $\beta_{0,j}$ is the random intercept by state. 
* $\beta_{1}$ measures the fixed effect for date_num, the date variable.
* $\beta_{2}$ measures the fixed effect for mgstr, the dosage strength of reported drug in milligrams.
* $\beta_{3}$ measures the fixed effect for bulk_purchase is a binary variable that is represented by an indicator function I(bulk_purchase = Yes). 
* $\beta_{4}$ measures the fixed effect for the categorical variable Primary_Reason which contains three categories: Other or unknown, Self-treat and Enjoyment.
* $\beta_{5}$ measures the fixed effect for the categorical variable source

Model Selection discussions here.

```{r}
mod1 <- lmer(data=df, log_ppm ~ (1 | state) + date_num + mgstr + bulk_purchase + Primary_Reason + source)
```

#### Model Diagnostics
```{r, fig.align='center', warning=FALSE, message=FALSE, error=FALSE}
resqq <- plot_resqq(mod1)
ranefci <- plot_ranef(mod1)
grid.arrange(resqq, ranefci, ncol = 2)
```

Model diagnostics discussions here.

## Results
#### Fixed Effects
```{r, message=FALSE}
#knitr::kable(fixef(mod1))
#confint(mod1)[3:11,]

# extract coefficients
coefs <- data.frame(coef(summary(mod1)))
# use normal distribution to approximate p-value
coefs$p.z <- 2 * (1 - pnorm(abs(coefs$t.value)))
knitr::kable(coefs, col.names = c("Estimate", "Std.Error", "t-value", "P-value"))
```

* Bulk_purchase: “Not bulk purchase” is the reference group.
Controlling for other variables, being bulk purchase increases the price per mg by exp(-0.16) = 0.8521. 
* Primary_Reason: “Primary_Reason: Other or Unknown” is the reference group. 
Controlling for other variables, taking oxymorphone with the primary reason of self-treating tends to receive a exp(0.05) = 1.0513 times higher price per mg than people whose primary reason is “other or unknown”.
Controlling for other variables, taking oxymorphone with the primary reason of enjoyment tends to receive a exp(0.02) = 1.0202 times higher price per mg than people whose primary reason is “other or unknown”.
* Source: “Source: Unknown” is the reference group. 
Controlling for other variables, purchasing oxymorphone with Internet source tends to receive a exp(0.02) = 1.0202 times higher price per mg than people with unknown drug source.
Controlling for other variables, purchasing oxymorphone with personal source tends to receive a exp(0.07) = 1.0725 times higher price per mg than people with unknown drug source.
Controlling for other variables, purchasing oxymorphone with “heard it” source tends to receive a exp(0.24) = 1.2712 times higher price per mg than people with unknown drug source.

#### Random Effects
```{r}
#VarCorr(mod1)
Variance <- c(0.04748, 0.96369)
Std.Dev <- c(0.2179, 0.9817)
Groups <- c("state (Intercept)", "Residual")

REvar <- cbind(Groups, Variance, Std.Dev)

knitr::kable(REvar, col.names = c("Group", "Variance", "Std.Dev"))

```
The within-state variance $\sigma^2$ is 0.96369, which is much larger than the across-state variance $\tau^2$, 0.04748.

```{r, fig.align='center', fig.height=8}
dotplot(ranef(mod1, condVar=TRUE))$state
```
From the result of the random effects, we can see that the random intercept for Tennessee is the highest and it is significantly different from zero. The random intercept for Nevada is the lowest and this intercept is also significant.

## Appendix
```{r, ref.label=knitr::all_labels(),echo=TRUE,eval=FALSE}
```

### 2021-03-18 STA 610 Case Study 1 Report_Team 4.Rmd
added interpretations for fixed effects of mgstr and date_num

@@
 Controlling for other variables, purchasing oxymorphone with “heard it” source tends to receive a exp(0.24) = 1.2712 times higher price per mg than people with unknown drug source.
+* For the date_num variable, all else quals, for every one unit increase in the date_num, the price per mg of oxymorphone decreases to exp(-0.0001290) =  0.999871 times the original price.
+* For the mgstr variable, controlling for for other variables, for every one unit increase in mgstr, the price per mg of oxymorphone decrease to exp(-0.0241262) =  0.9761625 times the original price.
 

### 2021-03-20 STA 610 Case Study 1 Report_Team 4.Rmd
Added pacman to handle dependencies

--- 2021-03-18 STA 610 Case Study 1 Report_Team 4.Rmd
@@
 ```{r, echo=FALSE, warning=FALSE, message=FALSE}
-library(tidyverse)
-library(lme4)
-library(gridExtra)
-library(grid)
-library(ggplot2)
-library(lattice)
+if (!require(pacman)) install.packages("pacman")
+pacman::p_load(tidyverse, lme4, gridExtra, grid, ggplot2, lattice, redres)
 #devtools::install_github("goodekat/redres")
-library(redres)
 ```

### 2021-03-20 STA 610 Case Study 1 Report_Team 4.Rmd
Started writeup on model specification

@@
 ## Model
+
+Per the research question of interest, we are interested in exploring factors related to price per mg of oxymorphone.
+We are also specifically asked to control for location and how drug price changes with it.
+Cities seemed to be too noisy to use as a location predictor.
+On the flip side, region did not seem to be associated with a significant change in drug prices.
+Therefore, state provides the best granularity of accounting for location in this instance.
+
+Linear regression models are usually a good standard approach to this, providing clear interpretations of how various predictors are associated with drug price.
+As seen in the EDA, however, many states have excessively small sample sizes.
+This produces corresponding $\hat{\beta}_{OLS}$ estimates with large variance.
+We would like to be able to *partially pool* information across states, letting us borrow information to help stabilize state estimates with small sample sizes.
+
+This prompts a **mixed effects model**, where we instead model the state coefficients as having been drawn $N(0, \tau^2)$.
+
 #### Model Specification
@@
 ```{r, fig.align='center', warning=FALSE, message=FALSE, error=FALSE}
-resqq <- plot_resqq(mod1)
-ranefci <- plot_ranef(mod1)
-grid.arrange(resqq, ranefci, ncol = 2)
+# resqq <- plot_resqq(mod1)
+# ranefci <- plot_ranef(mod1)
+# grid.arrange(resqq, ranefci, ncol = 2)
 ```

### 2021-03-20 STA 610 Case Study 1 Report_Team 4.Rmd
Small formatting changes

@@
 * i stands for individuals and j stands for each state. 
-* $\beta_{0,j}$ is the random intercept by state. 
-* $\beta_{1}$ measures the fixed effect for date_num, the date variable.
-* $\beta_{2}$ measures the fixed effect for mgstr, the dosage strength of reported drug in milligrams.
-* $\beta_{3}$ measures the fixed effect for bulk_purchase is a binary variable that is represented by an indicator function I(bulk_purchase = Yes). 
-* $\beta_{4}$ measures the fixed effect for the categorical variable Primary_Reason which contains three categories: Other or unknown, Self-treat and Enjoyment.
+* $\beta_{0,j}$: Random intercept by state
+* $\beta_{1}$: Fixed effect for date_num, the date variable
+* $\beta_{2}$: Fixed effect for mgstr, the dosage strength of reported drug in milligrams.
+* $\beta_{3}$: Fixed effect for bulk_purchase is a binary variable that is represented by an indicator function I(bulk_purchase = Yes). 
+* $\beta_{4}$: Fixed effect for the categorical variable Primary_Reason which contains three categories:     - Other or unknown
+    - Self-treat
+    - Enjoymen
 * $\beta_{5}$ measures the fixed effect for the categorical variable source

### 2021-03-20 report_draft_fz_lp_4.Rmd
working draft of report - 3/20
tweaked EDA, starting writing the report, explored model selection and diagnostics more...

---
title: "STA610 Case Study 1 Team 4 Report"
author:
- Cole Juracek (Coordinator)
- Lauren Palazzo (Programmer)
- Lingyu Zhou (Checker)
- Fan Zhu (Presenter)
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
library(lme4)
library(gridExtra)
library(grid)
library(ggplot2)
library(lattice)
# devtools::install_github("goodekat/redres")
library(redres)
library(stringr)
library(influence.ME)
```

```{r, echo = FALSE}

load("~/Documents/Spring 2021/STA 610 - H Models/cs1/streetrx.RData")
df <- streetrx %>% filter(api_temp=="oxymorphone", !is.na(ppm))
```

# Introduction

This study is an exploration of the relationship between the sale price per milligram of diverted pharmaceutical substances and other factors related to the sale, as reflected in self-reported data from the website StreetRx. StreetRx maintains a partnership with the Researched Abuse, Diversion, and Addiction-Related Surveillance System (RADARS), which aims for these data to inform public health measures such as health policy related to illicit drugs.

[insert something about our specific drug here]

Among the data available to us are location data pertaining to the city, state, and United States region of the sale. These variables present a natural grouping structure for the price data, which suggest that we employ a hierarchical model in our analysis. The benefits of a hierarchical model include the ability to "borrow information" from the data set as a whole to stabilize estimates of price for groups (e.g., states) with relatively small sample size, and to establish a estimated distribution of how price varies with respect to groups, which would in theory allow generalization to new groups not included in the original analysis.

# Exploratory data analysis and data cleaning

## Data cleaning

### Group variables

We performed an initial examination of two grouping variables, `state` and `USA_region`. The state North Dakota had only one observation, and as such was not considered a well-defined "group" for which a variance could be estimated; thus this state's observation was removed. States labeled simply as "USA" were recoded to "Unknown". Region was also sometimes labeled as uknown or other; these observations coincided with the unknown states (see appendix Table X). (The variable `city` as a grouping variable is addressed later in the report.)

```{r, echo = FALSE}
# cleaning state:

# table(df$state)

# remove levels corresponding to states with 0 observations in df
df$state <- droplevels(df$state)
# table(df$state)

# drop North Dakota - one observation
df <- df %>% filter(state != "North Dakota")
df$state <- droplevels(df$state)
# table(df$state)

# replace state "USA" with "Unknown"
df$state <- recode_factor(df$state, "USA" = "Unknown")
# table(df$state)

### PUT IN APPENDIX: the state-by-region table of coutns below
# table(df$state, df$USA_region)
```

### Date variable

We considered two different ways to engineer the date-of-sale feature. One intuitive and commonly-used approach would be to recode this as a continuous variables counting days from some start date, the intent of which would be to capture the influence of inflation for this price data that spans years. Another approach would be to consider whether there might be an effect of seasonality on drug prices.

For recoding date into a continuous variable, `date_num`, we filter out observations dated prior to the year 2010, as that is the year that the StreetRx web site was founded. For the seasonality variable, `date_quarter`, we extract the "quarter" of the year (numbered 1 through 4) as its own variable. Below we see the distributions of these variables:

```{r, echo = FALSE}
# Change date to numeric (easier for interpretation)

# this sets 
df$date_num <- as.numeric(as.Date(df$price_date, "%m/%d/%y")) - 14621 # sets 1/12/2010 (first post-2010 date) as zero

tempdf <- df %>% select(date_num, price_date)

# this filters out all data before the 1/1/2000, which probably shouldn't be included
df <- df %>% filter(date_num >= 0)
```

```{r}
df$date_quarter <- as.numeric(substr(df$yq_pdate, 5, 5))
```

```{r}
hist(df$date_num)
knitr::kable(table(df$date_quarter))
```

### Sources variable

As the variable `source` that describes the source of the price information is a user-input field, it is an inherently "messy" variable that calls for some recoding in order to make the levels of the variable more meaningful and interpretable, and to avoid overfitting. In the case of `source`, some user input is simply blank, while a small number of other entries indicate in some other way that the source is unknown or unclear. Since it is quite uncommon in the data for a source to be entered yet indicated as unknown, we decided to drop these unkown-source observations as being indicative of poor data quality. (If the user cannot identify their source of price information, why should we trust their price data?) We decided to distinguish between "Internet pharmacy" sources and other internet sources, since there may be a distinction between a source that is selling a drug versus perhaps just discussing a drug price in a forum, which may yield less legitimate information. Ultimately, we recoded the categorical variable "sources" into the following levels.

```{r, echo = FALSE}
# recode sources

# remove unused levels
df$source <- droplevels(df$source)

source_df <- data.frame(table(df$source))

# levels(df$source)
levels(df$source)[1] <- "Blank"

source_df <- data.frame(table(df$source))

df$source <- recode_factor(df$source, "Blank" = "Blank",
                           "Internet" = "Internet",
                           "Internet Pharmacy" = "Internet Pharmacy",
                           "Personal" = "Personal",
                           "Heard it" = "Heard it",
                           "Idk" = "Unknown",
                           "w" = "Unknown",
                           "STREET PRICE" = "Heard it",
                           .default = "Internet") # everything else is a URL

source_df_2 <- data.frame(table(df$source))

# drop unknowns

df <- df %>% filter(df$source != "Unknown")
df$source <- droplevels(df$source)

knitr::kable(table(df$source), col.names = c("Level", "Freq"))
```

### Bulk purchase variable

Our visualization of the bulk purchase variable indicates nothing unexpected that calls for data cleaning.

```{r}
#table(df$bulk_purchase, useNA = "always")
# looks fine
```

### Purchase reason variable

`Primary_Reason` for purchase is another user-input variable that calls for recoding. Similarly to the `source` variable, we retain the information about which entries were blank. We also try to distinguish this from a different response level that is not blank, yet indicates the user did not provide a reason. We further still distinguished between these responses and a non-blank response of refusal to answer. We strove to avoid making presumptions of no difference between these similar-seeming response types, as they may indicate variations in data processing that hold some predictive power that we are not yet aware of. We ultimately recoded the categorical variable `Primary_Reason` into the following levels.

```{r}
# remove unused levels
df$Primary_Reason <- droplevels(df$Primary_Reason)

reason_df <- data.frame(table(df$Primary_Reason))

# levels(df$Primary_Reason)
levels(df$Primary_Reason)[1] <- "Blank"

reason_df <- data.frame(table(df$Primary_Reason))

df$Primary_Reason <- recode_factor(df$Primary_Reason, "Blank" = "Blank",
                                   "0 Reporter did not answer the question" = "Left unanswered",
                                   "Other or unknown" = "Other or unknown",
                                   "9 To self-treat my pain" = "Self-treat",
                                   "3 To prevent or treat withdrawal" = "Self-treat",
                                   "10 To treat a medical condition other than pain" = "Self-treat",
                                   "4 For enjoyment/to get high" = "Enjoyment",
                                   "5 To resell" = "Resell",
                                   "8 Prefer not to answer" = "Refuse to answer",
                           .default = "Other or unknown")

reason_df_2 <- data.frame(table(df$Primary_Reason))

#print(reason_df)
knitr::kable(reason_df_2, col.names = c("Level", "Freq"))
```

### Drug strength variable

We renamed this variable `dose_per_mg` to make its meaning clearer to the reader. This is a continuous variable that nevertheless tends to take on only a few values, as depicted in the histogram. We considered a recoding of this variable into levels, but decided against it as the dose per milligram technically may take on any value and the generalizability of our model may be improved if we are able to consider dose levels that are in a realistic range yet not present in our current data.

```{r}
hist(df$mgstr)
# rename variable for ease of interpretability
df <- df %>% rename(dose_per_mg = mgstr)
```

### Response variable: Price per milligram

We are considering fitting a hierarchical model with a response variable that has normally distributed errors, and so we would like our response variable, here `ppm`, or price per milligram, to be close to normally-distributed. The histogram of the `ppm` is extremely right skewed (not unexpected for data related to financial measures such as price, salary, earnings, etc.), while the distribution of the log(`ppm`) looks roughly normal. Thus, we will use `log_ppm` as the dependent variable in our model.

```{r, fig.height=3, fig.align='center'}
# log transform ppm
df$log_ppm <- log(df$ppm)

hist_ppm <- ggplot(df, aes(x=ppm)) + 
 geom_histogram(aes(y=..density..), colour="black", fill="white", bins = 50)+
 geom_density(alpha=.2, fill="red")

hist_log_ppm <- ggplot(df, aes(x=log_ppm)) + 
 geom_histogram(aes(y=..density..), colour="black", fill="white", bins = 50)+
 geom_density(alpha=.2, fill="blue")

grid.arrange(hist_ppm, hist_log_ppm, ncol=2)
```

In order to cut off extreme price values that may skew the results for typical prices, we restrict the `log_ppm` distribution to more realistic and representative values:

```{r}
df <- df %>% filter(log_ppm > -2 & log_ppm < 2)
```

## Exploratory data analysis for random intercepts

We now take a closer look at our grouping variables with a focus on the question of whether random intercepts might be appropriate.

### State

We show boxplots of `log_ppm` by `state`, arranged in the order of increasing sample size from the bottom to the top.

```{r, fig.align='center', fig.height=6}
ggplot(df, aes(x=log_ppm, y=reorder(state, state, length), fill=state)) +
geom_boxplot() +
labs(title="State", x="Log PPM by state",y="Log PPM")
```

We see that some states with small sample sizes have distributions of `log_price` that seem unstable, with respect to how much they differ from other boxplots; this provides support for the idea that the stabilizing effect of a hierarchical model would benefit our analysis. The fact that some distributions differ from each other in a noteworthy way (e.g., boxplots hardly overlapping) suggests group-based differences in price that should be taken into account in modeling, as opposed to lumping all states' data together without distinction.

### Region

Viewing a similar plot for the regions of the USA, we do see some variability in price among regions, but it is not as pronounced as for the states.

```{r, fig.align='center'}
ggplot(df, aes(x=reorder(USA_region, USA_region, length), y=log_ppm, fill=USA_region)) +
geom_boxplot() +
labs(title="log PPM by region", x="region",y="Log PPM")
```

### City

We find that there are many cities with a sample size of one, which makes `city` perhaps not the most natural choice for a grouping variable. Example visualizations shown here are for all cities in the states of Florida and California.

```{r, fig.align='center'}
fl_lppmbycity <- ggplot(df[df$state=="Florida",], aes(x=log_ppm, y=reorder(city, city, length))) +
geom_boxplot() +
labs(title="log PPM by city (FL)", x="Log PPM",y="city")

ca_lppmbycity <- ggplot(df[df$state=="California",], aes(x=reorder(city, city, length), y=log_ppm)) +
geom_boxplot() +
labs(title="log PPM by city (CA)", x="Log PPM",y="city")

grid.arrange(fl_lppmbycity, ca_lppmbycity, ncol= 2)
```

We decided to focus on `state` as our primary candidate for a grouping variable moving forward.

## Exploratory data analysis for random slopes

To consider whether we may want to include random slopes by state for our predictor variables, we examined plots of the predictor versus response among each state individually to determine whether it seemed that slopes fit to the group-level data would differ. (We present a sample of visualizations here rather than visualizations for every state, due to space constraints.)

For out continuous variables (continuous date and dosage strength), in by-state plots we often saw strikingly similar patterns in predictor vs. response, suggesting that random slopes may not be necessary for these variables.

### dose_per_mg
```{r, fig.align='center'}
ggplot(data = df[df$state=="Florida" | df$state=="Texas" | df$state=="North Carolina",], aes(x = dose_per_mg, y = log_ppm)) + 
  geom_point() +
  facet_grid(~ state)
```

### date_num
```{r, fig.align='center'}
ggplot(data = df[df$state=="Florida" | df$state=="Texas" | df$state=="North Carolina",], aes(x = date_num, y = log_ppm)) + 
  geom_point() +
  facet_grid(~ state)
```

For categorical variables, we generally observed was there was indeed variation among the states, but it was not clear whether this variation had any well-defined structure. It was not clear whether our models would benefit from random slopes for these variables, so we decided to further investigate this in our modeling process.

### bulk_purchase
```{r, fig.align='center'}
# categorical variables -- looking at South only
ggplot(data = df[df$USA_region == "South",], aes(x = bulk_purchase, y = log_ppm)) + 
  geom_boxplot() +
  facet_grid(~ state)
```

### date_quarter
```{r}
ggplot(data = df[df$state=="Florida" | df$state=="Texas" | df$state=="North Carolina",], aes(x = date_quarter, y = log_ppm, group = date_quarter)) + 
  geom_boxplot() +
  facet_grid(~ state)
```

### Primary_Reason
```{r}
ggplot(data = df[df$state=="Florida" | df$state=="Texas" | df$state=="North Carolina",], aes(x = Primary_Reason, y = log_ppm, group = Primary_Reason)) + 
  geom_boxplot() +
  facet_grid(~ state)
```

### source
```{r}
ggplot(data = df[df$state=="Florida" | df$state=="Texas" | df$state=="North Carolina",], aes(x = source, y = log_ppm, group = source)) + 
  geom_boxplot() +
  facet_grid(~ state)
```
## More EDA: slope and intercept distributions for predictors by state

To get a better sense of how state-level groups and slopes are distributed, we created the following intercepts for our coninuous variables:

### Dose strength variable

```{r}
group_int_list <- NULL 
group_slope_list <- NULL

for(statej in
  unique(df$state))
{
m <- lm(data = df[df$state==statej, ], log_ppm ~ dose_per_mg)

group_int_list <- c(group_int_list, coef(m)[1])
group_slope_list <- c(group_slope_list, coef(m)[2])
}

hist(group_int_list)
hist(group_slope_list)
```

### Continuous date variable

```{r}
group_int_list <- NULL 
group_slope_list <- NULL

for(statej in
  unique(df$state))
{
m <- lm(data = df[df$state==statej, ], log_ppm ~ date_num)

group_int_list <- c(group_int_list, coef(m)[1])
group_slope_list <- c(group_slope_list, coef(m)[2])
}

hist(group_int_list)
hist(group_slope_list)
```

# Model selection and diagnostics ---- Unfinished! ------

## Model specification

We are considering models of the following form (with some coefficients potentially equal to zero):

$$
y_{ij} = \beta_{0,j} + \beta_{1,j} date\text{_}num + \beta_{2,j} mgstr + \beta_{3,j} bulk\text{_}purchase + \beta_{4,j} Primary\text{_}Reason + \beta_{5,j} source + \beta_{6,j} date\text{_}quarter + \epsilon_{ij}
$$

Where: 
* $\beta_{0,j}$ represents the intercept, which we plan to fit as a random intercept for each group $j$, so that it would have the form:
$\beta_{0,j} = \beta_0 + b_{0,j}, b_{0,j} \sim N(0,\tau_0^2)$

* The $\beta_k$ for $k=1,...,6$ are the coefficients for our fixed effect predictors. We will consider random slopes by group for these, in which case they would have the form: 
$\beta_{k,j} = \beta_k + b_{k,j}, b_{k,j} \sim N(0,\tau_k^2)$

We also have that $\epsilon_{ij} \perp b_{0,j} \overset{iid}{\sim} N(0,\sigma^2)$ for all $i, j$.

## Model specification

### Grouping variables and random intercepts

We see that `state` appears to be best for grouping; region is good as well.

```{r}
mod_s <- lmer(data=df, log_ppm ~ (1 |state) + date_num + date_quarter + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F)
mod_c <- lmer(data=df, log_ppm ~ (1 |city) + date_num + date_quarter + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F)
mod_r <- lmer(data=df, log_ppm ~ (1 |USA_region) + date_num + date_quarter + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F)

summary(mod_s)
# summary(mod_c)
# summary(mod_r)

LR <- 2*(logLik(mod_s)-logLik(mod_c))
0.5*(1-pchisq(LR[1],1)+1-pchisq(LR[1],1)) # choose s over c

LR <- 2*(logLik(mod_r)-logLik(mod_c))
0.5*(1-pchisq(LR[1],1)+1-pchisq(LR[1],1)) # choose c over r

LR <- 2*(logLik(mod_s)-logLik(mod_r))
0.5*(1-pchisq(LR[1],1)+1-pchisq(LR[1],1)) # choose s over r

# conclusion: choose state random intercept
```

```{r}
anova(mod_s,
      lmer(data=df, log_ppm ~ (1 |state) + date_quarter + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F))
# drop date_num

anova(lmer(data=df, log_ppm ~ (1 |state) + date_quarter + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F),
      lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F))
# drop date_quarter

anova(lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F),
      lmer(data=df, log_ppm ~ (1 |state) + bulk_purchase + Primary_Reason + source, REML=F))
# don't drop dose_per_mg

anova(lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F),
      lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + Primary_Reason + source, REML=F))
# don't srop bulk_purchase

anova(lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F),
      lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase + source, REML=F))
# do drop Primary_reason

anova(lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase + source, REML=F),
      lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase, REML=F))
# do drop source
```

Now investigate random slopes:

```{r}
anova(lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase, REML=F),
      lmer(data=df, log_ppm ~ (dose_per_mg |state) + dose_per_mg + bulk_purchase, REML=F))
# random slopes for dose_per_mg are good - BIC = 8624.6

anova(lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase, REML=F),
      lmer(data=df, log_ppm ~ (bulk_purchase |state) + dose_per_mg + bulk_purchase, REML=F))
# random slopes for nulk_purchase not helpful
```

Our candidate for final model is then:

```{r}
mod_final <- lmer(data=df, log_ppm ~ (dose_per_mg |state) + dose_per_mg + bulk_purchase, REML=F)
```

There are convergence inssues. Try removing states that are influential? Based on the "dfbetas" visualization below, they are Wyoming, Nevada, Alaska.

```{r}
mod_infl <- influence(mod_final,"state")
print(2/sqrt(length(unique(df_filter$state))))
diagn_df <- data.frame(dfbetas(mod_infl))
plot(mod_infl,which="dfbetas",xlab="DFBETAS",ylab="State")

resqq <- plot_resqq(mod_final)
ranefci <- plot_ranef(mod_final)
grid.arrange(resqq, ranefci, ncol = 2)
```

```{r}
df_filter <- df %>% filter(state != "Alaska" & state != "Wyoming" & state != "Nevada")
```

```{r}
mod_final <- lmer(data=df_filter, log_ppm ~ (dose_per_mg |state) + dose_per_mg + bulk_purchase, REML=F)

mod_infl <- influence(mod_final,"state")
print(2/sqrt(length(unique(df_filter$state))))
diagn_df <- data.frame(dfbetas(mod_infl))
plot(mod_infl,which="dfbetas",xlab="DFBETAS",ylab="State")

resqq <- plot_resqq(mod_final)
ranefci <- plot_ranef(mod_final)
grid.arrange(resqq, ranefci, ncol = 2)
```

See below: looks like the singularity is because some states don't have enough data for `bulk_purchase`. Ugh. Should we use `region` instead???

```{r}
View(data.frame(table(df$state, df$bulk_purchase)))
```

### 2021-03-21 STA 610 Case Study 1 Report_Team 4.Rmd
Various work on coding for modeling + selection

@@
 author: 
-  - Cole Juracek (Coordinator) 
+  - Cole Juracek (Programmer) 
   - Lauren Palazzo (Programmer) 
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE)
+knitr::opts_chunk$set(echo = FALSE, cache = TRUE)
 ```
@@
 if (!require(pacman)) install.packages("pacman")
-pacman::p_load(tidyverse, lme4, gridExtra, grid, ggplot2, lattice, redres)
+pacman::p_load(tidyverse, lme4, gridExtra, grid, ggplot2, lattice, redres, lubridate)
 #devtools::install_github("goodekat/redres")
@@
 
-#print(reason_df)
+print(reason_df)
 knitr::kable(reason_df_2, col.names = c("Level", "Freq"))
@@
 $$y_{ij} = \beta_{0,j} + \beta_{1} date_num + \beta_{2} mgstr + \beta_{3} bulk_purchase + \beta_{4} Primary_Reason + \beta_{5} source + \epsilon_{ij}$$
+
+$$log(y_{ij}) = \beta_{0,j} + \beta_{1} x_{ij1} + \beta_{2} x_{ij2} + \beta_{3} x_{ij3} + \beta_{4} x_{ij4} + \beta_{5} x_{ij5} + \epsilon_{ij}$$
+
 $$
@@
 
-* i stands for individuals and j stands for each state. 
-* $\beta_{0,j}$: Random intercept by state
-* $\beta_{1}$: Fixed effect for date_num, the date variable
-* $\beta_{2}$: Fixed effect for mgstr, the dosage strength of reported drug in milligrams.
-* $\beta_{3}$: Fixed effect for bulk_purchase is a binary variable that is represented by an indicator function I(bulk_purchase = Yes). 
-* $\beta_{4}$: Fixed effect for the categorical variable Primary_Reason which contains three categories:     - Other or unknown
-    - Self-treat
-    - Enjoymen
-* $\beta_{5}$ measures the fixed effect for the categorical variable source
+This linear model estimates the log price per miligram $log(y)$ of oxymorphonefor an invidual $i$ in state $j$ ($x_{ij}$). The following predictors are defined as:
+
+* $x_1$: The number of days since the recorded transaction and January 1st, 2000
+* $x_2$: Dosage strength in mg of the units purchased
+* $x_3$: Bulk purchase, an indicator for whether 10+ units were purchased at once
+* $x_4$: The primary reason for purchase. Factors include "other/unknown", "self-treatment", and "enjoyment"
+* $x_5$: Source of information (allows for reporting of transactions). Factors include "unkown", "internet", "heard it", and "personal"
 
@@
 
+- mgstr: good
+- bulk_purchase: good
+
 ```{r}
-mod1 <- lmer(data=df, log_ppm ~ (1 | state) + date_num + mgstr + bulk_purchase + Primary_Reason + source)
+mod1b.inf <- influence(mod1b, maxfun=1)
+infIndexPlot(mod1b.inf, vars='Cook')
+```
+
+```{r}
+#mod1 <- lmer(data=df, log_ppm ~ (1 | state) + date_num + mgstr + bulk_purchase + Primary_Reason + source)
+#mod1b <- lmer(data=df, log_ppm ~ (1 | state) + year + mgstr + bulk_purchase + Primary_Reason + source)
+big_data <- df[!influential,]
+mod1c <- lmer(data=big_data, log_ppm ~ (1 | state) + year + mgstr + bulk_purchase + Primary_Reason + source)
+qqmath(mod1c)
 ```

### 2021-03-21 report.Rmd
draft interpretation

--- 2021-03-20 STA 610 Case Study 1 Report_Team 4.Rmd
@@
 ```{r, echo=FALSE, warning=FALSE, message=FALSE}
-if (!require(pacman)) install.packages("pacman")
-pacman::p_load(tidyverse, lme4, gridExtra, grid, ggplot2, lattice, redres)
+library(tidyverse)
+library(lme4)
+library(gridExtra)
+library(grid)
+library(ggplot2)
+library(lattice)
 #devtools::install_github("goodekat/redres")
+library(redres)
 ```
@@
 ## Model
-
-Per the research question of interest, we are interested in exploring factors related to price per mg of oxymorphone.
-We are also specifically asked to control for location and how drug price changes with it.
-Cities seemed to be too noisy to use as a location predictor.
-On the flip side, region did not seem to be associated with a significant change in drug prices.
-Therefore, state provides the best granularity of accounting for location in this instance.
-
-Linear regression models are usually a good standard approach to this, providing clear interpretations of how various predictors are associated with drug price.
-As seen in the EDA, however, many states have excessively small sample sizes.
-This produces corresponding $\hat{\beta}_{OLS}$ estimates with large variance.
-We would like to be able to *partially pool* information across states, letting us borrow information to help stabilize state estimates with small sample sizes.
-
-This prompts a **mixed effects model**, where we instead model the state coefficients as having been drawn $N(0, \tau^2)$.
-
 #### Model Specification
@@
 ```{r, fig.align='center', warning=FALSE, message=FALSE, error=FALSE}
-# resqq <- plot_resqq(mod1)
-# ranefci <- plot_ranef(mod1)
-# grid.arrange(resqq, ranefci, ncol = 2)
+resqq <- plot_resqq(mod1)
+ranefci <- plot_ranef(mod1)
+resi <- plot_redres(mod1, type = "std_cond")
+grid.arrange(resi, resqq, ranefci, ncol = 3)
 ```
@@
 Model diagnostics discussions here.
+
+From the esiduals versus the fitted values plot, there are several outliers in the lower half of the plot. 
+From the normal quantile plot of the raw conditional residuals, the residuals are skewed and there are a lot of outliers. Thus, the residuals are not normally distributed. 
+From the normal quantile plot on random effect state, there are only two outliers. Thus, we can accept random effect of state is normally distributed.
 
@@
 
-* Bulk_purchase: “Not bulk purchase” is the reference group.
-Controlling for other variables, being bulk purchase increases the price per mg by exp(-0.16) = 0.8521. 
+* date_num: For the date_num variable, controlling for other variables, for every one unit increase in the date_num, the price per mg of oxymorphone reduces by a multiplicative effect of exp(-0.0001290) =  0.999871, about a 0.0129% decrease.
+
+* mgstr: For the mgstr variable, controlling for for other variables, for every one unit increase in mgstr, the price per mg of oxymorphone reduces by a multiplicative effect of exp(-0.0241262) =  0.9761625, about a 2.38375% decrease.
+
+* bulk_purchase: “Not bulk purchase” is the reference group.
+  * Controlling for other variables, being bulk purchase reduces the price per mg by a multiplicative effect of exp(-0.161) = 0.8521, about a 14.79% reduction.
+
 * Primary_Reason: “Primary_Reason: Other or Unknown” is the reference group. 
-Controlling for other variables, taking oxymorphone with the primary reason of self-treating tends to receive a exp(0.05) = 1.0513 times higher price per mg than people whose primary reason is “other or unknown”.
-Controlling for other variables, taking oxymorphone with the primary reason of enjoyment tends to receive a exp(0.02) = 1.0202 times higher price per mg than people whose primary reason is “other or unknown”.
-* Source: “Source: Unknown” is the reference group. 
-Controlling for other variables, purchasing oxymorphone with Internet source tends to receive a exp(0.02) = 1.0202 times higher price per mg than people with unknown drug source.
-Controlling for other variables, purchasing oxymorphone with personal source tends to receive a exp(0.07) = 1.0725 times higher price per mg than people with unknown drug source.
-Controlling for other variables, purchasing oxymorphone with “heard it” source tends to receive a exp(0.24) = 1.2712 times higher price per mg than people with unknown drug source.
+  * Controlling for other variables, taking oxymorphone with the primary reason of self-treating increase the price per mg by a multiplicative effect of exp(0.05) = 1.0513, about a 5.13% increase than the price per mg of oxymorphone for people whose primary reason is “other or unknown”.
+  * Controlling for other variables, taking oxymorphone with the primary reason of enjoyment tends to increase the price per mg by a multiplicative effect of exp(0.2) = 1.221403, about a 22.1403% increase than the price per mg of oxymorphone for people whose primary reason is “other or unknown”.
+
+* source: “Source: Unknown” is the reference group. 
+  * Controlling for other variables, purchasing oxymorphone with Internet source tends to increase the price per mg by a multiplicative effect of exp(0.02) = 1.0202, about a 2.02% increase than the price per mg of oxymorphone for people with unknown drug source.
+  * Controlling for other variables, purchasing oxymorphone with personal source tends to increase the price per mg by a multiplicative effect of exp(0.07) = 1.0725, about a 7.25% increase than the price per mg of oxymorphone for people with unknown drug source.
+  * Controlling for other variables, purchasing oxymorphone with “heard it” source tends to increase the price per mg by a multiplicative effect of exp(0.24) = 1.2712, about a 27.12% increase than the price per mg of oxymorphone for people with unknown drug source.
 
@@
 
+```{r, include=FALSE}
+library(sjstats)
+icc(mod1)
+```
+
+```{r}
+ICC <- 0.047
+knitr::kable(ICC, col.names = "Intraclass Correlation")
+```
+Intraclass correlation is 0.047, which means values are not very similar from the same state.
+
 ```{r, fig.align='center', fig.height=8}

### 2021-03-21 Draft Report 3.21.Rmd
draft

--- 2021-03-20 report_draft_fz_lp_4.Rmd
@@
 output:
+  pdf_document: default
   html_document:
@@
 
-load("~/Documents/Spring 2021/STA 610 - H Models/cs1/streetrx.RData")
+load("streetrx.RData")
 df <- streetrx %>% filter(api_temp=="oxymorphone", !is.na(ppm))
@@
 # summary(mod_r)
-
+BIC(mod_s)
+BIC(mod_c)
+BIC(mod_r)
 LR <- 2*(logLik(mod_s)-logLik(mod_c))
@@
 
+## Changes start here
+
 ```{r}
 mod_final <- lmer(data=df_filter, log_ppm ~ (dose_per_mg |state) + dose_per_mg + bulk_purchase, REML=F)
+BIC(mod_final)
+
+#model with no random slope is better
+#best model!!!
+mod_final <- lmer(data=df_filter, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase + source, REML=F)
+BIC(mod_final)
+#8412.338
+
+#class model
+mod1 <- lmer(data=df_filter, log_ppm ~ (1 | state) + date_num + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F)
+BIC(mod1)
 
@@
 ranefci <- plot_ranef(mod_final)
-grid.arrange(resqq, ranefci, ncol = 2)
+#take a long time to run
+#resi <- plot_redres(mod_final, type = "std_cond")
+grid.arrange(resqq, ranefci, ncol = 3)
 ```
@@
 
+## Results
+#### Fixed Effects
+```{r, message=FALSE}
+#knitr::kable(fixef(mod1))
+#confint(mod1)[3:11,]
+
+# extract coefficients
+coefs <- data.frame(coef(summary(mod_final)))
+# use normal distribution to approximate p-value
+coefs$p.z <- 2 * (1 - pnorm(abs(coefs$t.value)))
+knitr::kable(coefs, col.names = c("Estimate", "Std.Error", "t-value", "P-value"))
+```
+
+#### need to change numbers after finalizing model
+* date_num: For the date_num variable, controlling for other variables, for every one unit increase in the date_num, the price per mg of oxymorphone reduces by a multiplicative effect of exp(-0.0001290) =  0.999871, about a 0.0129% decrease.
+
+* mgstr: For the mgstr variable, controlling for for other variables, for every one unit increase in mgstr, the price per mg of oxymorphone reduces by a multiplicative effect of exp(-0.0241262) =  0.9761625, about a 2.38375% decrease.
+
+* bulk_purchase: “Not bulk purchase” is the reference group.
+  * Controlling for other variables, being bulk purchase reduces the price per mg by a multiplicative effect of exp(-0.161) = 0.8521, about a 14.79% reduction.
+
+* Primary_Reason: “Primary_Reason: Other or Unknown” is the reference group. 
+  * Controlling for other variables, taking oxymorphone with the primary reason of self-treating increase the price per mg by a multiplicative effect of exp(0.05) = 1.0513, about a 5.13% increase than the price per mg of oxymorphone for people whose primary reason is “other or unknown”.
+  * Controlling for other variables, taking oxymorphone with the primary reason of enjoyment tends to increase the price per mg by a multiplicative effect of exp(0.2) = 1.221403, about a 22.1403% increase than the price per mg of oxymorphone for people whose primary reason is “other or unknown”.
+
+* source: “Source: Unknown” is the reference group.
+
+  * Controlling for other variables, purchasing oxymorphone with Internet source tends to increase the price per mg by a multiplicative effect of exp(0.02) = 1.0202, about a 2.02% increase than the price per mg of oxymorphone for people with unknown drug source.
+  * Controlling for other variables, purchasing oxymorphone with personal source tends to increase the price per mg by a multiplicative effect of exp(0.07) = 1.0725, about a 7.25% increase than the price per mg of oxymorphone for people with unknown drug source.
+  * Controlling for other variables, purchasing oxymorphone with “heard it” source tends to increase the price per mg by a multiplicative effect of exp(0.24) = 1.2712, about a 27.12% increase than the price per mg of oxymorphone for people with unknown drug source.
+
+#### Random Effects
 ```{r}
-View(data.frame(table(df$state, df$bulk_purchase)))
+#VarCorr(mod1)
+Variance <- c(0.04748, 0.96369)
+Std.Dev <- c(0.2179, 0.9817)
+Groups <- c("state (Intercept)", "Residual")
+
+REvar <- cbind(Groups, Variance, Std.Dev)
+
+knitr::kable(REvar, col.names = c("Group", "Variance", "Std.Dev"))
 ```
+The within-state variance $\sigma^2$ is 0.96369, which is much larger than the across-state variance $\tau^2$, 0.04748.
+
+```{r, include=FALSE}
+library(sjstats)
+icc(mod_final)
+```
+
+```{r}
+ICC <- 0.047
+knitr::kable(ICC, col.names = "Intraclass Correlation")
+```
+Intraclass correlation is 0.047, which means values are not very similar from the same state.
+
+```{r, fig.align='center', fig.height=8}
+dotplot(ranef(mod_final, condVar=TRUE))$state
+```
+From the result of the random effects, we can see that the random intercept for Tennessee is the highest and it is significantly different from zero.
+The random intercept for Nevada is the lowest and this intercept is also significant.
+
+## Appendix
+```{r, ref.label=knitr::all_labels(),echo=TRUE,eval=FALSE}
+```

### 2021-03-21 Draft Report 3.21.Rmd
updated the interpretations of the fixed effects

@@
 #### need to change numbers after finalizing model
-* date_num: For the date_num variable, controlling for other variables, for every one unit increase in the date_num, the price per mg of oxymorphone reduces by a multiplicative effect of exp(-0.0001290) =  0.999871, about a 0.0129% decrease.
-
-* mgstr: For the mgstr variable, controlling for for other variables, for every one unit increase in mgstr, the price per mg of oxymorphone reduces by a multiplicative effect of exp(-0.0241262) =  0.9761625, about a 2.38375% decrease.
+* dose_per_mg: Controlling for other variables, for every one milligram increase of dosage, the price per mg of oxymorphone increases by a multiplicative effect of exp(-0.01524395) = 0.9848717, about a 1.51% reduction.
 
 * bulk_purchase: “Not bulk purchase” is the reference group.
-  * Controlling for other variables, being bulk purchase reduces the price per mg by a multiplicative effect of exp(-0.161) = 0.8521, about a 14.79% reduction.
-
-* Primary_Reason: “Primary_Reason: Other or Unknown” is the reference group. 
-  * Controlling for other variables, taking oxymorphone with the primary reason of self-treating increase the price per mg by a multiplicative effect of exp(0.05) = 1.0513, about a 5.13% increase than the price per mg of oxymorphone for people whose primary reason is “other or unknown”.
-  * Controlling for other variables, taking oxymorphone with the primary reason of enjoyment tends to increase the price per mg by a multiplicative effect of exp(0.2) = 1.221403, about a 22.1403% increase than the price per mg of oxymorphone for people whose primary reason is “other or unknown”.
+  * Controlling for other variables, bulk purchase reduces the price per mg of oxymorphone by a multiplicative effect of exp(-0.17013501) = 0.8435509, about a 15.64% reduction.
 
 * source: “Source: Unknown” is the reference group. 
-
-  * Controlling for other variables, purchasing oxymorphone with Internet source tends to increase the price per mg by a multiplicative effect of exp(0.02) = 1.0202, about a 2.02% increase than the price per mg of oxymorphone for people with unknown drug source.
-  * Controlling for other variables, purchasing oxymorphone with personal source tends to increase the price per mg by a multiplicative effect of exp(0.07) = 1.0725, about a 7.25% increase than the price per mg of oxymorphone for people with unknown drug source.
-  * Controlling for other variables, purchasing oxymorphone with “heard it” source tends to increase the price per mg by a multiplicative effect of exp(0.24) = 1.2712, about a 27.12% increase than the price per mg of oxymorphone for people with unknown drug source.
+  * Controlling for other variables, purchasing oxymorphone with Internet source tends to increase the price per mg by a multiplicative effect of exp(0.12686001) = 1.135258, about a 13.53% increase than the price per mg of oxymorphone for people with unknown drug source.
+  * Controlling for other variables, purchasing oxymorphone with Internet Pharmacy source tends to increase the price per mg by a multiplicative effect of exp(-0.44349281) = 0.6417908, about a 35.82% reduction of the price per mg of oxymorphone for people with unknown drug source.
+  * Controlling for other variables, purchasing oxymorphone with personal source tends to increase the price per mg by a multiplicative effect of exp(0.06140303) = 1.063327, about a 6.33% increase than the price per mg of oxymorphone for people with unknown drug source.
+  * Controlling for other variables, purchasing oxymorphone with “heard it” source tends to increase the price per mg by a multiplicative effect of exp(0.18025958) = 1.197528, about a 19.75% increase than the price per mg of oxymorphone for people with unknown drug source.
 

### 2021-03-21 Draft Report 3.21.Rmd
updated the random effect results

@@
 #VarCorr(mod1)
-Variance <- c(0.04748, 0.96369)
-Std.Dev <- c(0.2179, 0.9817)
+Variance <- c(0.0233, 0.5028)
+Std.Dev <- c(0.1526, 0.7091)
 Groups <- c("state (Intercept)", "Residual")
@@
 ```
-The within-state variance $\sigma^2$ is 0.96369, which is much larger than the across-state variance $\tau^2$, 0.04748.
+The within-state variance $\sigma^2$ is 0.5028, which is much larger than the across-state variance $\tau^2$, 0.0233.
 
@@
 From the result of the random effects, we can see that the random intercept for Tennessee is the highest and it is significantly different from zero.
-The random intercept for Nevada is the lowest and this intercept is also significant.
+The random intercept for Arizona is the lowest and this intercept is also significant.
 

### 2021-03-21 STA 610 Case Study 1 Report_Team 4.Rmd
Code added for finding influential observations based on Cook's distance

@@
 mod1b.inf <- influence(mod1b, maxfun=1)
+
 infIndexPlot(mod1b.inf, vars='Cook')
+
+# Find the most influential observations
+high_inf_idx <- order(cooks.distance(mod1b.inf), decreasing = TRUE)
+
+# Take a look at some of the most influential observations
+df[high_inf_idx[1:25], ]
 ```
@@
 #mod1b <- lmer(data=df, log_ppm ~ (1 | state) + year + mgstr + bulk_purchase + Primary_Reason + source)
-big_data <- df[!influential,]
-mod1c <- lmer(data=big_data, log_ppm ~ (1 | state) + year + mgstr + bulk_purchase + Primary_Reason + source)
+better_data <- df %>% filter(ppm > 0.2)
+mod1c <- lmer(data=better_data, log_ppm ~ (1 | state) + date_num + year + mgstr + bulk_purchase + Primary_Reason + source)
 qqmath(mod1c)

### 2021-03-21 Draft Report 3.21.Rmd
corrected a minor mistake in fixed effect interpretations

@@
   * Controlling for other variables, purchasing oxymorphone with Internet source tends to increase the price per mg by a multiplicative effect of exp(0.12686001) = 1.135258, about a 13.53% increase than the price per mg of oxymorphone for people with unknown drug source.
-  * Controlling for other variables, purchasing oxymorphone with Internet Pharmacy source tends to increase the price per mg by a multiplicative effect of exp(-0.44349281) = 0.6417908, about a 35.82% reduction of the price per mg of oxymorphone for people with unknown drug source.
+  * Controlling for other variables, purchasing oxymorphone with Internet Pharmacy source tends to reduce the price per mg by a multiplicative effect of exp(-0.44349281) = 0.6417908, about a 35.82% reduction of the price per mg of oxymorphone for people with unknown drug source.
   * Controlling for other variables, purchasing oxymorphone with personal source tends to increase the price per mg by a multiplicative effect of exp(0.06140303) = 1.063327, about a 6.33% increase than the price per mg of oxymorphone for people with unknown drug source.

### 2021-03-21 STA 610 Case Study 1 Report_Team 4.Rmd
Added most recent changes from Lauren's EDA

@@
 author:
-  - Cole Juracek (Programmer) 
+- Cole Juracek (Coordinator)
 - Lauren Palazzo (Programmer)
@@
 date: "`r Sys.Date()`"
-output: pdf_document
+output:
+  html_document:
+    df_print: paged
 ---
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, cache = TRUE)
+knitr::opts_chunk$set(echo = FALSE)
 ```
@@
 ```{r, echo=FALSE, warning=FALSE, message=FALSE}
-if (!require(pacman)) install.packages("pacman")
-pacman::p_load(tidyverse, lme4, gridExtra, grid, ggplot2, lattice, redres, lubridate)
-#devtools::install_github("goodekat/redres")
+library(tidyverse)
+library(lme4)
+library(gridExtra)
+library(grid)
+library(ggplot2)
+library(lattice)
+# devtools::install_github("goodekat/redres")
+library(redres)
+library(stringr)
+library(influence.ME)
 ```
@@
 ```{r, echo = FALSE}
-load("streetrx.RData")
+#load("~/Documents/Spring 2021/STA 610 - H Models/cs1/streetrx.RData")
+load('streetrx.rdata')
 df <- streetrx %>% filter(api_temp=="oxymorphone", !is.na(ppm))
 ```
+
+# Introduction
+
+This study is an exploration of the relationship between the sale price per milligram of diverted pharmaceutical substances and other factors related to the sale, as reflected in self-reported data from the website StreetRx.
+StreetRx maintains a partnership with the Researched Abuse, Diversion, and Addiction-Related Surveillance System (RADARS), which aims for these data to inform public health measures such as health policy related to illicit drugs.
+
+[insert something about our specific drug here]
+
+Among the data available to us are location data pertaining to the city, state, and United States region of the sale.
+These variables present a natural grouping structure for the price data, which suggest that we employ a hierarchical model in our analysis.
+The benefits of a hierarchical model include the ability to "borrow information" from the data set as a whole to stabilize estimates of price for groups (e.g., states) with relatively small sample size, and to establish a estimated distribution of how price varies with respect to groups, which would in theory allow generalization to new groups not included in the original analysis.
+
+# Exploratory data analysis and data cleaning
+
+## Data cleaning
+
+### Group variables
+
+We performed an initial examination of two grouping variables, `state` and `USA_region`.
+The state North Dakota had only one observation, and as such was not considered a well-defined "group" for which a variance could be estimated; thus this state's observation was removed.
+States labeled simply as "USA" were recoded to "Unknown".
+Region was also sometimes labeled as uknown or other; these observations coincided with the unknown states (see appendix Table X).
+(The variable `city` as a grouping variable is addressed later in the report.)
 
@@
 
+### PUT IN APPENDIX: the state-by-region table of coutns below
+# table(df$state, df$USA_region)
+```
+
+### Date variable
+
+We considered two different ways to engineer the date-of-sale feature.
+One intuitive and commonly-used approach would be to recode this as a continuous variables counting days from some start date, the intent of which would be to capture the influence of inflation for this price data that spans years.
+Another approach would be to consider whether there might be an effect of seasonality on drug prices.
+
+For recoding date into a continuous variable, `date_num`, we filter out observations dated prior to the year 2010, as that is the year that the StreetRx web site was founded.
+For the seasonality variable, `date_quarter`, we extract the "quarter" of the year (numbered 1 through 4) as its own variable.
+Below we see the distributions of these variables:
+
+```{r, echo = FALSE}
+# Change date to numeric (easier for interpretation)
+
+# this sets 
+df$date_num <- as.numeric(as.Date(df$price_date, "%m/%d/%y")) - 14621 # sets 1/12/2010 (first post-2010 date) as zero
+
+tempdf <- df %>% select(date_num, price_date)
+
+# this filters out all data before the 1/1/2000, which probably shouldn't be included
+df <- df %>% filter(date_num >= 0)
+```
+
+```{r}
+df$date_quarter <- as.numeric(substr(df$yq_pdate, 5, 5))
+```
+
+```{r}
+hist(df$date_num)
+knitr::kable(table(df$date_quarter))
+```
+
+### Sources variable
+
+As the variable `source` that describes the source of the price information is a user-input field, it is an inherently "messy" variable that calls for some recoding in order to make the levels of the variable more meaningful and interpretable, and to avoid overfitting.
+In the case of `source`, some user input is simply blank, while a small number of other entries indicate in some other way that the source is unknown or unclear.
+Since it is quite uncommon in the data for a source to be entered yet indicated as unknown, we decided to drop these unkown-source observations as being indicative of poor data quality.
+(If the user cannot identify their source of price information, why should we trust their price data?) We decided to distinguish between "Internet pharmacy" sources and other internet sources, since there may be a distinction between a source that is selling a drug versus perhaps just discussing a drug price in a forum, which may yield less legitimate information.
+Ultimately, we recoded the categorical variable "sources" into the following levels.
+
+```{r, echo = FALSE}
+# recode sources
+
+# remove unused levels
+df$source <- droplevels(df$source)
+
+source_df <- data.frame(table(df$source))
+
+# levels(df$source)
+levels(df$source)[1] <- "Blank"
+
+source_df <- data.frame(table(df$source))
+
+df$source <- recode_factor(df$source, "Blank" = "Blank",
+                           "Internet" = "Internet",
+                           "Internet Pharmacy" = "Internet Pharmacy",
+                           "Personal" = "Personal",
+                           "Heard it" = "Heard it",
+                           "Idk" = "Unknown",
+                           "w" = "Unknown",
+                           "STREET PRICE" = "Heard it",
+                           .default = "Internet") # everything else is a URL
+
+source_df_2 <- data.frame(table(df$source))
+
+# drop unknowns
+
+df <- df %>% filter(df$source != "Unknown")
+df$source <- droplevels(df$source)
+
+knitr::kable(table(df$source), col.names = c("Level", "Freq"))
+```
+
+### Bulk purchase variable
+
+Our visualization of the bulk purchase variable indicates nothing unexpected that calls for data cleaning.
+
+```{r}
+#table(df$bulk_purchase, useNA = "always")
+# looks fine
+```
+
+### Purchase reason variable
+
+`Primary_Reason` for purchase is another user-input variable that calls for recoding.
+Similarly to the `source` variable, we retain the information about which entries were blank.
+We also try to distinguish this from a different response level that is not blank, yet indicates the user did not provide a reason.
+We further still distinguished between these responses and a non-blank response of refusal to answer.
+We strove to avoid making presumptions of no difference between these similar-seeming response types, as they may indicate variations in data processing that hold some predictive power that we are not yet aware of.
+We ultimately recoded the categorical variable `Primary_Reason` into the following levels.
+
+```{r}
+# remove unused levels
+df$Primary_Reason <- droplevels(df$Primary_Reason)
+
+reason_df <- data.frame(table(df$Primary_Reason))
+
+# levels(df$Primary_Reason)
+levels(df$Primary_Reason)[1] <- "Blank"
+
+reason_df <- data.frame(table(df$Primary_Reason))
+
+df$Primary_Reason <- recode_factor(df$Primary_Reason, "Blank" = "Blank",
+                                   "0 Reporter did not answer the question" = "Left unanswered",
+                                   "Other or unknown" = "Other or unknown",
+                                   "9 To self-treat my pain" = "Self-treat",
+                                   "3 To prevent or treat withdrawal" = "Self-treat",
+                                   "10 To treat a medical condition other than pain" = "Self-treat",
+                                   "4 For enjoyment/to get high" = "Enjoyment",
+                                   "5 To resell" = "Resell",
+                                   "8 Prefer not to answer" = "Refuse to answer",
+                           .default = "Other or unknown")
+
+reason_df_2 <- data.frame(table(df$Primary_Reason))
+
+#print(reason_df)
+knitr::kable(reason_df_2, col.names = c("Level", "Freq"))
+```
+
+### Drug strength variable
+
+We renamed this variable `dose_per_mg` to make its meaning clearer to the reader.
+This is a continuous variable that nevertheless tends to take on only a few values, as depicted in the histogram.
+We considered a recoding of this variable into levels, but decided against it as the dose per milligram technically may take on any value and the generalizability of our model may be improved if we are able to consider dose levels that are in a realistic range yet not present in our current data.
+
+```{r}
+hist(df$mgstr)
+# rename variable for ease of interpretability
+df <- df %>% rename(dose_per_mg = mgstr)
+```
+
+### Response variable: Price per milligram
+
+We are considering fitting a hierarchical model with a response variable that has normally distributed errors, and so we would like our response variable, here `ppm`, or price per milligram, to be close to normally-distributed.
+The histogram of the `ppm` is extremely right skewed (not unexpected for data related to financial measures such as price, salary, earnings, etc.), while the distribution of the log(`ppm`) looks roughly normal.
+Thus, we will use `log_ppm` as the dependent variable in our model.
+
+```{r, fig.height=3, fig.align='center'}
 # log transform ppm
 df$log_ppm <- log(df$ppm)
-```
 
-## EDA
-#### PPM
-
-```{r, fig.height=3, fig.align='center'}
 hist_ppm <- ggplot(df, aes(x=ppm)) + 
@@
 ```
-The histogram of the original ppm is extremely right skewed, while the distribution of the log(ppm) looks roughly normal. Thus, we will use the log(ppm) as the dependent variable in our model.
 
-```{r, echo = FALSE}
-# Change date to numeric (easier for interpretation)
+In order to cut off extreme price values that may skew the results for typical prices, we restrict the `log_ppm` distribution to more realistic and representative values:
 
-# this sets 
-df$date_num <- as.numeric(as.Date(df$price_date, "%m/%d/%y")) - 10957 # sets 1/1/2000 as zero
-
-tempdf <- df %>% select(date_num, price_date)
-
-# this filters out all data before the 1/1/2000, which probably shouldn't be included
-df <- df %>% filter(date_num >= 0)
-
+```{r}
+df <- df %>% filter(log_ppm > -2 & log_ppm < 2)
 ```
 
-#### Sources
+## Exploratory data analysis for random intercepts
 
-We recoded the categorical variable "sources" into the following 4 levels.
-```{r, echo = FALSE}
-# recode sources
+We now take a closer look at our grouping variables with a focus on the question of whether random intercepts might be appropriate.
 
-# remove unused levels
-df$source <- droplevels(df$source)
+### State
 
-source_df <- data.frame(table(df$source))
+We show boxplots of `log_ppm` by `state`, arranged in the order of increasing sample size from the bottom to the top.
 
-# levels(df$source)
-levels(df$source)[1] <- "Unknown"
-
-source_df <- data.frame(table(df$source))
-
-df$source <- recode_factor(df$source, "Unknown" = "Unknown",
-                           "Internet" = "Internet",
-                           "Internet Pharmacy" = "Internet",
-                           "Personal" = "Personal",
-                           "Heard it" = "Heard it",
-                           "Idk" = "Unknown",
-                           "w" = "Unknown",
-                           "STREET PRICE" = "Heard it",
-                           .default = "Internet") # everything else is a URL
-
-source_df_2 <- data.frame(table(df$source))
-
-knitr::kable(source_df_2, col.names = c("Level", "Freq"))
-```
-
-```{r}
-#table(df$bulk_purchase, useNA = "always")
-# looks fine
-```
-
-#### Primary_Reason
-
-We recoded the categorical variable "Primary_Reason" into the following 3 levels.
-```{r}
-# remove unused levels
-df$Primary_Reason <- droplevels(df$Primary_Reason)
-
-reason_df <- data.frame(table(df$Primary_Reason))
-
-# levels(df$Primary_Reason)
-levels(df$Primary_Reason)[1] <- "Other or unknown"
-
-reason_df <- data.frame(table(df$Primary_Reason))
-
-df$Primary_Reason <- recode_factor(df$Primary_Reason, "Other or unknown" = "Other or unknown",
-                                    "9 To self-treat my pain" = "Self-treat",
-                                    "3 To prevent or treat withdrawal" = "Self-treat",
-                                    "10 To treat a medical condition other than pain" = "Self-treat",
-                                    "4 For enjoyment/to get high" = "Enjoyment",
-                            .default = "Other or unknown")
-
-reason_df_2 <- data.frame(table(df$Primary_Reason))
-
-print(reason_df)
-knitr::kable(reason_df_2, col.names = c("Level", "Freq"))
-```
-
-#### State
 ```{r, fig.align='center', fig.height=6}
@@
 geom_boxplot() +
-labs(title="state", x="log PPM by state",y="Log PPM")
+labs(title="State", x="Log PPM by state",y="Log PPM")
 ```
-The boxplot of log PPM by state is arranged in the order of increasing sample size from the bottom to the top.
 
-#### Region
+We see that some states with small sample sizes have distributions of `log_price` that seem unstable, with respect to how much they differ from other boxplots; this provides support for the idea that the stabilizing effect of a hierarchical model would benefit our analysis.
+The fact that some distributions differ from each other in a noteworthy way (e.g., boxplots hardly overlapping) suggests group-based differences in price that should be taken into account in modeling, as opposed to lumping all states' data together without distinction.
+
+### Region
+
+Viewing a similar plot for the regions of the USA, we do see some variability in price among regions, but it is not as pronounced as for the states.
 
@@
 
-#### City
+### City
+
+We find that there are many cities with a sample size of one, which makes `city` perhaps not the most natural choice for a grouping variable.
+Example visualizations shown here are for all cities in the states of Florida and California.
+
 ```{r, fig.align='center'}
@@
 
-#### mgstr
+We decided to focus on `state` as our primary candidate for a grouping variable moving forward.
+
+## Exploratory data analysis for random slopes
+
+To consider whether we may want to include random slopes by state for our predictor variables, we examined plots of the predictor versus response among each state individually to determine whether it seemed that slopes fit to the group-level data would differ.
+(We present a sample of visualizations here rather than visualizations for every state, due to space constraints.)
+
+For out continuous variables (continuous date and dosage strength), in by-state plots we often saw strikingly similar patterns in predictor vs.
+response, suggesting that random slopes may not be necessary for these variables.
+
+### dose_per_mg
 ```{r, fig.align='center'}
-# random slopes for mgstr?
-
-ggplot(data = df[df$state=="Florida" | df$state=="Texas" | df$state=="North Carolina",], aes(x = mgstr, y = log_ppm)) + 
+ggplot(data = df[df$state=="Florida" | df$state=="Texas" | df$state=="North Carolina",], aes(x = dose_per_mg, y = log_ppm)) + 
   geom_point() +
@@
 
-#### bulk_purchase
+### date_num
+```{r, fig.align='center'}
+ggplot(data = df[df$state=="Florida" | df$state=="Texas" | df$state=="North Carolina",], aes(x = date_num, y = log_ppm)) + 
+  geom_point() +
+  facet_grid(~ state)
+```
+
+For categorical variables, we generally observed was there was indeed variation among the states, but it was not clear whether this variation had any well-defined structure.
+It was not clear whether our models would benefit from random slopes for these variables, so we decided to further investigate this in our modeling process.
+
+### bulk_purchase
 ```{r, fig.align='center'}
@@
 
-#### date_num
-```{r, fig.align='center'}
-# categorical variables -- looking at South only
-ggplot(data = df[df$state=="Florida" | df$state=="Texas" | df$state=="North Carolina",], aes(x = date_num, y = log_ppm)) + 
-  geom_point() +
+### date_quarter
+```{r}
+ggplot(data = df[df$state=="Florida" | df$state=="Texas" | df$state=="North Carolina",], aes(x = date_quarter, y = log_ppm, group = date_quarter)) + 
+  geom_boxplot() +
   facet_grid(~ state)
+```
+
+### Primary_Reason
+```{r}
+ggplot(data = df[df$state=="Florida" | df$state=="Texas" | df$state=="North Carolina",], aes(x = Primary_Reason, y = log_ppm, group = Primary_Reason)) + 
+  geom_boxplot() +
+  facet_grid(~ state)
+```
+
+### source
+```{r}
+ggplot(data = df[df$state=="Florida" | df$state=="Texas" | df$state=="North Carolina",], aes(x = source, y = log_ppm, group = source)) + 
+  geom_boxplot() +
+  facet_grid(~ state)
+```
+## More EDA: slope and intercept distributions for predictors by state
+
+To get a better sense of how state-level groups and slopes are distributed, we created the following intercepts for our coninuous variables:
+
+### Dose strength variable
+
+```{r}
+group_int_list <- NULL 
+group_slope_list <- NULL
+
+for(statej in
+  unique(df$state))
+{
+m <- lm(data = df[df$state==statej, ], log_ppm ~ dose_per_mg)
+
+group_int_list <- c(group_int_list, coef(m)[1])
+group_slope_list <- c(group_slope_list, coef(m)[2])
+}
+
+hist(group_int_list)
+hist(group_slope_list)
+```
+
+### Continuous date variable
+
+```{r}
+group_int_list <- NULL 
+group_slope_list <- NULL
+
+for(statej in
+  unique(df$state))
+{
+m <- lm(data = df[df$state==statej, ], log_ppm ~ date_num)
+
+group_int_list <- c(group_int_list, coef(m)[1])
+group_slope_list <- c(group_slope_list, coef(m)[2])
+}
+
+hist(group_int_list)
+hist(group_slope_list)
 ```

### 2021-03-21 STA 610 Case Study 1 Report_Team 4.Rmd
Start of writeup on model selection

@@
 # devtools::install_github("goodekat/redres")
-library(redres)
+#library(redres)
 library(stringr)
@@
 ### dose_per_mg
+
 ```{r, fig.align='center'}
@@
 ### date_num
+
 ```{r, fig.align='center'}
@@
 ### source
+
 ```{r}
-ggplot(data = df[df$state=="Florida" | df$state=="Texas" | df$state=="North Carolina",], aes(x = source, y = log_ppm, group = source)) + 
+ggplot(data = df %>% filter(state %in% c('Florida', 'Texas', 'North Carolina')), aes(x = source, y = log_ppm, group = source)) + 
   geom_boxplot() +
@@
 This prompts a **mixed effects model**, where we instead model the state coefficients as having been drawn $N(0, \tau^2)$.
+
+### Model Selection
+
+Based off of the results of the EDA, we want to initially fit a model with the following variables:
+
+- source
+- dose_per_mg
+- bulk_purchase
+- Primary_Reason
+- date_num
+- date_quarter
+
+In terms of the research question, we would also like to group based on location. There are 3 options for this:
+
+- Region
+- State
+- City
+
+In order to test whether keeping a variable in the model is "worth it", we may perform a likelihood ratio test for nested models.
+For adding $k$ fixed effects, the LRT test statistic follows a $\chi^2(k)$ distribution, and for adding $k$ random effect, the LRT test statistic follows a 50-50 mixture of $\chi^2(k)$ and $\chi^2(k + 1)$ distribution.
+For non-nested models, we can use BIC as a rough indicator of which model is "better".
+BIC balances likelihood of model with how parsimonious it is.
 

### 2021-03-21 STA 610 Case Study 1 Report_Team 4.Rmd
Added model selection for grouping variable

@@
 
+First, let's decide what to group on for region:
+
+```{r}
+# Fit 3 models grouping on different location specificity
+mod_s <- lmer(data=df, log_ppm ~ (1 |state) + date_num + date_quarter + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F)
+print(paste("BIC for grouping on state", round(BIC(mod_s), 3)))
+
+mod_c <- lmer(data=df, log_ppm ~ (1 |city) + date_num + date_quarter + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F)
+print(paste("BIC for grouping on city", round(BIC(mod_c), 3)))
+
+mod_r <- lmer(data=df, log_ppm ~ (1 |USA_region) + date_num + date_quarter + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F)
+print(paste("BIC for grouping on region", round(BIC(mod_r), 3)))
+```
+
+State provides the best level of BIC, so we will choose this for our grouping variable.
+
+Now let's investigate dropping various
+
 #### Model Specification

### 2021-03-21 STA 610 Case Study 1 Report_Team 4.Rmd
Perform nested model tests to decide on dropping variables

@@
 
-Now let's investigate dropping various
+Now let's investigate dropping various fixed effects and see results of our hypothesis tests:
+
+```{r}
+# Test date_num
+test <- anova(mod_s,
+      lmer(data=df, log_ppm ~ (1 |state) + date_quarter + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F))
+print(test$`Pr(>Chisq)`)  # Drop date_num
+
+# Test date_quarter
+test <- anova(lmer(data=df, log_ppm ~ (1 |state) + date_quarter + dose_per_mg + 
+                    bulk_purchase + Primary_Reason + source, REML=F),
+      lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F))
+print(test$`Pr(>Chisq)`) # drop date_quarter
+
+# Test dose_per_mg
+test <- anova(lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F),
+      lmer(data=df, log_ppm ~ (1 |state) + bulk_purchase + Primary_Reason + source, REML=F))
+print(test$`Pr(>Chisq)`)  # Don't drop dose_per_mg
+
+# Test bulk purchase
+test <- anova(lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F),
+      lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + Primary_Reason + source, REML=F))
+print(test$`Pr(>Chisq)`) # Don't drop bulk_purchase
+
+# Test primary reason
+test <- anova(lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F),
+      lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase + source, REML=F))
+print(test$`Pr(>Chisq)`) # Don't drop Primary_reason
+
+# Test source
+test <- anova(lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase + source, REML=F),
+      lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase, REML=F))
+print(test$`Pr(>Chisq)`) # Don't drop source
+```
+
+Thus these results suggest we want to drop
 

### 2021-03-21 STA_610_Case_Study_1_Report_Team_4.Rmd
File renamed

--- 2021-03-21 STA 610 Case Study 1 Report_Team 4.Rmd
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE)
+knitr::opts_chunk$set(echo = FALSE, cache=TRUE)
 ```
@@
 library(influence.ME)
+library(knitr)
 ```
@@
 
+First, let's decide what to group on for region:
+
+```{r}
+# Fit 3 models grouping on different location specificity
+mod_s <- lmer(data=df, log_ppm ~ (1 |state) + date_num + date_quarter + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F)
+
+mod_c <- lmer(data=df, log_ppm ~ (1 |city) + date_num + date_quarter + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F)
+
+mod_r <- lmer(data=df, log_ppm ~ (1 |USA_region) + date_num + date_quarter + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F)
+```
+
+```{r}
+BIC <- sapply(c(mod_s, mod_c, mod_r), BIC)
+re_results <- data.frame('Grouping' = c('State', 'City', 'Region'), 'BIC' = BIC)
+kable(re_results, caption = 'BIC for Models with Different Random Intercepts')
+```
+
+State provides the best level of BIC, so we will choose this for our grouping variable.
+
+Now let's investigate dropping various fixed effects and see results of our hypothesis tests:
+
+```{r}
+# Test date_num
+drop_date_num <- anova(mod_s,
+      lmer(data=df, log_ppm ~ (1 |state) + date_quarter + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F))
+date_num_pval <- drop_date_num$`Pr(>Chisq)`[2]
+
+# Test date_quarter
+drop_quarter <- anova(lmer(data=df, log_ppm ~ (1 |state) + date_quarter + dose_per_mg + 
+                    bulk_purchase + Primary_Reason + source, REML=F),
+      lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F))
+quarter_pval <- drop_quarter$`Pr(>Chisq)`[2]
+
+# Test dose_per_mg
+drop_dose <- anova(lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F),
+      lmer(data=df, log_ppm ~ (1 |state) + bulk_purchase + Primary_Reason + source, REML=F))
+dose_pval <- drop_dose$`Pr(>Chisq)`[2]
+
+# Test bulk purchase
+drop_bulk <- anova(lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F),
+      lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + Primary_Reason + source, REML=F))
+bulk_pval <- drop_bulk$`Pr(>Chisq)`[2]
+
+# Test primary reason
+drop_reason <- anova(lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F),
+      lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase + source, REML=F))
+reason_pval <- drop_reason$`Pr(>Chisq)`[2]
+
+# Test source
+drop_source <- anova(lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F),
+      lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase + Primary_Reason, REML=F))
+source_pval <- drop_source$`Pr(>Chisq)`[2]
+```
+
+The outline is as follows:
+
+* Start with date_num, date_quarter, dose_per_mg, bulk_purchase, Primary_Reason, source
+* Perform a LRT between competing models with and without the date variable.
+Observe a p-value of `r round(date_num_pval, 3)` and fail to reject the null (simpler model).
+Drop `date_num` from the subsequent model tests
+* Perform a LRT between competing models with and without the quarter of the sale.
+Observe a p-value of `r round(quarter_pval, 3)` and fail to reject the null (simpler model).
+Drop `date_quarter` from the subsequent model tests
+* Perform a LRT between competing models with and without dose per mg. Observe a p-value of ~`r round(dose_pval, 3)` and reject the null (simpler model). Decide on model keeping `dose_per_mg`
+* Perform a LRT between competing models with and without the bulk purchase predictor.
+Observe a p-value of ~`r round(bulk_pval, 3)` and reject the null (simpler model).
+Decide on model keeping `bulk_purchase`
+* Perform a LRT between competing models with and without the primary reason predictor.
+Observe a p-value of `r round(reason_pval, 3)` and reject the null (simpler model).
+Decide on keeping `primary_reason` in the model
+* Perform a LRT between competing models with and without the source predictor.
+Observe a p-value of ~`r round(source_pval, 3)` and reject the null (simpler model).
+Decide on keeping `source` in the model
+
+Thus these results suggest we want to drop `date_num` and `date_quarter`, while keeping
+
+* dose_per_mg
+* bulk_purchase
+* primary_reason
+* source
+
+We will now investigate random slopes in addition to random intercepts. We are having difficulties fitting a random slope for `dose_per_mg`, but we can check the random slope for bulk purchase:
+
+```{r}
+random_bulk <- anova(lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F),
+      lmer(data=df, log_ppm ~ (bulk_purchase |state) + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F))
+random_bulk$`Pr(>Chisq)`[2]  # Random slopes for bulk_purchase not helpful
+
+random_reason <- anova(lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F),
+                       lmer(data=df, log_ppm ~ (Primary_Reason |state) + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F))
+random_reason$`Pr(>Chisq)`[2]
+
+random_source <- anova(lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F),
+                       lmer(data=df, log_ppm ~ (source |state) + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F))
+random_source$`Pr(>Chisq)`[2]
+
+print(paste('For testing between competing models with and without a random slope on '))
+```
+
+These results suggest that we should stick with a random intercept only.
+
 #### Model Specification
@@
 
-This linear model estimates the log price per miligram $log(y)$ of oxymorphonefor an invidual $i$ in state $j$ ($x_{ij}$). The following predictors are defined as:
+This linear model estimates the log price per miligram $log(y)$ of oxymorphone for an invidual $i$ in state $j$ ($x_{ij}$). The following predictors are defined as:
 
@@
 ```{r}
-mod1b.inf <- influence(mod1b, maxfun=1)
-
-infIndexPlot(mod1b.inf, vars='Cook')
-
-# Find the most influential observations
-high_inf_idx <- order(cooks.distance(mod1b.inf), decreasing = TRUE)
-
-# Take a look at some of the most influential observations
-df[high_inf_idx[1:25], ]
+# mod1b.inf <- influence(mod1b, maxfun=1)
+# 
+# infIndexPlot(mod1b.inf, vars='Cook')
+# 
+# # Find the most influential observations
+# high_inf_idx <- order(cooks.distance(mod1b.inf), decreasing = TRUE)
+# 
+# # Take a look at some of the most influential observations
+# df[high_inf_idx[1:25], ]
 ```
@@
 #mod1b <- lmer(data=df, log_ppm ~ (1 | state) + year + mgstr + bulk_purchase + Primary_Reason + source)
-better_data <- df %>% filter(ppm > 0.2)
-mod1c <- lmer(data=better_data, log_ppm ~ (1 | state) + date_num + year + mgstr + bulk_purchase + Primary_Reason + source)
-qqmath(mod1c)
+# better_data <- df %>% filter(ppm > 0.2)
+# mod1c <- lmer(data=better_data, log_ppm ~ (1 | state) + date_num + year + mgstr + bulk_purchase + Primary_Reason + source)
+# qqmath(mod1c)
+```
+
+```{r}
+mod1 <- lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase + source, REML=F)
 ```

### 2021-03-21 Draft Report 3.21.Rmd
new version

@@
 #best model!!!
-mod_final <- lmer(data=df_filter, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase + source, REML=F)
+mod_final <- lmer(data=df_filter, log_ppm ~ (1 | state) + dose_per_mg + bulk_purchase + source + Primary_Reason, REML=F)
 BIC(mod_final)
@@
 
+## Start here
+
 ## Results
@@
 #confint(mod1)[3:11,]
-
 # extract coefficients
@@
 
-#### need to change numbers after finalizing model
-* dose_per_mg: Controlling for other variables, for every one milligram increase of dosage, the price per mg of oxymorphone increases by a multiplicative effect of exp(-0.01524395) = 0.9848717, about a 1.51% reduction.
+* dose_per_mg: Controlling for other variables, for every one milligram increase of dosage, the price per mg of oxymorphone increases by a multiplicative effect of exp(-0.0151868) = 0.9849279, about a 1.51% reduction.
 
 * bulk_purchase: “Not bulk purchase” is the reference group.
-  * Controlling for other variables, bulk purchase reduces the price per mg of oxymorphone by a multiplicative effect of exp(-0.17013501) = 0.8435509, about a 15.64% reduction.
+  * Controlling for other variables, bulk purchase reduces the price per mg of oxymorphone by a multiplicative effect of exp(-0.1691121) = 0.8444142, about a 15.56% reduction.
 
 * source: “Source: Unknown” is the reference group. 
-  * Controlling for other variables, purchasing oxymorphone with Internet source tends to increase the price per mg by a multiplicative effect of exp(0.12686001) = 1.135258, about a 13.53% increase than the price per mg of oxymorphone for people with unknown drug source.
-  * Controlling for other variables, purchasing oxymorphone with Internet Pharmacy source tends to reduce the price per mg by a multiplicative effect of exp(-0.44349281) = 0.6417908, about a 35.82% reduction of the price per mg of oxymorphone for people with unknown drug source.
-  * Controlling for other variables, purchasing oxymorphone with personal source tends to increase the price per mg by a multiplicative effect of exp(0.06140303) = 1.063327, about a 6.33% increase than the price per mg of oxymorphone for people with unknown drug source.
-  * Controlling for other variables, purchasing oxymorphone with “heard it” source tends to increase the price per mg by a multiplicative effect of exp(0.18025958) = 1.197528, about a 19.75% increase than the price per mg of oxymorphone for people with unknown drug source.
+  * Controlling for other variables, purchasing oxymorphone with Internet source tends to increase the price per mg by a multiplicative effect of exp(0.1339468) = 1.143332, about a 14.33% increase than the price per mg of oxymorphone for people with unknown drug source.
+  * Controlling for other variables, purchasing oxymorphone with Internet Pharmacy source tends to reduce the price per mg by a multiplicative effect of exp(-0.4584378) = 0.6322706, about a 36.77% reduction than the price per mg of oxymorphone for people with unknown drug source.
+  * Controlling for other variables, purchasing oxymorphone with personal source tends to increase the price per mg by a multiplicative effect of exp(0.0565153) = 1.058143, about a 5.81% increase than the price per mg of oxymorphone for people with unknown drug source.
+  * Controlling for other variables, purchasing oxymorphone with “heard it” source tends to increase the price per mg by a multiplicative effect of exp(0.1790027) = 1.196024, about a 19.6% increase than the price per mg of oxymorphone for people with unknown drug source.
+
+* Primary Reason: “Primary_Reason: Blank” is the reference group. 
+  * Controlling for other variables, purchasing oxymorphone with primary reason Left unanswered tends to increase the price per mg by a multiplicative effect of exp(0.0176846) = 1.017842, about a 1.78% increase than the price per mg of oxymorphone for people with Blank primary reason.
+  * Controlling for other variables, purchasing oxymorphone with primary reason Other or unknown tends to increase the price per mg by a multiplicative effect of exp(0.0486307) = 1.049833, about a 4.98% increase than the price per mg of oxymorphone for people with Blank primary reason.
+  * Controlling for other variables, purchasing oxymorphone with primary reason Self-treat tends to increase the price per mg by a multiplicative effect of exp(0.1034974) = 1.109043, about a 10.9% increase than the price per mg of oxymorphone for people with Blank primary reason.
+  * Controlling for other variables, purchasing oxymorphone with primary reason Enjoyment tends to increase the price per mg by a multiplicative effect of exp(0.1838890) = 1.201882, about a 20.19% increase than the price per mg of oxymorphone for people with Blank primary reason.
+  * Controlling for other variables, purchasing oxymorphone with primary reason Resell tends to reduce the price per mg by a multiplicative effect of exp(-0.3179197) = 0.7276612, about a 27.23% reduction than the price per mg of oxymorphone for people with Blank primary reason.
+  * Controlling for other variables, purchasing oxymorphone with primary reason Refuse to answer tends to increase the price per mg by a multiplicative effect of exp(0.0564516) = 1.058075, about a 5.81% increase than the price per mg of oxymorphone for people with Blank primary reason.
 
@@
 ```{r}
-#VarCorr(mod1)
-Variance <- c(0.0233, 0.5028)
-Std.Dev <- c(0.1526, 0.7091)
+#VarCorr(mod_final)
+Variance <- c(0.02292, 0.50060)
+Std.Dev <- c(0.15140 , 0.70753)
 Groups <- c("state (Intercept)", "Residual")
-
 REvar <- cbind(Groups, Variance, Std.Dev)
-
 knitr::kable(REvar, col.names = c("Group", "Variance", "Std.Dev"))
 ```
-The within-state variance $\sigma^2$ is 0.5028, which is much larger than the across-state variance $\tau^2$, 0.0233.
+The within-state variance $\sigma^2$ is 0.5006, which is much larger than the across-state variance $\tau^2$, 0.02292.
 
@@
 ```{r}
-ICC <- 0.047
+ICC <- 0.044
 knitr::kable(ICC, col.names = "Intraclass Correlation")
 ```
-Intraclass correlation is 0.047, which means values are not very similar from the same state.
+Intraclass correlation is 0.044, which means values are not very similar from the same state.
 

### 2021-03-21 STA_610_Case_Study_1_Report_Team_4.Rmd
Altered model specification

@@
 In order to test whether keeping a variable in the model is "worth it", we may perform a likelihood ratio test for nested models.
-For adding $k$ fixed effects, the LRT test statistic follows a $\chi^2(k)$ distribution, and for adding $k$ random effect, the LRT test statistic follows a 50-50 mixture of $\chi^2(k)$ and $\chi^2(k + 1)$ distribution.
+For adding $k$ fixed effects, the LRT test statistic follows a $\chi^2(k)$ distribution, and for adding $k$ random effects, the LRT test statistic follows a 50-50 mixture of $\chi^2(k)$ and $\chi^2(k + 1)$ distribution.
 For non-nested models, we can use BIC as a rough indicator of which model is "better".
@@
 
-$$y_{ij} = \beta_{0,j} + \beta_{1} date_num + \beta_{2} mgstr + \beta_{3} bulk_purchase + \beta_{4} Primary_Reason + \beta_{5} source + \epsilon_{ij}$$
-
-$$log(y_{ij}) = \beta_{0,j} + \beta_{1} x_{ij1} + \beta_{2} x_{ij2} + \beta_{3} x_{ij3} + \beta_{4} x_{ij4} + \beta_{5} x_{ij5} + \epsilon_{ij}$$
+$$log(y_{ij}) = \beta_{0,j} + \beta_{1} x_{ij1} + \beta_{2} x_{ij2} + \beta_{3} x_{ij3} + \beta_{4} x_{ij4} + \epsilon_{ij}$$
 
@@
 
-* $x_1$: The number of days since the recorded transaction and January 1st, 2000
-* $x_2$: Dosage strength in mg of the units purchased
-* $x_3$: Bulk purchase, an indicator for whether 10+ units were purchased at once
-* $x_4$: The primary reason for purchase. Factors include "other/unknown", "self-treatment", and "enjoyment"
-* $x_5$: Source of information (allows for reporting of transactions). Factors include "unkown", "internet", "heard it", and "personal"
-
-Model Selection discussions here.
-
-- mgstr: good
-- bulk_purchase: good
+* $x_1$: Dosage strength in mg of the units purchased
+* $x_2$: Bulk purchase, an indicator for whether 10+ units were purchased at once
+* $x_3$: The primary reason for purchase. Factors include "blank", "left unanswered", "other or unknown", "self-treatment", "enjoyment", "resell", and "refuse to answer"
+* $x_4$: Source of information (allows for reporting of transactions). Factors include "unkown", "internet", "internet pharmacy", "heard it", and "personal"
 

### 2021-03-21 final_report.Rmd
Added Lauren's EDA

--- 2021-03-20 report_draft_fz_lp_4.Rmd
@@
 ```{r, echo = FALSE}
-
 load("~/Documents/Spring 2021/STA 610 - H Models/cs1/streetrx.RData")
@@
 # cleaning state:
-
 # table(df$state)
-
 # remove levels corresponding to states with 0 observations in df
@@
 # table(df$state)
-
 # drop North Dakota - one observation
@@
 # table(df$state)
-
 # replace state "USA" with "Unknown"
@@
 # table(df$state)
-
 ### PUT IN APPENDIX: the state-by-region table of coutns below
@@
 # Change date to numeric (easier for interpretation)
-
 # this sets 
 df$date_num <- as.numeric(as.Date(df$price_date, "%m/%d/%y")) - 14621 # sets 1/12/2010 (first post-2010 date) as zero
-
 tempdf <- df %>% select(date_num, price_date)
-
 # this filters out all data before the 1/1/2000, which probably shouldn't be included
@@
 # recode sources 
-
 # remove unused levels
 df$source <- droplevels(df$source)
-
 source_df <- data.frame(table(df$source))
-
 # levels(df$source)
 levels(df$source)[1] <- "Blank"
-
 source_df <- data.frame(table(df$source))
-
 df$source <- recode_factor(df$source, "Blank" = "Blank",
@@
                            .default = "Internet") # everything else is a URL
-
 source_df_2 <- data.frame(table(df$source))
-
 # drop unknowns
-
 df <- df %>% filter(df$source != "Unknown")
 df$source <- droplevels(df$source)
-
 knitr::kable(table(df$source), col.names = c("Level", "Freq"))
@@
 df$Primary_Reason <- droplevels(df$Primary_Reason)
-
 reason_df <- data.frame(table(df$Primary_Reason))
-
 # levels(df$Primary_Reason)
 levels(df$Primary_Reason)[1] <- "Blank"
-
 reason_df <- data.frame(table(df$Primary_Reason))
-
 df$Primary_Reason <- recode_factor(df$Primary_Reason, "Blank" = "Blank",
@@
                            .default = "Other or unknown")
-
 reason_df_2 <- data.frame(table(df$Primary_Reason))
-
 #print(reason_df)
@@
 df$log_ppm <- log(df$ppm)
-
 hist_ppm <- ggplot(df, aes(x=ppm)) + 
@@
  geom_density(alpha=.2, fill="red") 
-
 hist_log_ppm <- ggplot(df, aes(x=log_ppm)) + 
@@
  geom_density(alpha=.2, fill="blue") 
-
 grid.arrange(hist_ppm, hist_log_ppm, ncol=2)
@@
 labs(title="log PPM by city (FL)", x="Log PPM",y="city")
-
 ca_lppmbycity <- ggplot(df[df$state=="California",], aes(x=reorder(city, city, length), y=log_ppm)) +
@@
 labs(title="log PPM by city (CA)", x="Log PPM",y="city")
-
 grid.arrange(fl_lppmbycity, ca_lppmbycity, ncol= 2)
@@
 group_slope_list <- NULL
-
 for(statej in
@@
 m <- lm(data = df[df$state==statej, ], log_ppm ~ dose_per_mg)
-
 group_int_list <- c(group_int_list, coef(m)[1])
@@
 }
-
 hist(group_int_list)
@@
 m <- lm(data = df[df$state==statej, ], log_ppm ~ date_num)
-
 group_int_list <- c(group_int_list, coef(m)[1])
@@
 ```
-
-# Model selection and diagnostics ---- Unfinished! ------
-
-## Model specification
-
-We are considering models of the following form (with some coefficients potentially equal to zero):
-
-$$
-y_{ij} = \beta_{0,j} + \beta_{1,j} date\text{_}num + \beta_{2,j} mgstr + \beta_{3,j} bulk\text{_}purchase + \beta_{4,j} Primary\text{_}Reason + \beta_{5,j} source + \beta_{6,j} date\text{_}quarter + \epsilon_{ij}
-$$
-
-Where: 
-* $\beta_{0,j}$ represents the intercept, which we plan to fit as a random intercept for each group $j$, so that it would have the form:
-$\beta_{0,j} = \beta_0 + b_{0,j}, b_{0,j} \sim N(0,\tau_0^2)$
-
-* The $\beta_k$ for $k=1,...,6$ are the coefficients for our fixed effect predictors. We will consider random slopes by group for these, in which case they would have the form: 
-$\beta_{k,j} = \beta_k + b_{k,j}, b_{k,j} \sim N(0,\tau_k^2)$
-
-We also have that $\epsilon_{ij} \perp b_{0,j} \overset{iid}{\sim} N(0,\sigma^2)$ for all $i, j$.
-
-## Model specification
-
-### Grouping variables and random intercepts
-
-We see that `state` appears to be best for grouping; region is good as well.
-
-```{r}
-mod_s <- lmer(data=df, log_ppm ~ (1 |state) + date_num + date_quarter + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F)
-mod_c <- lmer(data=df, log_ppm ~ (1 |city) + date_num + date_quarter + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F)
-mod_r <- lmer(data=df, log_ppm ~ (1 |USA_region) + date_num + date_quarter + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F)
-
-summary(mod_s)
-# summary(mod_c)
-# summary(mod_r)
-
-LR <- 2*(logLik(mod_s)-logLik(mod_c))
-0.5*(1-pchisq(LR[1],1)+1-pchisq(LR[1],1)) # choose s over c
-
-LR <- 2*(logLik(mod_r)-logLik(mod_c))
-0.5*(1-pchisq(LR[1],1)+1-pchisq(LR[1],1)) # choose c over r
-
-LR <- 2*(logLik(mod_s)-logLik(mod_r))
-0.5*(1-pchisq(LR[1],1)+1-pchisq(LR[1],1)) # choose s over r
-
-# conclusion: choose state random intercept
-```
-
-```{r}
-anova(mod_s,
-      lmer(data=df, log_ppm ~ (1 |state) + date_quarter + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F))
-# drop date_num
-
-anova(lmer(data=df, log_ppm ~ (1 |state) + date_quarter + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F),
-      lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F))
-# drop date_quarter
-
-anova(lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F),
-      lmer(data=df, log_ppm ~ (1 |state) + bulk_purchase + Primary_Reason + source, REML=F))
-# don't drop dose_per_mg
-
-anova(lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F),
-      lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + Primary_Reason + source, REML=F))
-# don't srop bulk_purchase
-
-anova(lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F),
-      lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase + source, REML=F))
-# do drop Primary_reason
-
-anova(lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase + source, REML=F),
-      lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase, REML=F))
-# do drop source
-```
-
-Now investigate random slopes:
-
-```{r}
-anova(lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase, REML=F),
-      lmer(data=df, log_ppm ~ (dose_per_mg |state) + dose_per_mg + bulk_purchase, REML=F))
-# random slopes for dose_per_mg are good - BIC = 8624.6
-
-anova(lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase, REML=F),
-      lmer(data=df, log_ppm ~ (bulk_purchase |state) + dose_per_mg + bulk_purchase, REML=F))
-# random slopes for nulk_purchase not helpful
-```
-
-Our candidate for final model is then:
-
-```{r}
-mod_final <- lmer(data=df, log_ppm ~ (dose_per_mg |state) + dose_per_mg + bulk_purchase, REML=F)
-```
-
-There are convergence inssues. Try removing states that are influential? Based on the "dfbetas" visualization below, they are Wyoming, Nevada, Alaska.
-
-```{r}
-mod_infl <- influence(mod_final,"state")
-print(2/sqrt(length(unique(df_filter$state))))
-diagn_df <- data.frame(dfbetas(mod_infl))
-plot(mod_infl,which="dfbetas",xlab="DFBETAS",ylab="State")
-
-resqq <- plot_resqq(mod_final)
-ranefci <- plot_ranef(mod_final)
-grid.arrange(resqq, ranefci, ncol = 2)
-```
-
-```{r}
-df_filter <- df %>% filter(state != "Alaska" & state != "Wyoming" & state != "Nevada")
-```
-
-```{r}
-mod_final <- lmer(data=df_filter, log_ppm ~ (dose_per_mg |state) + dose_per_mg + bulk_purchase, REML=F)
-
-mod_infl <- influence(mod_final,"state")
-print(2/sqrt(length(unique(df_filter$state))))
-diagn_df <- data.frame(dfbetas(mod_infl))
-plot(mod_infl,which="dfbetas",xlab="DFBETAS",ylab="State")
-
-resqq <- plot_resqq(mod_final)
-ranefci <- plot_ranef(mod_final)
-grid.arrange(resqq, ranefci, ncol = 2)
-```
-
-See below: looks like the singularity is because some states don't have enough data for `bulk_purchase`. Ugh. Should we use `region` instead???
-
-```{r}
-View(data.frame(table(df$state, df$bulk_purchase)))
-```

### 2021-03-21 final_report.Rmd
Added my section for model selection and specification

@@
 ```
+
+## Model
+
+Per the research question of interest, we are interested in exploring factors related to price per mg of oxymorphone.
+We are also specifically asked to control for location and how drug price changes with it.
+Cities seemed to be too noisy to use as a location predictor.
+On the flip side, region did not seem to be associated with a significant change in drug prices.
+Therefore, state provides the best granularity of accounting for location in this instance.
+
+Linear regression models are usually a good standard approach to this, providing clear interpretations of how various predictors are associated with drug price.
+As seen in the EDA, however, many states have excessively small sample sizes.
+This produces corresponding $\hat{\beta}_{OLS}$ estimates with large variance.
+We would like to be able to *partially pool* information across states, letting us borrow information to help stabilize state estimates with small sample sizes.
+
+This prompts a **mixed effects model**, where we instead model the state coefficients as having been drawn $N(0, \tau^2)$.
+
+### Model Selection
+
+Based off of the results of the EDA, we want to initially fit a model with the following variables:
+
+- source
+- dose_per_mg
+- bulk_purchase
+- Primary_Reason
+- date_num
+- date_quarter
+
+In terms of the research question, we would also like to group based on location. There are 3 options for this:
+
+- Region
+- State
+- City
+
+In order to test whether keeping a variable in the model is "worth it", we may perform a likelihood ratio test for nested models.
+For adding $k$ fixed effects, the LRT test statistic follows a $\chi^2(k)$ distribution, and for adding $k$ random effects, the LRT test statistic follows a 50-50 mixture of $\chi^2(k)$ and $\chi^2(k + 1)$ distribution.
+For non-nested models, we can use BIC as a rough indicator of which model is "better".
+BIC balances likelihood of model with how parsimonious it is.
+
+First, let's decide what to group on for region:
+
+```{r}
+# Fit 3 models grouping on different location specificity
+mod_s <- lmer(data=df, log_ppm ~ (1 |state) + date_num + date_quarter + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F)
+mod_c <- lmer(data=df, log_ppm ~ (1 |city) + date_num + date_quarter + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F)
+mod_r <- lmer(data=df, log_ppm ~ (1 |USA_region) + date_num + date_quarter + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F)
+```
+
+```{r}
+BIC <- sapply(c(mod_s, mod_c, mod_r), BIC)
+re_results <- data.frame('Grouping' = c('State', 'City', 'Region'), 'BIC' = BIC)
+kable(re_results, caption = 'BIC for Models with Different Random Intercepts')
+```
+
+State provides the best level of BIC, so we will choose this for our grouping variable.
+
+Now let's investigate dropping various fixed effects and see results of our hypothesis tests:
+
+```{r}
+# Test date_num
+drop_date_num <- anova(mod_s,
+      lmer(data=df, log_ppm ~ (1 |state) + date_quarter + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F))
+date_num_pval <- drop_date_num$`Pr(>Chisq)`[2]
+# Test date_quarter
+drop_quarter <- anova(lmer(data=df, log_ppm ~ (1 |state) + date_quarter + dose_per_mg + 
+                    bulk_purchase + Primary_Reason + source, REML=F),
+      lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F))
+quarter_pval <- drop_quarter$`Pr(>Chisq)`[2]
+# Test dose_per_mg
+drop_dose <- anova(lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F),
+      lmer(data=df, log_ppm ~ (1 |state) + bulk_purchase + Primary_Reason + source, REML=F))
+dose_pval <- drop_dose$`Pr(>Chisq)`[2]
+# Test bulk purchase
+drop_bulk <- anova(lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F),
+      lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + Primary_Reason + source, REML=F))
+bulk_pval <- drop_bulk$`Pr(>Chisq)`[2]
+# Test primary reason
+drop_reason <- anova(lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F),
+      lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase + source, REML=F))
+reason_pval <- drop_reason$`Pr(>Chisq)`[2]
+# Test source
+drop_source <- anova(lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F),
+      lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase + Primary_Reason, REML=F))
+source_pval <- drop_source$`Pr(>Chisq)`[2]
+```
+
+The outline is as follows:
+
+* Start with date_num, date_quarter, dose_per_mg, bulk_purchase, Primary_Reason, source
+* Perform a LRT between competing models with and without the date variable.
+Observe a p-value of `r round(date_num_pval, 3)` and fail to reject the null (simpler model).
+Drop `date_num` from the subsequent model tests
+* Perform a LRT between competing models with and without the quarter of the sale.
+Observe a p-value of `r round(quarter_pval, 3)` and fail to reject the null (simpler model).
+Drop `date_quarter` from the subsequent model tests
+* Perform a LRT between competing models with and without dose per mg. Observe a p-value of ~`r round(dose_pval, 3)` and reject the null (simpler model). Decide on model keeping `dose_per_mg`
+* Perform a LRT between competing models with and without the bulk purchase predictor.
+Observe a p-value of ~`r round(bulk_pval, 3)` and reject the null (simpler model).
+Decide on model keeping `bulk_purchase`
+* Perform a LRT between competing models with and without the primary reason predictor.
+Observe a p-value of `r round(reason_pval, 3)` and reject the null (simpler model).
+Decide on keeping `primary_reason` in the model
+* Perform a LRT between competing models with and without the source predictor.
+Observe a p-value of ~`r round(source_pval, 3)` and reject the null (simpler model).
+Decide on keeping `source` in the model
+
+Thus these results suggest we want to drop `date_num` and `date_quarter`, while keeping
+
+* dose_per_mg
+* bulk_purchase
+* primary_reason
+* source
+
+We will now investigate random slopes in addition to random intercepts. We are having difficulties fitting a random slope for `dose_per_mg`, but we can check the random slope for bulk purchase:
+
+```{r}
+random_bulk <- anova(lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F),
+      lmer(data=df, log_ppm ~ (bulk_purchase |state) + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F))
+random_bulk$`Pr(>Chisq)`[2]  # Random slopes for bulk_purchase not helpful
+random_reason <- anova(lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F),
+                       lmer(data=df, log_ppm ~ (Primary_Reason |state) + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F))
+random_reason$`Pr(>Chisq)`[2]
+random_source <- anova(lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F),
+                       lmer(data=df, log_ppm ~ (source |state) + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F))
+random_source$`Pr(>Chisq)`[2]
+print(paste('For testing between competing models with and without a random slope on '))
+```
+
+These results suggest that we should stick with a random intercept only.
+
+#### Model Specification
+
+$$log(y_{ij}) = \beta_{0,j} + \beta_{1} x_{ij1} + \beta_{2} x_{ij2} + \beta_{3} x_{ij3} + \beta_{4} x_{ij4} + \epsilon_{ij}$$
+
+$$
+\beta_{0,j} = \beta_0 + b_{0,j}, b_{0,j} \sim N(0,\tau^2)
+$$
+$$
+\epsilon_{ij} \perp b_{0,j} \overset{iid}{\sim} N(0,\sigma^2)
+$$
+
+This linear model estimates the log price per miligram $log(y)$ of oxymorphone for an invidual $i$ in state $j$ ($x_{ij}$). The following predictors are defined as:
+
+* $x_1$: Dosage strength in mg of the units purchased
+* $x_2$: Bulk purchase, an indicator for whether 10+ units were purchased at once
+* $x_3$: The primary reason for purchase. Factors include "blank", "left unanswered", "other or unknown", "self-treatment", "enjoyment", "resell", and "refuse to answer"
+* $x_4$: Source of information (allows for reporting of transactions). Factors include "unkown", "internet", "internet pharmacy", "heard it", and "personal"
+
+```{r}
+# mod1b.inf <- influence(mod1b, maxfun=1)
+# 
+# infIndexPlot(mod1b.inf, vars='Cook')
+# 
+# # Find the most influential observations
+# high_inf_idx <- order(cooks.distance(mod1b.inf), decreasing = TRUE)
+# 
+# # Take a look at some of the most influential observations
+# df[high_inf_idx[1:25], ]
+```
+
+```{r}
+#mod1 <- lmer(data=df, log_ppm ~ (1 | state) + date_num + mgstr + bulk_purchase + Primary_Reason + source)
+#mod1b <- lmer(data=df, log_ppm ~ (1 | state) + year + mgstr + bulk_purchase + Primary_Reason + source)
+# better_data <- df %>% filter(ppm > 0.2)
+# mod1c <- lmer(data=better_data, log_ppm ~ (1 | state) + date_num + year + mgstr + bulk_purchase + Primary_Reason + source)
+# qqmath(mod1c)
+```
+
+```{r}
+mod1 <- lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase + source, REML=F)
+```
+
+#### Model Diagnostics
+```{r, fig.align='center', warning=FALSE, message=FALSE, error=FALSE}
+# resqq <- plot_resqq(mod1)
+# ranefci <- plot_ranef(mod1)
+# grid.arrange(resqq, ranefci, ncol = 2)
+```
+
+Model diagnostics discussions here.

### 2021-03-21 final_report.Rmd
Added Fan/Lingyu's section on results

@@
 Model diagnostics discussions here.
+
+## Results
+#### Fixed Effects
+```{r, message=FALSE}
+#knitr::kable(fixef(mod1))
+#confint(mod1)[3:11,]
+# extract coefficients
+coefs <- data.frame(coef(summary(mod_final)))
+# use normal distribution to approximate p-value
+coefs$p.z <- 2 * (1 - pnorm(abs(coefs$t.value)))
+knitr::kable(coefs, col.names = c("Estimate", "Std.Error", "t-value", "P-value"))
+```
+
+* dose_per_mg: Controlling for other variables, for every one milligram increase of dosage, the price per mg of oxymorphone increases by a multiplicative effect of exp(-0.0151868) = 0.9849279, about a 1.51% reduction.
+
+* bulk_purchase: “Not bulk purchase” is the reference group.
+  * Controlling for other variables, bulk purchase reduces the price per mg of oxymorphone by a multiplicative effect of exp(-0.1691121) = 0.8444142, about a 15.56% reduction.
+
+* source: “Source: Unknown” is the reference group. 
+  * Controlling for other variables, purchasing oxymorphone with Internet source tends to increase the price per mg by a multiplicative effect of exp(0.1339468) = 1.143332, about a 14.33% increase than the price per mg of oxymorphone for people with unknown drug source.
+  * Controlling for other variables, purchasing oxymorphone with Internet Pharmacy source tends to reduce the price per mg by a multiplicative effect of exp(-0.4584378) = 0.6322706, about a 36.77% reduction than the price per mg of oxymorphone for people with unknown drug source.
+  * Controlling for other variables, purchasing oxymorphone with personal source tends to increase the price per mg by a multiplicative effect of exp(0.0565153) = 1.058143, about a 5.81% increase than the price per mg of oxymorphone for people with unknown drug source.
+  * Controlling for other variables, purchasing oxymorphone with “heard it” source tends to increase the price per mg by a multiplicative effect of exp(0.1790027) = 1.196024, about a 19.6% increase than the price per mg of oxymorphone for people with unknown drug source.
+
+* Primary Reason: “Primary_Reason: Blank” is the reference group. 
+  * Controlling for other variables, purchasing oxymorphone with primary reason Left unanswered tends to increase the price per mg by a multiplicative effect of exp(0.0176846) = 1.017842, about a 1.78% increase than the price per mg of oxymorphone for people with Blank primary reason.
+  * Controlling for other variables, purchasing oxymorphone with primary reason Other or unknown tends to increase the price per mg by a multiplicative effect of exp(0.0486307) = 1.049833, about a 4.98% increase than the price per mg of oxymorphone for people with Blank primary reason.
+  * Controlling for other variables, purchasing oxymorphone with primary reason Self-treat tends to increase the price per mg by a multiplicative effect of exp(0.1034974) = 1.109043, about a 10.9% increase than the price per mg of oxymorphone for people with Blank primary reason.
+  * Controlling for other variables, purchasing oxymorphone with primary reason Enjoyment tends to increase the price per mg by a multiplicative effect of exp(0.1838890) = 1.201882, about a 20.19% increase than the price per mg of oxymorphone for people with Blank primary reason.
+  * Controlling for other variables, purchasing oxymorphone with primary reason Resell tends to reduce the price per mg by a multiplicative effect of exp(-0.3179197) = 0.7276612, about a 27.23% reduction than the price per mg of oxymorphone for people with Blank primary reason.
+  * Controlling for other variables, purchasing oxymorphone with primary reason Refuse to answer tends to increase the price per mg by a multiplicative effect of exp(0.0564516) = 1.058075, about a 5.81% increase than the price per mg of oxymorphone for people with Blank primary reason.
+
+#### Random Effects
+```{r}
+#VarCorr(mod_final)
+Variance <- c(0.02292, 0.50060)
+Std.Dev <- c(0.15140 , 0.70753)
+Groups <- c("state (Intercept)", "Residual")
+REvar <- cbind(Groups, Variance, Std.Dev)
+knitr::kable(REvar, col.names = c("Group", "Variance", "Std.Dev"))
+```
+The within-state variance $\sigma^2$ is 0.5006, which is much larger than the across-state variance $\tau^2$, 0.02292.
+
+```{r, include=FALSE}
+library(sjstats)
+icc(mod_final)
+```
+
+```{r}
+ICC <- 0.044
+knitr::kable(ICC, col.names = "Intraclass Correlation")
+```
+Intraclass correlation is 0.044, which means values are not very similar from the same state.
+
+```{r, fig.align='center', fig.height=8}
+dotplot(ranef(mod_final, condVar=TRUE))$state
+```
+From the result of the random effects, we can see that the random intercept for Tennessee is the highest and it is significantly different from zero.
+The random intercept for Arizona is the lowest and this intercept is also significant.
+
+## Appendix
+```{r, ref.label=knitr::all_labels(),echo=TRUE,eval=FALSE}
+```

### 2021-03-22 final_report.Rmd
Aesthetic changes, removal of unnecesary code, and general imporvements

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE)
+knitr::opts_chunk$set(echo = FALSE, cache=TRUE, fig.height = 3)
 ```
@@
 ```{r, echo=FALSE, warning=FALSE, message=FALSE}
-library(tidyverse)
-library(lme4)
-library(gridExtra)
-library(grid)
-library(ggplot2)
-library(lattice)
+library(pacman)
+pacman::p_load(tidyverse, lme4, gridExtra, grid, ggplot2, lattice, redres, stringr, influence.ME, knitr)
 # devtools::install_github("goodekat/redres")
-library(redres)
-library(stringr)
-library(influence.ME)
 ```
@@
 ```{r, echo = FALSE}
-load("~/Documents/Spring 2021/STA 610 - H Models/cs1/streetrx.RData")
-df <- streetrx %>% filter(api_temp=="oxymorphone", !is.na(ppm))
+load("streetrx.rdata")
+df <- streetrx %>% filter(api_temp == "oxymorphone", !is.na(ppm))
 ```
@@
 The benefits of a hierarchical model include the ability to "borrow information" from the data set as a whole to stabilize estimates of price for groups (e.g., states) with relatively small sample size, and to establish a estimated distribution of how price varies with respect to groups, which would in theory allow generalization to new groups not included in the original analysis.
-
 # Exploratory data analysis and data cleaning
@@
 # categorical variables -- looking at South only
-ggplot(data = df[df$USA_region == "South",], aes(x = bulk_purchase, y = log_ppm)) + 
+set.seed(42)
+states <- sample(df$state, 5)
+ggplot(data = df %>% filter(state %in% states), aes(x = bulk_purchase, y = log_ppm)) + 
   geom_boxplot() +
-  facet_grid(~ state)
+  facet_grid(~ state) +
+  theme(axis.text.x = element_text(angle = 60))
 ```
@@
   geom_boxplot() +
-  facet_grid(~ state)
+  facet_grid(~ state) +
+  theme(axis.text.x = element_text(angle = 90))
 ```
@@
   geom_boxplot() +
-  facet_grid(~ state)
+  facet_grid(~ state) + 
+  theme(axis.text.x = element_text(angle = 90))
 ```
+
 ## More EDA: slope and intercept distributions for predictors by state
@@
 group_slope_list <- NULL
-for(statej in
-  unique(df$state))
-{
+for(statej in unique(df$state)) {
     m <- lm(data = df[df$state==statej, ], log_ppm ~ dose_per_mg)
@@
 group_slope_list <- NULL
-for(statej in
-  unique(df$state))
-{
+for(statej in unique(df$state)) {
     m <- lm(data = df[df$state==statej, ], log_ppm ~ date_num)
@@
 }
+
 hist(group_int_list)
@@
 ```{r}
-# mod1b.inf <- influence(mod1b, maxfun=1)
-# 
-# infIndexPlot(mod1b.inf, vars='Cook')
-# 
-# # Find the most influential observations
-# high_inf_idx <- order(cooks.distance(mod1b.inf), decreasing = TRUE)
-# 
-# # Take a look at some of the most influential observations
-# df[high_inf_idx[1:25], ]
-```
-
-```{r}
-#mod1 <- lmer(data=df, log_ppm ~ (1 | state) + date_num + mgstr + bulk_purchase + Primary_Reason + source)
-#mod1b <- lmer(data=df, log_ppm ~ (1 | state) + year + mgstr + bulk_purchase + Primary_Reason + source)
-# better_data <- df %>% filter(ppm > 0.2)
-# mod1c <- lmer(data=better_data, log_ppm ~ (1 | state) + date_num + year + mgstr + bulk_purchase + Primary_Reason + source)
-# qqmath(mod1c)
-```
-
-```{r}
-mod1 <- lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase + source, REML=F)
+mod_final <- lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F)
 ```
@@
 ## Appendix
-```{r, ref.label=knitr::all_labels(),echo=TRUE,eval=FALSE}
+```{r, ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE}
 ```

### 2021-03-22 final_report.Rmd
Various edits to save space

@@
 author:
-- Cole Juracek (Coordinator)
-- Lauren Palazzo (Programmer)
-- Lingyu Zhou (Checker)
+- Cole Juracek (Programmer)
+- Lauren Palazzo (Writer)
+- Lingyu Zhou (Checker + Coordinator)
 - Fan Zhu (Presenter)
@@
   facet_grid(~ state) +
-  theme(axis.text.x = element_text(angle = 60))
+  theme(axis.text.x = element_text(angle = 90))
 ```
@@
 
-```{r}
+```{r fig.width=3}
 group_int_list <- NULL 
@@
 
-We will now investigate random slopes in addition to random intercepts. We are having difficulties fitting a random slope for `dose_per_mg`, but we can check the random slope for bulk purchase:
+#### Model Specification
 
-```{r}
-random_bulk <- anova(lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F),
-      lmer(data=df, log_ppm ~ (bulk_purchase |state) + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F))
-random_bulk$`Pr(>Chisq)`[2]  # Random slopes for bulk_purchase not helpful
-random_reason <- anova(lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F),
-                       lmer(data=df, log_ppm ~ (Primary_Reason |state) + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F))
-random_reason$`Pr(>Chisq)`[2]
-random_source <- anova(lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F),
-                       lmer(data=df, log_ppm ~ (source |state) + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F))
-random_source$`Pr(>Chisq)`[2]
-print(paste('For testing between competing models with and without a random slope on '))
-```
-
-These results suggest that we should stick with a random intercept only.
-
-#### Model Specification
+Our final model is as specified:
 

### 2021-03-22 final_report.Rmd

@@
 
-* dose_per_mg: Controlling for other variables, for every one milligram increase of dosage, the price per mg of oxymorphone increases by a multiplicative effect of exp(-0.0151868) = 0.9849279, about a 1.51% reduction.
+* *dose_per_mg*: Controlling for other variables, for every one milligram increase of dosage, the price per mg of oxymorphone increases by a multiplicative effect of exp(-0.0151868) = 0.9849279, about a 1.51% reduction.
 
-* bulk_purchase: “Not bulk purchase” is the reference group.
-  * Controlling for other variables, bulk purchase reduces the price per mg of oxymorphone by a multiplicative effect of exp(-0.1691121) = 0.8444142, about a 15.56% reduction.
+* *bulk_purchase* ("not bulk purchase" is the reference): Controlling for other variables, bulk purchase reduces the price per mg of oxymorphone by a multiplicative effect of exp(-0.1691121) = 0.8444142, about a 15.56% reduction.
 
-* source: “Source: Unknown” is the reference group. 
-  * Controlling for other variables, purchasing oxymorphone with Internet source tends to increase the price per mg by a multiplicative effect of exp(0.1339468) = 1.143332, about a 14.33% increase than the price per mg of oxymorphone for people with unknown drug source.
-  * Controlling for other variables, purchasing oxymorphone with Internet Pharmacy source tends to reduce the price per mg by a multiplicative effect of exp(-0.4584378) = 0.6322706, about a 36.77% reduction than the price per mg of oxymorphone for people with unknown drug source.
-  * Controlling for other variables, purchasing oxymorphone with personal source tends to increase the price per mg by a multiplicative effect of exp(0.0565153) = 1.058143, about a 5.81% increase than the price per mg of oxymorphone for people with unknown drug source.
-  * Controlling for other variables, purchasing oxymorphone with “heard it” source tends to increase the price per mg by a multiplicative effect of exp(0.1790027) = 1.196024, about a 19.6% increase than the price per mg of oxymorphone for people with unknown drug source.
+* *source* (“Source: Unknown” is the reference, other levels ommitted for brevity):
+    + Controlling for other variables, purchasing oxymorphone with Internet source tends to increase the price per mg by a multiplicative effect of exp(0.1339468) = 1.143332, about a 14.33% increase than the price per mg of oxymorphone for people with unknown drug source.
 
-* Primary Reason: “Primary_Reason: Blank” is the reference group. 
-  * Controlling for other variables, purchasing oxymorphone with primary reason Left unanswered tends to increase the price per mg by a multiplicative effect of exp(0.0176846) = 1.017842, about a 1.78% increase than the price per mg of oxymorphone for people with Blank primary reason.
-  * Controlling for other variables, purchasing oxymorphone with primary reason Other or unknown tends to increase the price per mg by a multiplicative effect of exp(0.0486307) = 1.049833, about a 4.98% increase than the price per mg of oxymorphone for people with Blank primary reason.
-  * Controlling for other variables, purchasing oxymorphone with primary reason Self-treat tends to increase the price per mg by a multiplicative effect of exp(0.1034974) = 1.109043, about a 10.9% increase than the price per mg of oxymorphone for people with Blank primary reason.
-  * Controlling for other variables, purchasing oxymorphone with primary reason Enjoyment tends to increase the price per mg by a multiplicative effect of exp(0.1838890) = 1.201882, about a 20.19% increase than the price per mg of oxymorphone for people with Blank primary reason.
-  * Controlling for other variables, purchasing oxymorphone with primary reason Resell tends to reduce the price per mg by a multiplicative effect of exp(-0.3179197) = 0.7276612, about a 27.23% reduction than the price per mg of oxymorphone for people with Blank primary reason.
-  * Controlling for other variables, purchasing oxymorphone with primary reason Refuse to answer tends to increase the price per mg by a multiplicative effect of exp(0.0564516) = 1.058075, about a 5.81% increase than the price per mg of oxymorphone for people with Blank primary reason.
+* *Primary Reason* ("Blank" is the reference, other levels omitted for brevity):
+    + Controlling for other variables, purchasing oxymorphone with primary reason Self-treat tends to increase the price per mg by a multiplicative effect of exp(0.1034974) = 1.109043, about a 10.9% increase than the price per mg of oxymorphone for people with Blank primary reason.
 

### 2021-03-22 final_report.Rmd
Moved some EDA to appendix

@@
 
-We find that there are many cities with a sample size of one, which makes `city` perhaps not the most natural choice for a grouping variable.
-Example visualizations shown here are for all cities in the states of Florida and California.
-
-```{r, fig.align='center'}
-fl_lppmbycity <- ggplot(df[df$state=="Florida",], aes(x=log_ppm, y=reorder(city, city, length))) +
-geom_boxplot() +
-labs(title="log PPM by city (FL)", x="Log PPM",y="city")
-ca_lppmbycity <- ggplot(df[df$state=="California",], aes(x=reorder(city, city, length), y=log_ppm)) +
-geom_boxplot() +
-labs(title="log PPM by city (CA)", x="Log PPM",y="city")
-grid.arrange(fl_lppmbycity, ca_lppmbycity, ncol= 2)
-```
-
+We find that there are many cities with a sample size of one, which makes `city` perhaps not the most natural choice for a grouping variable (please see appendix).
 We decided to focus on `state` as our primary candidate for a grouping variable moving forward.
@@
   theme(axis.text.x = element_text(angle = 90))
-```
-
-## More EDA: slope and intercept distributions for predictors by state
-
-To get a better sense of how state-level groups and slopes are distributed, we created the following intercepts for our coninuous variables:
-
-### Dose strength variable
-
-```{r fig.width=3}
-group_int_list <- NULL 
-group_slope_list <- NULL
-for(statej in unique(df$state)) {
-    m <- lm(data = df[df$state==statej, ], log_ppm ~ dose_per_mg)
-    group_int_list <- c(group_int_list, coef(m)[1])
-    group_slope_list <- c(group_slope_list, coef(m)[2])
-}
-hist(group_int_list)
-hist(group_slope_list)
-```
-
-### Continuous date variable
-
-```{r fig.width=3}
-group_int_list <- NULL 
-group_slope_list <- NULL
-for(statej in unique(df$state)) {
-    m <- lm(data = df[df$state==statej, ], log_ppm ~ date_num)
-    group_int_list <- c(group_int_list, coef(m)[1])
-    group_slope_list <- c(group_slope_list, coef(m)[2])
-}
-
-hist(group_int_list)
-hist(group_slope_list)
 ```
@@
 ```{r, ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE}
+# Example visualization shown here are for all cities in the state of Florida.
+
+ggplot(df[df$state=="Florida",], aes(x=log_ppm, y=reorder(city, city, length))) +
+geom_boxplot() +
+labs(title="log PPM by city (FL)", x="Log PPM",y="city")
+
+# More EDA: slope and intercept distributions for predictors by state
+
+# To get a better sense of how state-level groups and slopes are distributed, we created the following intercepts for our coninuous variables:
+
+### Dose strength variable
+
+group_int_list <- NULL 
+group_slope_list <- NULL
+for(statej in unique(df$state)) {
+    m <- lm(data = df[df$state==statej, ], log_ppm ~ dose_per_mg)
+    group_int_list <- c(group_int_list, coef(m)[1])
+    group_slope_list <- c(group_slope_list, coef(m)[2])
+}
+hist(group_int_list)
+hist(group_slope_list)
+
+### Continuous date variable
+
+group_int_list <- NULL 
+group_slope_list <- NULL
+for(statej in unique(df$state)) {
+    m <- lm(data = df[df$state==statej, ], log_ppm ~ date_num)
+    group_int_list <- c(group_int_list, coef(m)[1])
+    group_slope_list <- c(group_slope_list, coef(m)[2])
+}
+
+hist(group_int_list)
+hist(group_slope_list)
+
 ```

### 2021-03-22 final_report.Rmd
Made margins smaller

--- 2021-03-22 final_report.Rmd
@@
     df_print: paged
+fontsize: 11pt
+geometry: margin=2cm
 ---

### 2021-03-22 final_report.Rmd
Space saving

@@
 
-Based off of the results of the EDA, we want to initially fit a model with the following variables:
-
-- source
-- dose_per_mg
-- bulk_purchase
-- Primary_Reason
-- date_num
-- date_quarter
-
-In terms of the research question, we would also like to group based on location. There are 3 options for this:
-
-- Region
-- State
-- City
+Based off of the results of the EDA, we want to initially fit a model with the following variables: source, dose_per_mg, bulk_purchase, Primary_Reason, date_num, and date_quarter.
+In terms of the research question, we would also like to group based on location.
+There are 3 options for this: region, state, or city.
 
@@
 
-Thus these results suggest we want to drop `date_num` and `date_quarter`, while keeping
-
-* dose_per_mg
-* bulk_purchase
-* primary_reason
-* source
+Thus these results suggest we want to drop `date_num` and `date_quarter`, while keeping: dose_per_mg, bulk_purchase, primary_reason, source
 
@@
 
-* *dose_per_mg*: Controlling for other variables, for every one milligram increase of dosage, the price per mg of oxymorphone increases by a multiplicative effect of exp(-0.0151868) = 0.9849279, about a 1.51% reduction.
+*dose_per_mg*: Controlling for other variables, for every one milligram increase of dosage, the price per mg of oxymorphone increases by a multiplicative effect of exp(-0.0151868) = 0.9849279, about a 1.51% reduction.
 
-* *bulk_purchase* ("not bulk purchase" is the reference): Controlling for other variables, bulk purchase reduces the price per mg of oxymorphone by a multiplicative effect of exp(-0.1691121) = 0.8444142, about a 15.56% reduction.
+*bulk_purchase* ("not bulk purchase" is the reference): Controlling for other variables, bulk purchase reduces the price per mg of oxymorphone by a multiplicative effect of exp(-0.1691121) = 0.8444142, about a 15.56% reduction.
 
-* *source* (“Source: Unknown” is the reference, other levels ommitted for brevity):
-    + Controlling for other variables, purchasing oxymorphone with Internet source tends to increase the price per mg by a multiplicative effect of exp(0.1339468) = 1.143332, about a 14.33% increase than the price per mg of oxymorphone for people with unknown drug source.
+*source* (“Source: Unknown” is the reference, other levels ommitted for brevity): Controlling for other variables, purchasing oxymorphone with Internet source tends to increase the price per mg by a multiplicative effect of exp(0.1339468) = 1.143332, about a 14.33% increase than the price per mg of oxymorphone for people with unknown drug source.
 
-* *Primary Reason* ("Blank" is the reference, other levels omitted for brevity):
-    + Controlling for other variables, purchasing oxymorphone with primary reason Self-treat tends to increase the price per mg by a multiplicative effect of exp(0.1034974) = 1.109043, about a 10.9% increase than the price per mg of oxymorphone for people with Blank primary reason.
+*Primary Reason* ("Blank" is the reference, other levels omitted for brevity): Controlling for other variables, purchasing oxymorphone with primary reason Self-treat tends to increase the price per mg by a multiplicative effect of exp(0.1034974) = 1.109043, about a 10.9% increase than the price per mg of oxymorphone for people with Blank primary reason.
 

### 2021-03-22 final_report.Rmd
fixed table col names and fig.height for region

--- 2021-03-22 final_report.Rmd
@@
     df_print: paged
-fontsize: 11pt
-geometry: margin=2cm
 ---
@@
 ```{r}
-hist(df$date_num)
-knitr::kable(table(df$date_quarter))
+#hist(df$date_num)
+knitr::kable(table(df$date_quarter), col.names = c("Quarter","Freq"))
 ```
@@
 
-```{r, fig.align='center'}
+```{r, fig.align='center', fig.height = 2.5}
 ggplot(df, aes(x=reorder(USA_region, USA_region, length), y=log_ppm, fill=USA_region)) +

### 2021-03-22 STA_610_Case_Study_1_Report_Team_4.Rmd
small changes to the format

@@
 ```{r}
-hist(df$date_num)
-knitr::kable(table(df$date_quarter))
+knitr::kable(table(df$date_quarter), col.names = c("Quarter","Freq"))
 ```
@@
 
-```{r}
-hist(df$mgstr)
+```{r, fig.align='center', fig.height=3}
+#hist(df$mgstr)
 # rename variable for ease of interpretability
 df <- df %>% rename(dose_per_mg = mgstr)
+
+ggplot(df, aes(x=dose_per_mg)) + 
+ geom_histogram(aes(y=..density..), colour="black", fill="white", bins = 50)+
+ geom_density(alpha=.2, fill="red") 
 ```
@@
 geom_boxplot() +
-labs(title="State", x="Log PPM by state",y="Log PPM")
+labs(title="State", x="Log PPM",y="Log PPM by state")
 ```
@@
 
-```{r, fig.align='center'}
+```{r, fig.align='center', fig.height=2.5}
 ggplot(df, aes(x=reorder(USA_region, USA_region, length), y=log_ppm, fill=USA_region)) +
@@
 We find that there are many cities with a sample size of one, which makes `city` perhaps not the most natural choice for a grouping variable.
-Example visualizations shown here are for all cities in the states of Florida and California.
+We decided to focus on `state` as our primary candidate for a grouping variable moving forward.
 
-```{r, fig.align='center'}
+```{r, fig.align='center', include=FALSE}
 fl_lppmbycity <- ggplot(df[df$state=="Florida",], aes(x=log_ppm, y=reorder(city, city, length))) +
@@
 
-We decided to focus on `state` as our primary candidate for a grouping variable moving forward.
-
 ## Exploratory data analysis for random slopes
@@
 
-### dose_per_mg
+### dose_per_mg & date_num
 
 ```{r, fig.align='center'}
-ggplot(data = df[df$state=="Florida" | df$state=="Texas" | df$state=="North Carolina",], aes(x = dose_per_mg, y = log_ppm)) + 
+plot_dose_per_mg <- ggplot(data = df[df$state=="Florida" | df$state=="Texas" | df$state=="North Carolina",], aes(x = dose_per_mg, y = log_ppm)) + 
   geom_point() +
@@
 
-### date_num
-
-```{r, fig.align='center'}
-ggplot(data = df[df$state=="Florida" | df$state=="Texas" | df$state=="North Carolina",], aes(x = date_num, y = log_ppm)) + 
+```{r, fig.align='center', fig.height=3}
+plot_date_num <- ggplot(data = df[df$state=="Florida" | df$state=="Texas" | df$state=="North Carolina",], aes(x = date_num, y = log_ppm)) + 
   geom_point() +
   facet_grid(~ state)
+
+grid.arrange(plot_dose_per_mg, plot_date_num, ncol = 2)
 ```
@@
 
-### bulk_purchase
-```{r, fig.align='center'}
-# categorical variables -- looking at South only
-ggplot(data = df[df$USA_region == "South",], aes(x = bulk_purchase, y = log_ppm)) + 
+### bulk_purchase & date_quarter
+```{r, fig.align='center', fig.height=3}
+plot_bulk_purchase <- ggplot(data = df[df$state=="Florida" | df$state=="Texas" | df$state=="North Carolina",], aes(x = bulk_purchase, y = log_ppm, group = bulk_purchase)) + 
   geom_boxplot() +
@@
 
-### date_quarter
+```{r, fig.align='center'}
+plot_date_quarter <- ggplot(data = df[df$state=="Florida" | df$state=="Texas" | df$state=="North Carolina",], aes(x = date_quarter, y = log_ppm, group = date_quarter)) + 
+  geom_boxplot() +
+  facet_grid(~ state)
+
+grid.arrange(plot_bulk_purchase, plot_date_quarter, ncol = 2)
+```
+
+### Primary_Reason & source
 ```{r}
-ggplot(data = df[df$state=="Florida" | df$state=="Texas" | df$state=="North Carolina",], aes(x = date_quarter, y = log_ppm, group = date_quarter)) + 
+plot_Primary_Reason <- ggplot(data = df[df$state=="Florida" | df$state=="Texas" | df$state=="North Carolina",], aes(x = Primary_Reason, y = log_ppm, group = Primary_Reason)) + 
   geom_boxplot() +
@@
 
-### Primary_Reason
-```{r}
-ggplot(data = df[df$state=="Florida" | df$state=="Texas" | df$state=="North Carolina",], aes(x = Primary_Reason, y = log_ppm, group = Primary_Reason)) + 
+```{r, fig.align='center', fig.height=4}
+plot_source <- ggplot(data = df %>% filter(state %in% c('Florida', 'Texas', 'North Carolina')), aes(x = source, y = log_ppm, group = source)) + 
   geom_boxplot() +
   facet_grid(~ state)
-```
 
-### source
-
-```{r}
-ggplot(data = df %>% filter(state %in% c('Florida', 'Texas', 'North Carolina')), aes(x = source, y = log_ppm, group = source)) + 
-  geom_boxplot() +
-  facet_grid(~ state)
-```
-## More EDA: slope and intercept distributions for predictors by state
-
-To get a better sense of how state-level groups and slopes are distributed, we created the following intercepts for our coninuous variables:
-
-### Dose strength variable
-
-```{r}
-group_int_list <- NULL 
-group_slope_list <- NULL
-
-for(statej in
-  unique(df$state))
-{
-m <- lm(data = df[df$state==statej, ], log_ppm ~ dose_per_mg)
-
-group_int_list <- c(group_int_list, coef(m)[1])
-group_slope_list <- c(group_slope_list, coef(m)[2])
-}
-
-hist(group_int_list)
-hist(group_slope_list)
-```
-
-### Continuous date variable
-
-```{r}
-group_int_list <- NULL 
-group_slope_list <- NULL
-
-for(statej in
-  unique(df$state))
-{
-m <- lm(data = df[df$state==statej, ], log_ppm ~ date_num)
-
-group_int_list <- c(group_int_list, coef(m)[1])
-group_slope_list <- c(group_slope_list, coef(m)[2])
-}
-
-hist(group_int_list)
-hist(group_slope_list)
+grid.arrange(plot_Primary_Reason, plot_source, ncol= 2)
 ```

### 2021-03-22 final_report.Rmd
Shorter report

--- 2021-03-22 final_report.Rmd
@@
 ---
-title: "STA610 Case Study 1 Team 4 Report"
+title: \vspace{-1.5cm} "STA610 Case Study 1 Team 4 Report"
 author:
@@
 
-The outline is as follows:
-
-* Start with date_num, date_quarter, dose_per_mg, bulk_purchase, Primary_Reason, source
-* Perform a LRT between competing models with and without the date variable.
-Observe a p-value of `r round(date_num_pval, 3)` and fail to reject the null (simpler model).
-Drop `date_num` from the subsequent model tests
-* Perform a LRT between competing models with and without the quarter of the sale.
-Observe a p-value of `r round(quarter_pval, 3)` and fail to reject the null (simpler model).
-Drop `date_quarter` from the subsequent model tests
-* Perform a LRT between competing models with and without dose per mg. Observe a p-value of ~`r round(dose_pval, 3)` and reject the null (simpler model). Decide on model keeping `dose_per_mg`
-* Perform a LRT between competing models with and without the bulk purchase predictor.
-Observe a p-value of ~`r round(bulk_pval, 3)` and reject the null (simpler model).
-Decide on model keeping `bulk_purchase`
-* Perform a LRT between competing models with and without the primary reason predictor.
-Observe a p-value of `r round(reason_pval, 3)` and reject the null (simpler model).
-Decide on keeping `primary_reason` in the model
-* Perform a LRT between competing models with and without the source predictor.
-Observe a p-value of ~`r round(source_pval, 3)` and reject the null (simpler model).
-Decide on keeping `source` in the model
-
-Thus these results suggest we want to drop `date_num` and `date_quarter`, while keeping: dose_per_mg, bulk_purchase, primary_reason, source
+Model selection is omitted for brevity (see code appendix for full demonstration).
+Essentially we compared nested models with fewer and fewer variables, provided the results of the Anova test suggested the simpler model was preferred.
+Ultimately, this led to us dropping `date_num` and `date_quarter`, while keeping: dose_per_mg, bulk_purchase, primary_reason, source
 

### 2021-03-22 STA_610_Case_Study_1_Report_Team_4.Rmd
formated fig.height and plots

--- 2021-03-21 STA_610_Case_Study_1_Report_Team_4.Rmd
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, cache=TRUE)
+knitr::opts_chunk$set(echo = FALSE, cache=TRUE, fig.height = 3)
 ```
@@
 ```{r, echo=FALSE, warning=FALSE, message=FALSE}
-library(tidyverse)
-library(lme4)
-library(gridExtra)
-library(grid)
-library(ggplot2)
-library(lattice)
+library(pacman)
+pacman::p_load(tidyverse, lme4, gridExtra, grid, ggplot2, lattice, redres, stringr, influence.ME, knitr)
 # devtools::install_github("goodekat/redres")
-#library(redres)
-library(stringr)
-library(influence.ME)
-library(knitr)
 ```
@@
 ```{r, echo = FALSE}
-#load("~/Documents/Spring 2021/STA 610 - H Models/cs1/streetrx.RData")
-load('streetrx.rdata')
-df <- streetrx %>% filter(api_temp=="oxymorphone", !is.na(ppm))
+load("streetrx.rdata")
+df <- streetrx %>% filter(api_temp == "oxymorphone", !is.na(ppm))
 ```
@@
 StreetRx maintains a partnership with the Researched Abuse, Diversion, and Addiction-Related Surveillance System (RADARS), which aims for these data to inform public health measures such as health policy related to illicit drugs.
-
-[insert something about our specific drug here]
 
@@
 
-```{r}
-hist(df$mgstr)
+```{r, fig.align='center', fig.height=2.5}
+ggplot(df, aes(x=mgstr)) + 
+ geom_histogram(aes(y=..density..), colour="black", fill="white", bins = 50)+
+ geom_density(alpha=.2, fill="red") +
+labs(x="does per mg",y="density")
 # rename variable for ease of interpretability
@@
 
-```{r, fig.height=3, fig.align='center'}
+```{r, fig.height=2.5, fig.align='center'}
 # log transform ppm
 df$log_ppm <- log(df$ppm)
-
 hist_ppm <- ggplot(df, aes(x=ppm)) + 
@@
 
-We find that there are many cities with a sample size of one, which makes `city` perhaps not the most natural choice for a grouping variable.
-Example visualizations shown here are for all cities in the states of Florida and California.
-
-```{r, fig.align='center'}
-fl_lppmbycity <- ggplot(df[df$state=="Florida",], aes(x=log_ppm, y=reorder(city, city, length))) +
-geom_boxplot() +
-labs(title="log PPM by city (FL)", x="Log PPM",y="city")
-
-ca_lppmbycity <- ggplot(df[df$state=="California",], aes(x=reorder(city, city, length), y=log_ppm)) +
-geom_boxplot() +
-labs(title="log PPM by city (CA)", x="Log PPM",y="city")
-
-grid.arrange(fl_lppmbycity, ca_lppmbycity, ncol= 2)
-```
-
+We find that there are many cities with a sample size of one, which makes `city` perhaps not the most natural choice for a grouping variable (please see appendix).
 We decided to focus on `state` as our primary candidate for a grouping variable moving forward.
@@
 ### dose_per_mg
-
-```{r, fig.align='center'}
+```{r, fig.align='center', fig.height = 2.5}
 ggplot(data = df[df$state=="Florida" | df$state=="Texas" | df$state=="North Carolina",], aes(x = dose_per_mg, y = log_ppm)) + 
@@
 ### date_num
-
-```{r, fig.align='center'}
+```{r, fig.align='center', fig.height = 2.5}
 ggplot(data = df[df$state=="Florida" | df$state=="Texas" | df$state=="North Carolina",], aes(x = date_num, y = log_ppm)) + 
@@
 ### bulk_purchase
-```{r, fig.align='center'}
+```{r, fig.align='center', fig.height = 2.5}
 # categorical variables -- looking at South only
-ggplot(data = df[df$USA_region == "South",], aes(x = bulk_purchase, y = log_ppm)) + 
+set.seed(42)
+states <- sample(df$state, 5)
+ggplot(data = df %>% filter(state %in% states), aes(x = bulk_purchase, y = log_ppm)) + 
   geom_boxplot() +
-  facet_grid(~ state)
+  facet_grid(~ state) +
+  theme(axis.text.x = element_text(angle = 90))
 ```
@@
 ### date_quarter
-```{r}
+```{r, fig.height = 2.5, fig.align='center'}
 ggplot(data = df[df$state=="Florida" | df$state=="Texas" | df$state=="North Carolina",], aes(x = date_quarter, y = log_ppm, group = date_quarter)) + 
@@
 ### Primary_Reason
-```{r}
+```{r, fig.height = 2.5, fig.align='center'}
 ggplot(data = df[df$state=="Florida" | df$state=="Texas" | df$state=="North Carolina",], aes(x = Primary_Reason, y = log_ppm, group = Primary_Reason)) + 
   geom_boxplot() +
-  facet_grid(~ state)
+  facet_grid(~ state) +
+  theme(axis.text.x = element_text(angle = 90))
 ```
@@
 ### source
-
-```{r}
-ggplot(data = df %>% filter(state %in% c('Florida', 'Texas', 'North Carolina')), aes(x = source, y = log_ppm, group = source)) + 
+```{r, fig.height = 2.5, fig.align='center'}
+ggplot(data = df[df$state=="Florida" | df$state=="Texas" | df$state=="North Carolina",], aes(x = source, y = log_ppm, group = source)) + 
   geom_boxplot() +
-  facet_grid(~ state)
-```
-## More EDA: slope and intercept distributions for predictors by state
-
-To get a better sense of how state-level groups and slopes are distributed, we created the following intercepts for our coninuous variables:
-
-### Dose strength variable
-
-```{r}
-group_int_list <- NULL 
-group_slope_list <- NULL
-
-for(statej in
-  unique(df$state))
-{
-m <- lm(data = df[df$state==statej, ], log_ppm ~ dose_per_mg)
-
-group_int_list <- c(group_int_list, coef(m)[1])
-group_slope_list <- c(group_slope_list, coef(m)[2])
-}
-
-hist(group_int_list)
-hist(group_slope_list)
-```
-
-### Continuous date variable
-
-```{r}
-group_int_list <- NULL 
-group_slope_list <- NULL
-
-for(statej in
-  unique(df$state))
-{
-m <- lm(data = df[df$state==statej, ], log_ppm ~ date_num)
-
-group_int_list <- c(group_int_list, coef(m)[1])
-group_slope_list <- c(group_slope_list, coef(m)[2])
-}
-
-hist(group_int_list)
-hist(group_slope_list)
+  facet_grid(~ state) + 
+  theme(axis.text.x = element_text(angle = 90))
 ```
@@
 mod_s <- lmer(data=df, log_ppm ~ (1 |state) + date_num + date_quarter + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F)
-
 mod_c <- lmer(data=df, log_ppm ~ (1 |city) + date_num + date_quarter + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F)
-
 mod_r <- lmer(data=df, log_ppm ~ (1 |USA_region) + date_num + date_quarter + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F)
@@
 date_num_pval <- drop_date_num$`Pr(>Chisq)`[2]
-
 # Test date_quarter
@@
 quarter_pval <- drop_quarter$`Pr(>Chisq)`[2]
-
 # Test dose_per_mg
@@
 dose_pval <- drop_dose$`Pr(>Chisq)`[2]
-
 # Test bulk purchase
@@
 bulk_pval <- drop_bulk$`Pr(>Chisq)`[2]
-
 # Test primary reason
@@
 reason_pval <- drop_reason$`Pr(>Chisq)`[2]
-
 # Test source
@@
 
-We will now investigate random slopes in addition to random intercepts. We are having difficulties fitting a random slope for `dose_per_mg`, but we can check the random slope for bulk purchase:
+#### Model Specification
 
-```{r}
-random_bulk <- anova(lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F),
-      lmer(data=df, log_ppm ~ (bulk_purchase |state) + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F))
-random_bulk$`Pr(>Chisq)`[2]  # Random slopes for bulk_purchase not helpful
-
-random_reason <- anova(lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F),
-                       lmer(data=df, log_ppm ~ (Primary_Reason |state) + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F))
-random_reason$`Pr(>Chisq)`[2]
-
-random_source <- anova(lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F),
-                       lmer(data=df, log_ppm ~ (source |state) + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F))
-random_source$`Pr(>Chisq)`[2]
-
-print(paste('For testing between competing models with and without a random slope on '))
-```
-
-These results suggest that we should stick with a random intercept only.
-
-#### Model Specification
+Our final model is as specified:
 
@@
 #confint(mod1)[3:11,]
-
 # extract coefficients
-coefs <- data.frame(coef(summary(mod1)))
+coefs <- data.frame(coef(summary(mod_final)))
 # use normal distribution to approximate p-value
@@
 
-* Bulk_purchase: “Not bulk purchase” is the reference group.
-Controlling for other variables, being bulk purchase increases the price per mg by exp(-0.16) = 0.8521. 
-* Primary_Reason: “Primary_Reason: Other or Unknown” is the reference group. 
-Controlling for other variables, taking oxymorphone with the primary reason of self-treating tends to receive a exp(0.05) = 1.0513 times higher price per mg than people whose primary reason is “other or unknown”.
-Controlling for other variables, taking oxymorphone with the primary reason of enjoyment tends to receive a exp(0.02) = 1.0202 times higher price per mg than people whose primary reason is “other or unknown”.
-* Source: “Source: Unknown” is the reference group. 
-Controlling for other variables, purchasing oxymorphone with Internet source tends to receive a exp(0.02) = 1.0202 times higher price per mg than people with unknown drug source.
-Controlling for other variables, purchasing oxymorphone with personal source tends to receive a exp(0.07) = 1.0725 times higher price per mg than people with unknown drug source.
-Controlling for other variables, purchasing oxymorphone with “heard it” source tends to receive a exp(0.24) = 1.2712 times higher price per mg than people with unknown drug source.
+* *dose_per_mg*: Controlling for other variables, for every one milligram increase of dosage, the price per mg of oxymorphone increases by a multiplicative effect of exp(-0.0151868) = 0.9849279, about a 1.51% reduction.
+
+* *bulk_purchase* ("not bulk purchase" is the reference): Controlling for other variables, bulk purchase reduces the price per mg of oxymorphone by a multiplicative effect of exp(-0.1691121) = 0.8444142, about a 15.56% reduction.
+
+* *source* (“Source: Unknown” is the reference, other levels ommitted for brevity):
+    + Controlling for other variables, purchasing oxymorphone with Internet source tends to increase the price per mg by a multiplicative effect of exp(0.1339468) = 1.143332, about a 14.33% increase than the price per mg of oxymorphone for people with unknown drug source.
+
+* *Primary Reason* ("Blank" is the reference, other levels omitted for brevity):
+    + Controlling for other variables, purchasing oxymorphone with primary reason Self-treat tends to increase the price per mg by a multiplicative effect of exp(0.1034974) = 1.109043, about a 10.9% increase than the price per mg of oxymorphone for people with Blank primary reason.
 
@@
 ```{r}
-#VarCorr(mod1)
-Variance <- c(0.04748, 0.96369)
-Std.Dev <- c(0.2179, 0.9817)
+#VarCorr(mod_final)
+Variance <- c(0.02292, 0.50060)
+Std.Dev <- c(0.15140 , 0.70753)
 Groups <- c("state (Intercept)", "Residual")
+REvar <- cbind(Groups, Variance, Std.Dev)
+knitr::kable(REvar, col.names = c("Group", "Variance", "Std.Dev"))
+```
+The within-state variance $\sigma^2$ is 0.5006, which is much larger than the across-state variance $\tau^2$, 0.02292.
 
-REvar <- cbind(Groups, Variance, Std.Dev)
+```{r, include=FALSE}
+library(sjstats)
+icc(mod_final)
+```
 
-knitr::kable(REvar, col.names = c("Group", "Variance", "Std.Dev"))
+```{r}
+ICC <- 0.044
+knitr::kable(ICC, col.names = "Intraclass Correlation")
+```
+Intraclass correlation is 0.044, which means values are not very similar from the same state.
+
+```{r, fig.align='center', fig.height=8}
+dotplot(ranef(mod_final, condVar=TRUE))$state
+```
+From the result of the random effects, we can see that the random intercept for Tennessee is the highest and it is significantly different from zero.
+The random intercept for Arizona is the lowest and this intercept is also significant.
+
+## Appendix
+```{r, ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE}
+# Example visualization shown here are for all cities in the state of Florida.
+
+ggplot(df[df$state=="Florida",], aes(x=log_ppm, y=reorder(city, city, length))) +
+geom_boxplot() +
+labs(title="log PPM by city (FL)", x="Log PPM",y="city")
+
+# More EDA: slope and intercept distributions for predictors by state
+
+# To get a better sense of how state-level groups and slopes are distributed, we created the following intercepts for our coninuous variables:
+
+### Dose strength variable
+
+group_int_list <- NULL 
+group_slope_list <- NULL
+for(statej in unique(df$state)) {
+    m <- lm(data = df[df$state==statej, ], log_ppm ~ dose_per_mg)
+    group_int_list <- c(group_int_list, coef(m)[1])
+    group_slope_list <- c(group_slope_list, coef(m)[2])
+}
+hist(group_int_list)
+hist(group_slope_list)
+
+### Continuous date variable
+
+group_int_list <- NULL 
+group_slope_list <- NULL
+for(statej in unique(df$state)) {
+    m <- lm(data = df[df$state==statej, ], log_ppm ~ date_num)
+    group_int_list <- c(group_int_list, coef(m)[1])
+    group_slope_list <- c(group_slope_list, coef(m)[2])
+}
+
+hist(group_int_list)
+hist(group_slope_list)
 
 ```
-The within-state variance $\sigma^2$ is 0.96369, which is much larger than the across-state variance $\tau^2$, 0.04748.
-
-```{r, fig.align='center', fig.height=8}
-dotplot(ranef(mod1, condVar=TRUE))$state
-```
-From the result of the random effects, we can see that the random intercept for Tennessee is the highest and it is significantly different from zero.
-The random intercept for Nevada is the lowest and this intercept is also significant.
-
-## Appendix
-```{r, ref.label=knitr::all_labels(),echo=TRUE,eval=FALSE}
-```

### 2021-03-22 STA_610_Case_Study_1_Report_Team_4.Rmd
formatted the EDA sections.

@@
 We renamed this variable `dose_per_mg` to make its meaning clearer to the reader.
-This is a continuous variable that nevertheless tends to take on only a few values, as depicted in the histogram.
+This is a continuous variable that nevertheless tends to take on only a few values.
 We considered a recoding of this variable into levels, but decided against it as the dose per milligram technically may take on any value and the generalizability of our model may be improved if we are able to consider dose levels that are in a realistic range yet not present in our current data.
 
-```{r, fig.align='center', fig.height=2.5}
+```{r, fig.align='center', fig.height=2, include = FALSE}
 ggplot(df, aes(x=mgstr)) + 
@@
 
-### dose_per_mg
+### dose_per_mg & date_num
 ```{r, fig.align='center', fig.height = 2.5}
-ggplot(data = df[df$state=="Florida" | df$state=="Texas" | df$state=="North Carolina",], aes(x = dose_per_mg, y = log_ppm)) + 
+plot_dose_per_mg <- ggplot(data = df[df$state=="Florida" | df$state=="Texas" | df$state=="North Carolina",], aes(x = dose_per_mg, y = log_ppm)) + 
   geom_point() +
@@
 
-### date_num
 ```{r, fig.align='center', fig.height = 2.5}
-ggplot(data = df[df$state=="Florida" | df$state=="Texas" | df$state=="North Carolina",], aes(x = date_num, y = log_ppm)) + 
+plot_date_num <- ggplot(data = df[df$state=="Florida" | df$state=="Texas" | df$state=="North Carolina",], aes(x = date_num, y = log_ppm)) + 
   geom_point() +
   facet_grid(~ state)
+
+grid.arrange(plot_dose_per_mg, plot_date_num, ncol = 2)
 ```
@@
 
-### bulk_purchase
+### bulk_purchase & date_quarter
 ```{r, fig.align='center', fig.height = 2.5}
@@
 states <- sample(df$state, 5)
-ggplot(data = df %>% filter(state %in% states), aes(x = bulk_purchase, y = log_ppm)) + 
+box_bulk_purchase <- ggplot(data = df[df$state=="Florida" | df$state=="Texas" | df$state=="North Carolina",], aes(x = bulk_purchase, y = log_ppm, group = bulk_purchase)) + 
   geom_boxplot() +
@@
 
-### date_quarter
 ```{r, fig.height = 2.5, fig.align='center'}
-ggplot(data = df[df$state=="Florida" | df$state=="Texas" | df$state=="North Carolina",], aes(x = date_quarter, y = log_ppm, group = date_quarter)) + 
+box_date_quarter <- ggplot(data = df[df$state=="Florida" | df$state=="Texas" | df$state=="North Carolina",], aes(x = date_quarter, y = log_ppm, group = date_quarter)) + 
   geom_boxplot() +
   facet_grid(~ state)
+
+grid.arrange(box_bulk_purchase, box_date_quarter, ncol=2)
 ```
 
-### Primary_Reason
+### Primary_Reason & source
 ```{r, fig.height = 2.5, fig.align='center'}
-ggplot(data = df[df$state=="Florida" | df$state=="Texas" | df$state=="North Carolina",], aes(x = Primary_Reason, y = log_ppm, group = Primary_Reason)) + 
+box_Primary_Reason <- ggplot(data = df[df$state=="Florida" | df$state=="North Carolina",], aes(x = Primary_Reason, y = log_ppm, group = Primary_Reason)) + 
   geom_boxplot() +
@@
 
-### source
 ```{r, fig.height = 2.5, fig.align='center'}
-ggplot(data = df[df$state=="Florida" | df$state=="Texas" | df$state=="North Carolina",], aes(x = source, y = log_ppm, group = source)) + 
+box_source <- ggplot(data = df[df$state=="Florida" | df$state=="North Carolina",], aes(x = source, y = log_ppm, group = source)) + 
   geom_boxplot() +
@@
   theme(axis.text.x = element_text(angle = 90))
+
+grid.arrange(box_Primary_Reason, box_source, ncol = 2)
 ```

### 2021-03-22 STA_610_Case_Study_1_Report_Team_4.Rmd
updated the intro

@@
 StreetRx maintains a partnership with the Researched Abuse, Diversion, and Addiction-Related Surveillance System (RADARS), which aims for these data to inform public health measures such as health policy related to illicit drugs.
+
+Our drug of interest in this study is oxymorphone.
+Oxymorphone is used to relieve moderate to severe pain in people whose pain is not controlled with other medications, and may be habit-forming with prolonged use.
+Our aim is to investigate factors related to the price per mg of oxymorphone, accounting for potential clustering by location and exploring heterogeneity in pricing by location.
 

### 2021-03-22 final_report_2.Rmd

--- 2021-03-22 final_report.Rmd
@@
   pdf_document: default
-  html_document:
-    df_print: paged
 fontsize: 11pt
@@
 For the seasonality variable, `date_quarter`, we extract the "quarter" of the year (numbered 1 through 4) as its own variable.
-Below we see the distributions of these variables:
+Please see appendix for distributions.
 
@@
 df$date_quarter <- as.numeric(substr(df$yq_pdate, 5, 5))
-```
-
-```{r}
-#hist(df$date_num)
-knitr::kable(table(df$date_quarter), col.names = c("Quarter","Freq"))
 ```
@@
 We renamed this variable `dose_per_mg` to make its meaning clearer to the reader.
-This is a continuous variable that nevertheless tends to take on only a few values, as depicted in the histogram.
+This is a continuous variable that nevertheless tends to take on only a few values (please see appendix), as depicted in the histogram.
 We considered a recoding of this variable into levels, but decided against it as the dose per milligram technically may take on any value and the generalizability of our model may be improved if we are able to consider dose levels that are in a realistic range yet not present in our current data.
@@
 ```{r}
-hist(df$mgstr)
 # rename variable for ease of interpretability
@@
 To consider whether we may want to include random slopes by state for our predictor variables, we examined plots of the predictor versus response among each state individually to determine whether it seemed that slopes fit to the group-level data would differ.
-(We present a sample of visualizations here rather than visualizations for every state, due to space constraints.)
+(We present a sample of visualizations rather than visualizations for every state, due to space constraints.
+Please see appendix)
 
-For out continuous variables (continuous date and dosage strength), in by-state plots we often saw strikingly similar patterns in predictor vs.
+For our continuous variables (continuous date and dosage strength), in by-state plots we often saw strikingly similar patterns in predictor vs.
 response, suggesting that random slopes may not be necessary for these variables.
-
-### dose_per_mg
-```{r, fig.align='center'}
-ggplot(data = df[df$state=="Florida" | df$state=="Texas" | df$state=="North Carolina",], aes(x = dose_per_mg, y = log_ppm)) + 
-  geom_point() +
-  facet_grid(~ state)
-```
-
-### date_num
-```{r, fig.align='center'}
-ggplot(data = df[df$state=="Florida" | df$state=="Texas" | df$state=="North Carolina",], aes(x = date_num, y = log_ppm)) + 
-  geom_point() +
-  facet_grid(~ state)
-```
-
 For categorical variables, we generally observed was there was indeed variation among the states, but it was not clear whether this variation had any well-defined structure.
 It was not clear whether our models would benefit from random slopes for these variables, so we decided to further investigate this in our modeling process.
-
-### bulk_purchase
-```{r, fig.align='center'}
-# categorical variables -- looking at South only
-set.seed(42)
-states <- sample(df$state, 5)
-ggplot(data = df %>% filter(state %in% states), aes(x = bulk_purchase, y = log_ppm)) + 
-  geom_boxplot() +
-  facet_grid(~ state) +
-  theme(axis.text.x = element_text(angle = 90))
-```
-
-### date_quarter
-```{r}
-ggplot(data = df[df$state=="Florida" | df$state=="Texas" | df$state=="North Carolina",], aes(x = date_quarter, y = log_ppm, group = date_quarter)) + 
-  geom_boxplot() +
-  facet_grid(~ state)
-```
-
-### Primary_Reason
-```{r}
-ggplot(data = df[df$state=="Florida" | df$state=="Texas" | df$state=="North Carolina",], aes(x = Primary_Reason, y = log_ppm, group = Primary_Reason)) + 
-  geom_boxplot() +
-  facet_grid(~ state) +
-  theme(axis.text.x = element_text(angle = 90))
-```
-
-### source
-```{r}
-ggplot(data = df[df$state=="Florida" | df$state=="Texas" | df$state=="North Carolina",], aes(x = source, y = log_ppm, group = source)) + 
-  geom_boxplot() +
-  facet_grid(~ state) + 
-  theme(axis.text.x = element_text(angle = 90))
-```
 
@@
 
-## Appendix
+# Appendix
+
+## Code
+
 ```{r, ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE}
-# Example visualization shown here are for all cities in the state of Florida.
 
+```
+
+## Visualizations
+
+### Date variable  
+```{r}
+hist(df$date_num)
+knitr::kable(table(df$date_quarter))
+```
+
+### Dose strength variable 
+```{r}
+hist(df$dose_per_mg)
+```
+
+### Region box plots
+```{r}
+ggplot(df, aes(x=reorder(USA_region, USA_region, length), y=log_ppm, fill=USA_region)) +
+geom_boxplot() +
+labs(title="log PPM by region", x="region",y="Log PPM")
+```
+
+### dose_per_mg
+```{r, fig.align='center'}
+ggplot(data = df[df$state=="Florida" | df$state=="Texas" | df$state=="North Carolina",], aes(x = dose_per_mg, y = log_ppm)) + 
+  geom_point() +
+  facet_grid(~ state)
+```
+
+### date_num
+```{r, fig.align='center'}
+ggplot(data = df[df$state=="Florida" | df$state=="Texas" | df$state=="North Carolina",], aes(x = date_num, y = log_ppm)) + 
+  geom_point() +
+  facet_grid(~ state)
+```
+
+### bulk_purchase
+```{r, fig.align='center'}
+# categorical variables -- looking at South only
+set.seed(42)
+states <- sample(df$state, 5)
+ggplot(data = df %>% filter(state %in% states), aes(x = bulk_purchase, y = log_ppm)) + 
+  geom_boxplot() +
+  facet_grid(~ state) +
+  theme(axis.text.x = element_text(angle = 90))
+```
+
+### date_quarter
+```{r}
+ggplot(data = df[df$state=="Florida" | df$state=="Texas" | df$state=="North Carolina",], aes(x = date_quarter, y = log_ppm, group = date_quarter)) + 
+  geom_boxplot() +
+  facet_grid(~ state)
+```
+
+### Primary_Reason
+```{r}
+ggplot(data = df[df$state=="Florida" | df$state=="Texas" | df$state=="North Carolina",], aes(x = Primary_Reason, y = log_ppm, group = Primary_Reason)) + 
+  geom_boxplot() +
+  facet_grid(~ state) +
+  theme(axis.text.x = element_text(angle = 90))
+```
+
+### source
+```{r}
+ggplot(data = df[df$state=="Florida" | df$state=="Texas" | df$state=="North Carolina",], aes(x = source, y = log_ppm, group = source)) + 
+  geom_boxplot() +
+  facet_grid(~ state) + 
+  theme(axis.text.x = element_text(angle = 90))
+```
+
+### City visualization for FL
+```{r}
 ggplot(df[df$state=="Florida",], aes(x=log_ppm, y=reorder(city, city, length))) +
@@
 labs(title="log PPM by city (FL)", x="Log PPM",y="city")
+```
 
-# More EDA: slope and intercept distributions for predictors by state
-
+### More EDA: slope and intercept distributions for predictors by state
+```{r}
 # To get a better sense of how state-level groups and slopes are distributed, we created the following intercepts for our coninuous variables:
@@
 hist(group_slope_list)
-
 ```

### 2021-03-22 final_report_2.Rmd
Added description of oxymorphone

@@
 ---
-title: \vspace{-1.5cm} "STA610 Case Study 1 Team 4 Report"
+title: \vspace{-1.5cm} \textbf{STA610 Case Study 1 Team 4 Report}
 author:
@@
 
-[insert something about our specific drug here]
+Our drug of interest in this study is oxymorphone.
+Oxymorphone is used to relieve moderate to severe pain in people whose pain is not controlled with other medications, and may be habit-forming with prolonged use.
+Our aim is to investigate factors related to the price per mg of oxymorphone, accounting for potential clustering by location and exploring heterogeneity in pricing by location.
 
@@
 $$log(y_{ij}) = \beta_{0,j} + \beta_{1} x_{ij1} + \beta_{2} x_{ij2} + \beta_{3} x_{ij3} + \beta_{4} x_{ij4} + \epsilon_{ij}$$
-
-$$
-\beta_{0,j} = \beta_0 + b_{0,j}, b_{0,j} \sim N(0,\tau^2)
-$$
-$$
-\epsilon_{ij} \perp b_{0,j} \overset{iid}{\sim} N(0,\sigma^2)
-$$
+$$\beta_{0,j} = \beta_0 + b_{0,j}, b_{0,j} \sim N(0,\tau^2)$$
+$$\epsilon_{ij} \perp b_{0,j} \overset{iid}{\sim} N(0,\sigma^2)$$
 

### 2021-03-22 STA_610_Case_Study_1_Report_Team_4.Rmd
format changed

@@
 library(pacman)
-pacman::p_load(tidyverse, lme4, gridExtra, grid, ggplot2, lattice, redres, stringr, influence.ME, knitr)
+pacman::p_load(tidyverse, 
+               lme4, 
+               gridExtra, 
+               grid, 
+               ggplot2, 
+               lattice, 
+               redres, 
+               stringr, 
+               influence.ME, 
+               knitr, 
+               merTools)
 # devtools::install_github("goodekat/redres")
@@
 
-```{r, fig.height=2.5, fig.align='center'}
+```{r, fig.height=2, fig.align='center'}
 # log transform ppm
@@
 
-```{r, fig.align='center', fig.height = 2.5}
+```{r, fig.align='center', fig.height = 2}
 ggplot(df, aes(x=reorder(USA_region, USA_region, length), y=log_ppm, fill=USA_region)) +
@@
 ### dose_per_mg & date_num
-```{r, fig.align='center', fig.height = 2.5}
-plot_dose_per_mg <- ggplot(data = df[df$state=="Florida" | df$state=="Texas" | df$state=="North Carolina",], aes(x = dose_per_mg, y = log_ppm)) + 
+```{r, fig.align='center', fig.height = 2.5, warning=FALSE, message=FALSE}
+plot_dose_per_mg <- ggplot(data = df[df$state=="Florida" | df$state=="North Carolina",], aes(x = dose_per_mg, y = log_ppm)) + 
   geom_point() +
+  geom_smooth(method=lm) +
   facet_grid(~ state)
@@
 
-```{r, fig.align='center', fig.height = 2.5}
-plot_date_num <- ggplot(data = df[df$state=="Florida" | df$state=="Texas" | df$state=="North Carolina",], aes(x = date_num, y = log_ppm)) + 
+```{r, fig.align='center', fig.height = 2.5, warning=FALSE, message=FALSE}
+plot_date_num <- ggplot(data = df[df$state=="Florida" | df$state=="North Carolina",], aes(x = date_num, y = log_ppm)) + 
   geom_point() +
+  geom_smooth(method=lm) +
   facet_grid(~ state)
@@
 dotplot(ranef(mod_final, condVar=TRUE))$state
+
+#plotREsim(REsim(mod_final,n.sims=100),stat='median',sd=TRUE, labs = TRUE)
 ```

### 2021-03-22 final 3.Rmd
small changes

--- 2021-03-22 final_report_2.Rmd
@@
 ---
-title: \vspace{-1.5cm} \textbf{STA610 Case Study 1 Team 4 Report}
+title: "\vspace{-1.5cm} \textbf{STA610 Case Study 1 Team 4 Report}"
 author:
@@
 output:
+  html_document:
+    df_print: paged
   pdf_document: default
@@
 library(pacman)
-pacman::p_load(tidyverse, lme4, gridExtra, grid, ggplot2, lattice, redres, stringr, influence.ME, knitr)
+library(merTools)
+pacman::p_load(tidyverse, lme4, gridExtra, grid, ggplot2, lattice, redres, stringr,
+               influence.ME, knitr)
 # devtools::install_github("goodekat/redres")
@@
 The benefits of a hierarchical model include the ability to "borrow information" from the data set as a whole to stabilize estimates of price for groups (e.g., states) with relatively small sample size, and to establish a estimated distribution of how price varies with respect to groups, which would in theory allow generalization to new groups not included in the original analysis.
+
 # Exploratory data analysis and data cleaning
@@
 df$source <- droplevels(df$source)
-knitr::kable(table(df$source), col.names = c("Level", "Freq"))
+knitr::kable(table(df$source), col.names = c("Level", "Freq"), caption = "Frequency Table of Variable source")
 ```
@@
 Our visualization of the bulk purchase variable indicates nothing unexpected that calls for data cleaning.
-
-```{r}
-#table(df$bulk_purchase, useNA = "always")
-# looks fine
-```
 
@@
 #print(reason_df)
-knitr::kable(reason_df_2, col.names = c("Level", "Freq"))
+knitr::kable(reason_df_2, col.names = c("Level", "Freq"), caption = "Frequency Table of Variable Primary_Reason")
 ```
@@
 # Fit 3 models grouping on different location specificity
-mod_s <- lmer(data=df, log_ppm ~ (1 |state) + date_num + date_quarter + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F)
-mod_c <- lmer(data=df, log_ppm ~ (1 |city) + date_num + date_quarter + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F)
-mod_r <- lmer(data=df, log_ppm ~ (1 |USA_region) + date_num + date_quarter + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F)
+mod_s <- lmer(data=df, log_ppm ~ (1 |state) + date_num + date_quarter + dose_per_mg +
+                bulk_purchase + Primary_Reason + source, REML=F)
+mod_c <- lmer(data=df, log_ppm ~ (1 |city) + date_num + date_quarter + dose_per_mg +
+                bulk_purchase + Primary_Reason + source, REML=F)
+mod_r <- lmer(data=df, log_ppm ~ (1 |USA_region) + date_num + date_quarter + dose_per_mg +
+                bulk_purchase + Primary_Reason + source, REML=F)
 ```
@@
 drop_date_num <- anova(mod_s,
-      lmer(data=df, log_ppm ~ (1 |state) + date_quarter + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F))
+      lmer(data=df, log_ppm ~ (1 |state) + date_quarter + dose_per_mg + bulk_purchase +
+             Primary_Reason + source, REML=F))
 date_num_pval <- drop_date_num$`Pr(>Chisq)`[2]
@@
                     bulk_purchase + Primary_Reason + source, REML=F),
-      lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F))
+                    lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase +
+                           Primary_Reason + source, REML=F))
 quarter_pval <- drop_quarter$`Pr(>Chisq)`[2]
 # Test dose_per_mg
-drop_dose <- anova(lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F),
-      lmer(data=df, log_ppm ~ (1 |state) + bulk_purchase + Primary_Reason + source, REML=F))
+drop_dose <- anova(lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase 
+                        + Primary_Reason + source, REML=F),
+                   lmer(data=df, log_ppm ~ (1 |state) + bulk_purchase + Primary_Reason 
+                        + source, REML=F))
 dose_pval <- drop_dose$`Pr(>Chisq)`[2]
 # Test bulk purchase
-drop_bulk <- anova(lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F),
-      lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + Primary_Reason + source, REML=F))
+drop_bulk <- anova(lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase 
+                        + Primary_Reason + source, REML=F),
+                   lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + Primary_Reason 
+                        + source, REML=F))
 bulk_pval <- drop_bulk$`Pr(>Chisq)`[2]
 # Test primary reason
-drop_reason <- anova(lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F),
-      lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase + source, REML=F))
+drop_reason <- anova(lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase 
+                          + Primary_Reason + source, REML=F),
+                     lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase 
+                          + source, REML=F))
 reason_pval <- drop_reason$`Pr(>Chisq)`[2]
 # Test source
-drop_source <- anova(lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F),
-      lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase + Primary_Reason, REML=F))
+drop_source <- anova(lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase 
+                          + Primary_Reason + source, REML=F),
+                     lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase 
+                          + Primary_Reason, REML=F))
 source_pval <- drop_source$`Pr(>Chisq)`[2]
@@
 Essentially we compared nested models with fewer and fewer variables, provided the results of the Anova test suggested the simpler model was preferred.
-Ultimately, this led to us dropping `date_num` and `date_quarter`, while keeping: dose_per_mg, bulk_purchase, primary_reason, source
+Ultimately, this led to us dropping `date_num` and `date_quarter`, while keeping: dose_per_mg, bulk_purchase, primary_reason, source.
 
@@
 ```{r, fig.align='center', warning=FALSE, message=FALSE, error=FALSE}
-# resqq <- plot_resqq(mod1)
-# ranefci <- plot_ranef(mod1)
-# grid.arrange(resqq, ranefci, ncol = 2)
+resqq <- plot_resqq(mod_final)
+ranefci <- plot_ranef(mod_final)
+grid.arrange(resqq, ranefci, ncol = 2)
 ```
 
-Model diagnostics discussions here.
+From the normal quantile plot of the raw conditional residuals, although there are many residuals out of bounds, these residuals are not too far away from the bounds.
+Thus, we can still accept the raw conditional residuals are normally distributed.
+From the normal quantile plot on random effect state, there are only two outliers. Thus, we can accept random effect of state is normally distributed.
 
 ## Results
+
 #### Fixed Effects
+
 ```{r, message=FALSE}
-#knitr::kable(fixef(mod1))
-#confint(mod1)[3:11,]
-# extract coefficients
 coefs <- data.frame(coef(summary(mod_final)))
@@
 
-*dose_per_mg*: Controlling for other variables, for every one milligram increase of dosage, the price per mg of oxymorphone increases by a multiplicative effect of exp(-0.0151868) = 0.9849279, about a 1.51% reduction.
+* *dose_per_mg*: Controlling for other variables, for every one milligram increase of dosage, the price per mg of oxymorphone reduces by a multiplicative effect of exp(-0.0152224) = 0.9848929, about a 1.51% reduction.
 
-*bulk_purchase* ("not bulk purchase" is the reference): Controlling for other variables, bulk purchase reduces the price per mg of oxymorphone by a multiplicative effect of exp(-0.1691121) = 0.8444142, about a 15.56% reduction.
+* *bulk_purchase* ("not bulk purchase" is the reference): Controlling for other variables, bulk purchase reduces the price per mg of oxymorphone by a multiplicative effect of exp(-0.1716889) = 0.8422412, about a 15.76% reduction than the price per mg of oxymorphone for people without bulk purchase.
 
-*source* (“Source: Unknown” is the reference, other levels ommitted for brevity): Controlling for other variables, purchasing oxymorphone with Internet source tends to increase the price per mg by a multiplicative effect of exp(0.1339468) = 1.143332, about a 14.33% increase than the price per mg of oxymorphone for people with unknown drug source.
+* *source* (“Source: Unknown” is the reference, other levels ommitted for brevity):
+    + Controlling for other variables, purchasing oxymorphone with Internet source tends to increase the price per mg by a multiplicative effect of exp(0.1361942) = 1.145904, about a 14.59% increase than the price per mg of oxymorphone for people with unknown drug source.
 
-*Primary Reason* ("Blank" is the reference, other levels omitted for brevity): Controlling for other variables, purchasing oxymorphone with primary reason Self-treat tends to increase the price per mg by a multiplicative effect of exp(0.1034974) = 1.109043, about a 10.9% increase than the price per mg of oxymorphone for people with Blank primary reason.
+* *Primary Reason* ("Blank" is the reference, other levels omitted for brevity):
+    + Controlling for other variables, purchasing oxymorphone with primary reason Self-treat tends to increase the price per mg by a multiplicative effect of exp(0.1042256) = 1.109851, about a 10.99% increase than the price per mg of oxymorphone for people with Blank primary reason.
 
 #### Random Effects
+
+```{r, include=FALSE}
+VarCorr(mod_final)
+```
+
 ```{r}
-#VarCorr(mod_final)
 Variance <- c(0.02292, 0.50060)
@@
 ```{r, fig.align='center', fig.height=8}
-dotplot(ranef(mod_final, condVar=TRUE))$state
+plotREsim(REsim(mod_final,n.sims=100),stat='median',sd=TRUE, labs = TRUE)
 ```
@@
 The random intercept for Arizona is the lowest and this intercept is also significant.
+
+## Appendix
+
+```{r, ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE}
+```
 
@@
 ```{r}
-hist(df$date_num)
-knitr::kable(table(df$date_quarter))
+hist(df$date_num, xlab = "data_num", main = "Histogram of date_num")
+knitr::kable(table(df$date_quarter), caption = "Frequency Table of Quarter in date_num")
 ```
@@
 ```{r}
-hist(df$dose_per_mg)
+hist(df$dose_per_mg, xlab = "does_per_mg", main = "Histogram of dose_per_mg")
 ```
@@
 ```{r, fig.align='center'}
-ggplot(data = df[df$state=="Florida" | df$state=="Texas" | df$state=="North Carolina",], aes(x = dose_per_mg, y = log_ppm)) + 
+ggplot(data = df[df$state=="Florida" | df$state=="Texas" | df$state=="North Carolina",],
+       aes(x = dose_per_mg, y = log_ppm)) + 
   geom_point() +
@@
 ```{r, fig.align='center'}
-ggplot(data = df[df$state=="Florida" | df$state=="Texas" | df$state=="North Carolina",], aes(x = date_num, y = log_ppm)) + 
+ggplot(data = df[df$state=="Florida" | df$state=="Texas" | df$state=="North Carolina",],
+       aes(x = date_num, y = log_ppm)) + 
   geom_point() +
@@
 ```{r}
-ggplot(data = df[df$state=="Florida" | df$state=="Texas" | df$state=="North Carolina",], aes(x = date_quarter, y = log_ppm, group = date_quarter)) + 
+ggplot(data = df[df$state=="Florida" | df$state=="Texas" | df$state=="North Carolina",],
+       aes(x = date_quarter, y = log_ppm, group = date_quarter)) + 
   geom_boxplot() +
@@
 ```{r}
-ggplot(data = df[df$state=="Florida" | df$state=="Texas" | df$state=="North Carolina",], aes(x = Primary_Reason, y = log_ppm, group = Primary_Reason)) + 
+ggplot(data = df[df$state=="Florida" | df$state=="Texas" | df$state=="North Carolina",],
+       aes(x = Primary_Reason, y = log_ppm, group = Primary_Reason)) + 
   geom_boxplot() +
@@
 ```{r}
-ggplot(data = df[df$state=="Florida" | df$state=="Texas" | df$state=="North Carolina",], aes(x = source, y = log_ppm, group = source)) + 
+ggplot(data = df[df$state=="Florida" | df$state=="Texas" | df$state=="North Carolina",],
+       aes(x = source, y = log_ppm, group = source)) + 
   geom_boxplot() +
@@
 ### City visualization for FL
-```{r}
+```{r, fig.height=10}
 ggplot(df[df$state=="Florida",], aes(x=log_ppm, y=reorder(city, city, length))) +
@@
 }
+par(mfrow=c(1,1))
 hist(group_int_list)

### 2021-03-22 final_report_2.Rmd

@@
 df$source <- droplevels(df$source)
-knitr::kable(table(df$source), col.names = c("Level", "Freq"))
+knitr::kable(table(df$source), 
+             col.names = c("Level", "Freq"),
+             caption = "Frequency Table for Source")
 ```
@@
 #print(reason_df)
-knitr::kable(reason_df_2, col.names = c("Level", "Freq"))
+knitr::kable(reason_df_2, 
+             col.names = c("Level", "Freq"),
+             caption = "Frequency Table for Primary Reason")
 ```
@@
 
+\newpage 
 # Appendix
@@
 ```{r}
-hist(df$date_num)
-knitr::kable(table(df$date_quarter))
+ggplot(df, aes(x=date_num)) + 
+ geom_histogram(aes(y=..density..), colour="black", fill="white", bins = 50)+
+ geom_density(alpha=.2, fill="blue") 
+knitr::kable(table(df$date_quarter), col.names = c("Date Quarter", "Freq"))
 ```
@@
 ```{r}
-hist(df$dose_per_mg)
+ggplot(df, aes(x=dose_per_mg)) + 
+ geom_histogram(aes(y=..density..), colour="black", fill="white", bins = 50)+
+ geom_density(alpha=.2, fill="red") 
 ```
@@
 ### City visualization for FL
-```{r}
+```{r, fig.height=6, fig.align='center'}
 ggplot(df[df$state=="Florida",], aes(x=log_ppm, y=reorder(city, city, length))) +

### 2021-03-22 final_report_2.Rmd
Formatting changes

--- 2021-03-22 final_report_2.Rmd
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, cache=TRUE, fig.height = 3)
+knitr::opts_chunk$set(echo = FALSE, cache=TRUE, fig.height = 3, tidy.opts=list(width.cutoff=60), tidy=TRUE)
 ```
@@
 # Fit 3 models grouping on different location specificity
-mod_s <- lmer(data=df, log_ppm ~ (1 |state) + date_num + date_quarter + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F)
-mod_c <- lmer(data=df, log_ppm ~ (1 |city) + date_num + date_quarter + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F)
-mod_r <- lmer(data=df, log_ppm ~ (1 |USA_region) + date_num + date_quarter + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F)
+mod_s <- lmer(data=df, log_ppm ~ (1 |state) + date_num + date_quarter + 
+                  dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F)
+mod_c <- lmer(data=df, log_ppm ~ (1 |city) + date_num + date_quarter + 
+                  dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F)
+mod_r <- lmer(data=df, log_ppm ~ (1 |USA_region) + date_num + date_quarter +
+                  dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F)
 ```
@@
 drop_date_num <- anova(mod_s,
-      lmer(data=df, log_ppm ~ (1 |state) + date_quarter + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F))
+      lmer(data=df, log_ppm ~ (1 |state) + date_quarter + dose_per_mg + 
+               bulk_purchase + Primary_Reason + source, REML=F))
 date_num_pval <- drop_date_num$`Pr(>Chisq)`[2]
 # Test date_quarter
-drop_quarter <- anova(lmer(data=df, log_ppm ~ (1 |state) + date_quarter + dose_per_mg + 
+drop_quarter <- anova(lmer(data=df, log_ppm ~ (1 |state) + 
+                        date_quarter + dose_per_mg +
                         bulk_purchase + Primary_Reason + source, REML=F),
-      lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F))
+                lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + 
+                         bulk_purchase + Primary_Reason + source, REML=F))
 quarter_pval <- drop_quarter$`Pr(>Chisq)`[2]
+
 # Test dose_per_mg
-drop_dose <- anova(lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F),
-      lmer(data=df, log_ppm ~ (1 |state) + bulk_purchase + Primary_Reason + source, REML=F))
+drop_dose <- anova(lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + 
+                        bulk_purchase + Primary_Reason + source, REML=F),
+                   lmer(data=df, log_ppm ~ (1 |state) + bulk_purchase 
+                        + Primary_Reason + source, REML=F))
 dose_pval <- drop_dose$`Pr(>Chisq)`[2]
+
 # Test bulk purchase
-drop_bulk <- anova(lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F),
-      lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + Primary_Reason + source, REML=F))
+drop_bulk <- anova(lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + 
+                        bulk_purchase + Primary_Reason + source, REML=F),
+      lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + 
+               Primary_Reason + source, REML=F))
 bulk_pval <- drop_bulk$`Pr(>Chisq)`[2]
+
 # Test primary reason
-drop_reason <- anova(lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F),
-      lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase + source, REML=F))
+drop_reason <- anova(lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + 
+                              bulk_purchase + Primary_Reason + source, REML=F),
+      lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + 
+               bulk_purchase + source, REML=F))
 reason_pval <- drop_reason$`Pr(>Chisq)`[2]
+
 # Test source
-drop_source <- anova(lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F),
-      lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase + Primary_Reason, REML=F))
+drop_source <- anova(lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + 
+                              bulk_purchase + Primary_Reason + source, REML=F),
+      lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + 
+               bulk_purchase + Primary_Reason, REML=F))
 source_pval <- drop_source$`Pr(>Chisq)`[2]

### 2021-03-22 final_report_2.Rmd
limitations edit

--- 2021-03-22 final_report_2.Rmd
@@
 States labeled simply as "USA" were recoded to "Unknown".
-Region was also sometimes labeled as uknown or other; these observations coincided with the unknown states (see appendix Table X).
+Region was also sometimes labeled as uknown or other; these observations coincided with the unknown states (see appendix).
 (The variable `city` as a grouping variable is addressed later in the report.)
@@
 
+Our analysis has several limitations.
+Our filtering of the `log_ppm` variable to non-extreme values must be kept in mind, as we are essentially modeling typical values only, so some unusual price values may not be predicted correctly by our model.
+Small sample sizes for some states are not ideal, although our hierarchical model structure helps to mitigate this issue.
+User-input data is naturally messy and should be taken with a grain of salt.
+Furthermore, our sample is essentially self-selected (people who choose to enter their data into the StreetRx web site) rather than randomly sampled, and this may affect the generalizability of our results.
+
 # Appendix
@@
 
-## Visualizations
+## Visualizations, etc.
+
+```{r}
+#the state-by-region table of coutns below
+table(df$state, df$USA_region)
+```
 

### 2021-03-22 final 3.Rmd
changes

@@
 ---
-title: "\vspace{-1.5cm} \textbf{STA610 Case Study 1 Team 4 Report}"
+title: STA610 Case Study 1 Team 4 Report
 author:
@@
 output:
+  pdf_document: default
   html_document:
     df_print: paged
-  pdf_document: default
 fontsize: 11pt
-geometry: margin=2cm
+geometry: margin=1.5cm
 ---
@@
 
-```{r, fig.align='center', fig.height = 2.5}
+```{r, fig.align='center', fig.height = 1.5}
 ggplot(df, aes(x=reorder(USA_region, USA_region, length), y=log_ppm, fill=USA_region)) +
@@
 #### Model Diagnostics
-```{r, fig.align='center', warning=FALSE, message=FALSE, error=FALSE}
+```{r error=FALSE, fig.align='center', fig.height=, message=FALSE, warning=FALSE, fig.height=2}
 resqq <- plot_resqq(mod_final)
@@
 
-```{r, fig.align='center', fig.height=8}
+```{r, fig.align='center', fig.height=4}
 plotREsim(REsim(mod_final,n.sims=100),stat='median',sd=TRUE, labs = TRUE)
@@
 
+## Limitation
+
+Our analysis has several limitations.
+Our filtering of the `log_ppm` variable to non-extreme values must be kept in mind, as we are essentially modeling typical values only, so some unusual price values may not be predicted correctly by our model.
+Small sample sizes for some states are not ideal, although our hierarchical model structure helps to mitigate this issue.
+User-input data is naturally messy and should be taken with a grain of salt.
+Furthermore, our sample is essentially self-selected (people who choose to enter their data into the StreetRx web site) rather than randomly sampled, and this may affect the generalizability of our results.
+
+\newpage
 ## Appendix
@@
 ```
-
-# Appendix
 
@@
 ### More EDA: slope and intercept distributions for predictors by state
-```{r}
+```{r, fig.width=3}
 # To get a better sense of how state-level groups and slopes are distributed, we created the following intercepts for our coninuous variables:
@@
 }
-par(mfrow=c(1,1))
+
 hist(group_int_list)

### 2021-03-22 final 3.Rmd
Added FINAL writeup

@@
 ---
-title: STA610 Case Study 1 Team 4 Report
+title: \vspace{-1.5cm} STA610 Case Study 1 Team 4 Report
 author:
@@
 library(pacman)
-library(merTools)
+devtools::install_github("goodekat/redres")
 pacman::p_load(tidyverse, lme4, gridExtra, grid, ggplot2, lattice, redres, stringr,
-               influence.ME, knitr)
-# devtools::install_github("goodekat/redres")
+               influence.ME, knitr, merTools)
 ```
@@
 States labeled simply as "USA" were recoded to "Unknown".
-Region was also sometimes labeled as uknown or other; these observations coincided with the unknown states (see appendix Table X).
+Region was also sometimes labeled as uknown or other; these observations coincided with the unknown states.
 (The variable `city` as a grouping variable is addressed later in the report.)
@@
 df$date_num <- as.numeric(as.Date(df$price_date, "%m/%d/%y")) - 14621 # sets 1/12/2010 (first post-2010 date) as zero
-tempdf <- df %>% select(date_num, price_date)
+tempdf <- df %>% dplyr::select(date_num, price_date)
 # this filters out all data before the 1/1/2000, which probably shouldn't be included
@@
 
-```{r, fig.align='center', fig.height=4}
+```{r, fig.align='center', fig.height=6, fig.width=8}
 plotREsim(REsim(mod_final,n.sims=100),stat='median',sd=TRUE, labs = TRUE)
 ```
-From the result of the random effects, we can see that the random intercept for Tennessee is the highest and it is significantly different from zero.
-The random intercept for Arizona is the lowest and this intercept is also significant.
+
+The value of the random intecepts range from exp(0.450007131) = 1.568323 for Tenessee to exp(-0.417349014) = 0.658791 for Nevada.
+We interpret this as follows: the estimated average value for ppm of oxymorphone is highest for Tennessee.
+The price is estimated assuming a non-bulk purchase, unknown source, blank reason, and zero dose per mg (which is obviously not a realistic value, but is noted for the sake of interpretation).
 
@@
 ```{r, ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE}
-```
-
-## Code
-
-```{r, ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE}
-
 ```
@@
 
+par(mfrows=c(2, 1))
 hist(group_int_list)
